### 2017-08-17 deeplearn/presentation/paper.tex
Presentation draft paper

\documentclass[10pt]{amsart}
\usepackage{amsmath}

\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathcal{C}}

\begin{document}
\title{Recurrent Neural Networks}
\author{Jean-Martin Albert}
\date{\today}
\maketitle

\section{The basic problem}

Let $U=\R^n$ and $V=\R^m$ be vector spaces, and consider a function $f:U\to W$.
In theory, $f$ here can be any set theoretic function, but for our purpose, we
will assume that $f$ is at the very least square integrable. In general, we
consider a class $\C$ of functions $U\to V$, and try to find an element $g\in\C$
which is as close as possible to $f$.  In order to do so, we need a notion of distance
between functions.  There are many possible choices, but some of the most popular are
the $L^p$ norms and the $L^\infty$ norm.

While the class $\C$ van be virtually any class of functions, we usually choose a class
that is somewhat definable, in the sense that there is a (differentiable) function $F:U\times\R^k\to V$
such that $\C=\{F(x, w): w\in \R^k\}$.  Such classes of functions have the advantage of
being equicontinuous.

%\begin{example}
Let $f(x)=x(x-1)(x+1)$.  As an example, we will try to fit a linear function
$\ell(x)=ux+v$ to $f$.  The class $\C$ is defined by $\C=\{ux+v:u,v\in R\}$.
For an element $g(x, u, v)\in\C$, we have $$f(x)-g(x,u,v) = x^3-x - ux+v$$
From this we get
\begin{eqnarray*}
  (f(x)-g(x,u,v))^2 &=& (x^3 - (1-u)x + v)^2 \\
  {} &=& x^6 - 2(1-u)x^4 + 2vx^3 + (1-u)^2x^2 - \\
  {} & & 2(1-u)vx + v^2
\end{eqnarray*}
If we compute $\int_{-1}^1 (f(x)-g(x,u,v))^2$, we get

\begin{eqnarray*}
  \int_{-1}^1(f(x)-g(x; u,v))^2 &=& \int_{-1}^1(x^3 - (1-u)x + v)^2 \\
  {} &=& \int_{-1}^1x^6 - 2(1-u)x^4 + 2vx^3 + (1-u)^2x^2 - 2(1-u)vx + v^2\\
  {} &=& \frac{x^7}{7}-2v\frac{x^4}{4} + (1-u)^2\frac{x^3}{3} - 2(1-u)v\frac{x^2}{2}+v^2x\\
  {} &=& \frac{1}{7}-2v\frac{1}{4} + (1-u)^2\frac{1}{3} - 2(1-u)v\frac{1}{2}+v^2\\
  {} & & +\frac{1}{7}-2v\frac{1}{4} + (1-u)^2\frac{-1}{3} - 2(1-u)v\frac{1}{2}-v^2\\
  {} &=& \frac{2}{7} - (2-u)v+v^2 = \ell(u,v)
\end{eqnarray*}
and note that the last line above is a function of $u$ and $v$ alone, and we have to minimize it.
%\end{example}

\section{Dealing with sequences}

\end{document}

### 2017-08-18 deeplearn/presentation/paper_2.tex
Presentation files

\documentclass[10pt]{amsart}
\usepackage{amsmath}

\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathcal{C}}

\begin{document}
\title{Recurrent Neural Networks}
\author{Jean-Martin Albert}
\date{\today}
\maketitle

\section{Dealing with sequences}

There is more than one way to extend a recurrent function to a set of sequences

\section{Back Propagation}

In general, in order to minimize a function $f(w)$ of one variable, we make use of the fact that
the extreme values of $f$ all have $f'(x)=0$.  If $f$ is a function of two or more variables, then
if $x$ is such that $f(x)$ is a local max, or minimum, then all partial derivatives
$\frac{\partial f}{\partial w_i}(x)=0$.  If $f:\R^n\to\R$ is a function, then the
{\em gradient} of $f$ is the vector
$\nabla f(x) = \left(\frac{\partial f}{\partial w_1}(x),...,\frac{\partial f}{\partial w_n}(x)\right)$.
One interesting property of $\nabla f(x)$ is that, as a vector, it ``points'' in the direction in which
$f$ decreases the most.  That is to say, there is a good chance that $f(x-\alpha\nabla f(x))\leq f(x)$.
The chance is better the smaller $\alpha$ is.  We see that if we iterate that process, we get a
sequence $x_1,...,x_n,...$ such that $f(x_{i+1})\leq f(x_i)$ for every $i$.  Since $m\leq f(x_i)\leq f(x_0)$
for every $i$, we get a bounded decreasing sequence of real number, which converges.

Caveat: this ideal situation doesn't allways happen.

This iterative process is {\em (Stocastic) gradient descent}.

Suppose $f$ is a complsition of two functions, and that we can write $f(x,w)= g(h(x, w_1), w_2)$.
Then we can write $\nabla_w f(x, w)=\nabla g(h(x, w_1), w_2)\cdot \nabla h(x, w_1)$. Each level
of composition corresponds to a layer of neural network.  The chain rule transforms function composition
into a product. If a network becomes deep, then the derivative of the loss function becomes a long
product.  When gradients become small in norm (like they do when we approach a min), then the gradient
becomes very close to $0$, and the update rule for gradient descent stops changing the weights.

This is the vanishing gradient problem for very deep networks, and makes convergence slow.

\section{The base of recurrent networks}

Consider a function $f:U\times S\to V\times S$, where $U$, $V$ and $S$ are finite dimensional vector spaces.
Note that every vector space has a a distinguished element $0$, the zero-vector.  Let $u_1,...,u_n$ be a finite
sequence of vectors in $U$.  We construct a sequence of vectors in $V$ as follows. Write $f(u_1, 0)=(v_1, s_1)$,
and for every $i$, if $f(u_{i}, s_{i-1})=(v_i, s_i)$. Note that this is a state machine, close to a
deterministic finite automaton, except that here the state space $S$ is infinite.

The set $\{0,1\}^*$ of all finite sequences of $0$'s and $1$'s is countable, and $S$, being a real vector space,
is uncountable.  Therefore, we can encode every element $w\in\{0,1\}^*$ as a vector in $S$.  Informally,
$S$ is enough to encode any finite state space, and every possible value for the content of the tape of a
Turing machine. We get:

%\begin{theorem}
Recurrent neural networks are Turing-complete.
%\end{theorem}

which explains why recurrent neural networks seem to be able to produce results that other networks can't.
Training a recurrent network is the same as producing a Turing machine.

\section{Types of Recurrent Cells}

\subsection{Basic Recurrent Cell}

\subsection{Long Short-term Memory Cell}

\subsection{Gated Recurrent Unit Cell}

\section{Examples}

\subsection{Many-to-One}

The Buzzometer sentiment analysis tool uses a bi-directional GRU model.

\subsection{Many-to-many}

Text generator

\subsection{Many-to-one-to-many}

As an example of a recurrent neural network performing a task that ordinary neural network
should not be able to perform, we present a sorting function which uses GRU cells. The architecture
of the network is the reason we call it a many-to-one-to-many network.  One level takes a list
of numbers, and produces a vector which we can see as representing all the numbers in the list,
and a second layer uses this output vector as input, and produces the original list, sorted in
increasing order. The training was done using sequences of natural numbers $n\in\{1,...,32\}$, and each
of the training sequences of length $32$.  After about 6 days of training, the accuracy of the network
stabilized at around 95\%, and it is worth noting that eventhough it took a long time to breach
the 90\% accuracy mark, very early on in the training did the network output increasing sequences.

% \section{The basic problem}
%
% Let $U=\R^n$ and $V=\R^m$ be vector spaces, and consider a function $f:U\to W$.
% In theory, $f$ here can be any set theoretic function, but for our purpose, we
% will assume that $f$ is at the very least square integrable. In general, we
% consider a class $\C$ of functions $U\to V$, and try to find an element $g\in\C$
% which is as close as possible to $f$.  In order to do so, we need a notion of distance
% between functions.  There are many possible choices, but some of the most popular are
% the $L^p$ norms and the $L^\infty$ norm.
%
% While the class $\C$ van be virtually any class of functions, we usually choose a class
% that is somewhat definable, in the sense that there is a (differentiable) function $F:U\times\R^k\to V$
% such that $\C=\{F(x, w): w\in \R^k\}$.  Such classes of functions have the advantage of
% being equicontinuous.
%
%
% %\begin{example}
% Let $f(x)=x(x-1)(x+1)$.  As an example, we will try to fit a linear function
% $\ell(x)=ux+v$ to $f$.  The class $\C$ is defined by $\C=\{ux+v:u,v\in R\}$.
% For an element $g(x, u, v)\in\C$, we have $$f(x)-g(x,u,v) = x^3-x - ux+v$$
% From this we get
% \begin{eqnarray*}
%   (f(x)-g(x,u,v))^2 &=& (x^3 - (1-u)x + v)^2 \\
%   {} &=& x^6 - 2(1-u)x^4 + 2vx^3 + (1-u)^2x^2 - \\
%   {} & & 2(1-u)vx + v^2
% \end{eqnarray*}
% If we compute $\int_{-1}^1 (f(x)-g(x,u,v))^2$, we get
%
% \begin{eqnarray*}
%   \int_{-1}^1(f(x)-g(x; u,v))^2 &=& \int_{-1}^1(x^3 - (1-u)x + v)^2 \\
%   {} &=& \int_{-1}^1x^6 - 2(1-u)x^4 + 2vx^3 + (1-u)^2x^2 - 2(1-u)vx + v^2\\
%   {} &=& \frac{x^7}{7}-2v\frac{x^4}{4} + (1-u)^2\frac{x^3}{3} - 2(1-u)v\frac{x^2}{2}+v^2x\\
%   {} &=& \frac{1}{7}-2v\frac{1}{4} + (1-u)^2\frac{1}{3} - 2(1-u)v\frac{1}{2}+v^2\\
%   {} & & +\frac{1}{7}-2v\frac{1}{4} + (1-u)^2\frac{-1}{3} - 2(1-u)v\frac{1}{2}-v^2\\
%   {} &=& \frac{2}{7} - (2-u)v+v^2 = \ell(u,v)
% \end{eqnarray*}
% and note that the last line above is a function of $u$ and $v$ alone, and we have to minimize it.
% %\end{example}
%
%
% \section{Dealing with sequences}

\end{document}

### 2017-08-18 deeplearn/twitter_sentiment/models/text_generator/text_generator.py

@@ -178,7 +178,7 @@ def rnn_minibatch_sequencer(raw_data, batch_size, sequence_size, nb_epochs):
                    'train_y':  y,
                    'validate': None,
                    'batch_number':  batch,
-                   'epoch_number':  epoch,
+                   'epoch_number':  epoch+1,
                    'batch_index':   I,
                    'total_batches': total_num_batches,
                    'total_epochs':  nb_epochs}
@@ -186,7 +186,7 @@ def rnn_minibatch_sequencer(raw_data, batch_size, sequence_size, nb_epochs):
 
 def generate_text(length, session=None):
     generated_text = ''
-    character = [[ord('Z')]]
+    character = [[ord(' ')]]
     istate = np.zeros([Hyperparameters.n_layers, 1, Hyperparameters.hidden_states])
     while len(generated_text) < length:
         feed_dict = {Globals.model_input: character, Globals.initial_state: istate}
@@ -204,7 +204,14 @@ def generate_text(length, session=None):
 if __name__ == '__main__':
     root = os.path.dirname(__file__)
     root = os.path.join(root, 'text_gen_data')
-    data = read_data_files_from_folder(os.path.join(root, 'sicp-code-master/*.scm'))
+    save = os.path.join(os.path.expanduser('~'), '.text_gen', 'output')
+    chkpt = os.path.join(os.path.expanduser('~'), '.text_gen', 'model_chkpt')
+    if not os.path.exists(save):
+        os.makedirs(save)
+    if not os.path.exists(chkpt):
+        os.makedirs(chkpt)
+
+    data = read_data_files_from_folder(os.path.join(root, 'harry_potter/*.txt'))
 
     inference()
     train()
@@ -221,6 +228,12 @@ if __name__ == '__main__':
             _, ostate = session.run([Globals.minimize, Globals.state], feed_dict=feed_dict)
             print('x', end='', flush=True)
             istate = ostate
-            if index % 50 == 0:
-                print(generate_text(2000, session))
+            if index % batch['epoch_number']*150 == 0:
+                file_name = "sort-test-{}".format(index)
+                file_name = os.path.join(save, file_name)
+                with open(file_name, 'w') as test_file:
+                    test_file.write(generate_text(20000, session))
+            if index % 250 == 0:
+                s = tf.train.Saver()
+                s.save(session, os.path.join(chkpt, "t_gen_weights"))
             index += 1

### 2017-08-21 deeplearn/presentation/paper_2.tex
Presentation work

@@
 \usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{tikz}
+\usetikzlibrary{shapes.geometric, arrows}
+
+\usepackage{minted}
 
@@
 \newcommand{\C}{\mathcal{C}}
+
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{proposition}[theorem]{Proposition}
+\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{corollary}[theorem]{Corollary}
+
+\theoremstyle{definition}
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem{example}[theorem]{Example}
 
@@
 
-\section{Dealing with sequences}
+%\def\layersep{2.5cm}
 
-There is more than one way to extend a recurrent function to a set of sequences
+% \newcommand{\nodedef}{\tikzstyle{arrow} = [thick,->,>=stealth]
+%     \node (input) [] at (0cm,1cm) {};
+%     \node (output) [] at (0cm,-1cm) {};
+%     \node (s-input) [] at (-2cm,0cm) {};
+%     \node (s-output) [] at (2cm,0cm) {};
+%     \path[yshift=0.5cm]
+%         node[startstop] (N) at (0cm,-.5cm) {};
+%
+%     \draw [arrow] (input) -- (N);
+%     \draw [arrow] (s-input) -- (N);
+%     \draw [arrow] (N) -- (output);
+%     \draw [arrow] (N) -- (s-output);}
+%
+%
+% \newcommand{\nodedefinner}{\tikzstyle{arrow} = [thick,->,>=stealth]
+%     \node (input) [] at (0cm,1cm) {};
+%     \node (output) [] at (0cm,-1cm) {};
+%     \node (s-input) [] at (-2cm,0cm) {};
+%     \node (s-output) [] at (2cm,0cm) {};
+%     \path[yshift=0.5cm]
+%         node[startstop] (N) at (0cm,-.5cm) {};
+%
+%     %\draw [arrow] (input) -- (N);
+%     %\draw [arrow] (s-input) -- (N);
+%     \draw [arrow] (N) -- (output);
+%     \draw [arrow] (N) -- (s-output);}
+
+% \begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
+%     \tikzstyle{arrow} = [thick,->,>=stealth]
+%     \tikzstyle{startstop} = [rectangle, minimum width=2cm, minimum height=.75cm, text centered, draw=black, fill=black!20]
+%     \node (input) [] at (0cm,1cm) {};
+%     \node (output) [] at (0cm,-1cm) {};
+%     \node (s-input) [] at (-2cm,0cm) {};
+%     \node (s-output) [] at (2cm,0cm) {};
+%     \path[] node[startstop] (N) at (0cm,0cm) {};
+%     \draw [arrow] (input) -- (N);
+%     \draw [arrow] (s-input) -- (N);
+%     \draw [arrow] (N) -- (output);
+%     \draw [arrow] (N) -- (s-output);
+% \end{tikzpicture}
+%
+%
+
+ \section{Introduction}
+
+ Let $U=\R^n$ and $V=\R^m$ be vector spaces, and consider a function $f:U\to W$.
+ In theory, $f$ here can be any set theoretic function, but for our purpose, we
+ will assume that $f$ is at the very least square integrable. In general, we
+ consider a class $\C$ of functions $U\to V$, and try to find an element $g\in\C$
+ which is as close as possible to $f$.  In order to do so, we need a notion of distance
+ between functions.  There are many possible choices, but some of the most popular are
+ the $L^p$ norms and the $L^\infty$ norm.
+
+ While the class $\C$ van be virtually any class of functions, we usually choose a class
+ that is somewhat definable, in the sense that there is a (differentiable) function $F:U\times\R^k\to V$
+ such that $\C=\{F(x, w): w\in \R^k\}$.  Such classes of functions have the advantage of
+ being equicontinuous.
+
+ \begin{example}
+ Let $f(x)=x(x-1)(x+1)$.  As an example, we will try to fit a linear function
+ $\ell(x)=ux+v$ to $f$.  The class $\C$ is defined by $\C=\{ux+v:u,v\in R\}$.
+ For an element $g(x, u, v)\in\C$, we have $$f(x)-g(x,u,v) = x^3-x - ux+v$$
+ From this we get
+ \begin{eqnarray*}
+   (f(x)-g(x,u,v))^2 &=& (x^3 - (1-u)x + v)^2 \\
+   {} &=& x^6 - 2(1-u)x^4 + 2vx^3 + (1-u)^2x^2 - \\
+   {} & & 2(1-u)vx + v^2
+ \end{eqnarray*}
+ If we compute $\int_{-1}^1 (f(x)-g(x,u,v))^2$, we get
+
+ \begin{eqnarray*}
+   \int_{-1}^1(f(x)-g(x; u,v))^2 &=& \int_{-1}^1(x^3 - (1-u)x + v)^2 \\
+   {} &=& \int_{-1}^1x^6 - 2(1-u)x^4 + 2vx^3 + (1-u)^2x^2 - 2(1-u)vx + v^2\\
+   {} &=& \frac{x^7}{7}-2v\frac{x^4}{4} + (1-u)^2\frac{x^3}{3} - 2(1-u)v\frac{x^2}{2}+v^2x\\
+   {} &=& \frac{1}{7}-2v\frac{1}{4} + (1-u)^2\frac{1}{3} - 2(1-u)v\frac{1}{2}+v^2\\
+   {} & & +\frac{1}{7}-2v\frac{1}{4} + (1-u)^2\frac{-1}{3} - 2(1-u)v\frac{1}{2}-v^2\\
+   {} &=& \frac{2}{7} - (2-u)v+v^2 = \ell(u,v)
+ \end{eqnarray*}
+ and note that the last line above is a function of $u$ and $v$ alone, and we have to minimize it.
+ \end{example}
 
@@
 
+\section{Dealing with sequences}
+
+Let $A$ be a finite set.
+We define the set $A^*$ of all finite sequences of elements of $A$ to be the smallest set $X$ with the property that the empty sequence $\epsilon$ is in $X$, and whenever $w\in X$ and $a\in A$, the sequence $aw\in X$.
+If $A$ and $B$ are two finite sets, and $f:A\to B$ is any function, then we can extend $f$ extends uniquely to a function $f^*:A^*\to B^*$, which is defined by $f(a_1a_2,...,a_n)=f(a_1)f(a_2)\cdots f(a_n)$.
+This $f^*$ is the only function extending $f$ which satisfies $f(vw)=f(v)f(w)$.
+
+Let $S$ be another finite set with a distinguished element $\perp\in S$, and let $f:A\times S\to B\times S$ be a function.
+We can use $f$ to define a function $f^*:A^*\to B^*$ in several ways.
+Consider an auxiliary function $f^*:A^*\to B^*\times S$ by defining First we define $f^*(\epsilon)=(\epsilon, \perp)$.
+If $a\in A$ then $f^*(a)=f(a, \perp)$.
+If $v=aw$, then by induction we can compute $f^*(w)=(u, s)$.
+Write $f(a, s)=(b, s')$, and define $f^*(aw)=(bu, s')$.
+We can produce a function $f^*:A^*\to B^*$ by taking $f^*(w)=\pi_2 g(w)$.
+
+We can use $f$ to define a function $f^*:A^*\to B$.
+Consider an auxiliary function $f^*:A^*\to B\times S$ by defining First we define $f^*(\epsilon)=(b_0, \perp)$.
+If $a\in A$ then $f^*(a)=f(a, \perp)$.
+If $v=aw$, then by induction we can compute $f^*(w)=(u, s)$.
+Write $f(a, s)=(b, s')$, and define $f^*(aw)=(b, s')$.
+We can produce a function $f^*:A^*\to B$ by taking $f^*(w)=\pi_2 g(w)$.
+
+Let $S$ be another finite set with a distinguished element $\perp\in S$, and let $f:A\times S\to B\times S$ be a function.
+We can use $f$ to define a function $f^*:A\to B^N$ for every $N$ in several ways.
+Consider an auxiliary function $f^*:A\to B^N\times S$ by defining.
+If $a\in A$ then $f^*(a)_1=f(a, \perp)$.
+Suppose $f^*(a)_n=(w, s)$, and  $f(a, s) = (b, s')$, then $f^*(a)_{n+1}=(bw, s')$.
+This is an example of a one-to-many function.
+If $A$ has a distinguished element $a_0$, then we can define $f^*(a)_1=f(a, \perp)$, and $f^*(a)_n=(w, s)$, and  $f(a_0, s) = (b, s')$, then $f^*(a)_{n+1}=(bw, s')$.
+This is an example of a one-to-many function.
+Note that this is a special case of the many to many function, where $f^*(a)$ is just defined as $f^*(aaa\cdots a)$, or $f^*(aa_0\cdots a_0)$.
+
+%If $v=aw$, then by induction we can compute $f^*(w)=(u, s)$. Write $f(a, s)=(b, s')$, and define $f^*(aw)=(bu, %s')$.  We can produce a function $f^*:A^*\to B^*$ by taking $f^*(w)=\pi_2 g(w)$.
+%  We can then define $f^*(v)=f^*(aw)=f(a, s) = (w', s')$, or mode formally, $f^*(aw)=f(a, \pi_2f^*(w)$
+%
+%$f^*(v)=f^*()
+%
+%There is more than one way to extend a recurrent function to a set of sequences
+
 \section{The base of recurrent networks}
@@
 
-%\begin{theorem}
+\begin{theorem}
 Recurrent neural networks are Turing-complete.
-%\end{theorem}
+\end{theorem}
 
@@
 
+\begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
+    \tikzstyle{arrow} = [thick,->,>=stealth]
+    \tikzstyle{startstop} = [rectangle, minimum width=2cm, minimum height=.75cm, text centered, draw=black, fill=black!20]
+    \node (input) [] at (0cm,1cm) {};
+    \node (output) [] at (0cm,-1cm) {};
+    \node (s-input) [] at (-2cm,0cm) {};
+    \node (s-output) [] at (2cm,0cm) {};
+    \path[] node[startstop] (N) at (0cm,0cm) {};
+    \draw [arrow] (input) -- (N);
+    \draw [arrow] (s-input) -- (N);
+    \draw [arrow] (N) -- (output);
+    \draw [arrow] (N) -- (s-output);
+\end{tikzpicture}
+
+\begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
+    \tikzstyle{arrow} = [thick,->,>=stealth]
+    \tikzstyle{startstop} = [rectangle, minimum width=1.5cm, minimum height=.6cm, text centered, draw=black, fill=black!20]
+    \node (input) [] at (0cm,1cm) {};
+    \node (output) [] at (0cm,-3cm) {};
+    \node (s-input-A) [] at (-2cm,0cm) {};
+    \node (s-output-A) [] at (2cm,0cm) {};
+    \node (s-input-B) [] at (-2cm,-1cm) {};
+    \node (s-output-B) [] at (2cm,-1cm) {};
+    \node (s-input-C) [] at (-2cm,-2cm) {};
+    \node (s-output-C) [] at (2cm,-2cm) {};
+    \path[] node[startstop] (A) at (0cm,0cm) {};
+    \path[] node[startstop] (B) at (0cm,-1cm) {};
+    \path[] node[startstop] (C) at (0cm,-2cm) {};
+    \draw [arrow] (input) -- (A);
+    \draw [arrow] (s-input-A) -- (A);
+    \draw [arrow] (s-input-B) -- (B);
+    \draw [arrow] (s-input-C) -- (C);
+    \draw [arrow] (A) -- (s-output-A);
+    \draw [arrow] (B) -- (s-output-B);
+    \draw [arrow] (C) -- (s-output-C);
+    \draw [arrow] (A) -- (B);
+    \draw [arrow] (B) -- (C);
+    \draw [arrow] (C) -- (output);
+\end{tikzpicture}
+
+\section{Backward Propagation Through Time}
+
+How do we train recurrent neural networks?
+we can use back propagation, just like a regular neural network.
+Heuristically, we unroll the recurrent network ``infinitely'' many times, until it looks like an ordinary very deep neural network.
+In practice, since computers only have a finite amount of resources, we only unroll the network a large but finite number of times, and treat it like an ordinary neural network.
+
+\begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
+    \tikzstyle{arrow} = [thick,->,>=stealth]
+    \tikzstyle{startstop} = [rectangle, minimum width=1.5cm, minimum height=.6cm, text centered, draw=black, fill=black!20]
+    \node (input) [] at (0cm,1cm) {};
+    \node (output) [] at (0cm,-1cm) {};
+    \node (s-input-A) [] at (-2cm,0cm) {};
+    \node (s-output-A) [] at (10cm,0cm) {};
+    %\node (s-input-B) [] at (-2cm,-1cm) {};
+    %\node (s-output-B) [] at (10cm,-1cm) {};
+    %\node (s-input-C) [] at (-2cm,-2cm) {};
+    %\node (s-output-C) [] at (10cm,-2cm) {};
+    \path[] node[startstop] (A-0) at (0cm,0cm) {};
+    %\path[] node[startstop] (B-0) at (0cm,-1cm) {};
+    %\path[] node[startstop] (C-0) at (0cm,-2cm) {};
+    \draw [arrow] (input) -- (A);
+    \draw [arrow] (s-input-A) -- (A-0);
+    %\draw [arrow] (s-input-B) -- (B-0);
+    %\draw [arrow] (s-input-C) -- (C-0);
+    %\draw [arrow] (A) -- (B);
+    %\draw [arrow] (B) -- (C);
+    \draw [arrow] (A) -- (output);
+
+     \foreach \i/\j in {1/0,2/1,3/2,4/3}
+     {
+       \node (input-\i) [] at (2*\i cm,1cm) {};
+       \node (output-\i) [] at (2*\i cm,-1cm) {};
+       \path[] node[startstop] (A-\i) at (2*\i cm,0cm) {};
+    %   \path[] node[startstop] (B-\i) at (2*\i cm,-1cm) {};
+  %     \path[] node[startstop] (C-\i) at (2*\i cm,-2cm) {};
+  %     \draw [arrow] (A-\i) -- (B-\i);
+%       \draw [arrow] (B-\i) -- (C-\i);
+       \draw [arrow] (input-\i) -- (A-\i);
+       \draw [arrow] (A-\j) -- (A-\i);
+%       \draw [arrow] (B-\j) -- (B-\i);
+%       \draw [arrow] (C-\j) -- (C-\i);
+       \draw [arrow] (A-\i) -- (output-\i);
+     }
+     \draw [arrow] (A-4) -- (s-output-A);
+     %\draw [arrow] (B-4) -- (s-output-B);
+     %\draw [arrow] (C-4) -- (s-output-C);
+
+\end{tikzpicture}
+
 \section{Types of Recurrent Cells}
 
-\subsection{Basic Recurrent Cell}
+There are three main types of recurrent cells which are used to build recurrent neural networks.
+Let us describe them, and give some of their basic properties, along with a na\"ive implementation in tensorflow.
 
-\subsection{Long Short-term Memory Cell}
+\subsection{Basic Recurrent Cell} We begin with the most basic of recurrent cell.
+Abstractly, a function $f:U\times S\to V\times S$ can be defined using two functions $u:\R^n\times \R^\ell\to \R^n$ and $v:\R^n\times\R^\ell \to \R^\ell$.
+We se ethe former function as providing the output, and the latter function as a state updating function.
+The most common definition for $u$ and $v$ for basic recurrent cells is as follows: $u(x, s) = f(A_ux+B_us)$, where $A$ nad $B$ are matrices, and $f$ is a non-linear function, and $v(x, s) = \tanh(A_vx+B_vs)$.
+Te parameters in this definition are $A_u$, $B_u$, $A_v$ and $B_v$.
+In tehsorflow, the code looks like:
 
-\subsection{Gated Recurrent Unit Cell}
+\begin{minted}{python}
+def basic_rnn_cell(input_tensor, state_tensor):
+  A_u = tf.Variable(shape=[N, L])
+  B_u = tf.Variable(shape=[L, L])
+  A_v = tf.Variable(shape=[N, L])
+  B_v = tf.Variable(shape=[L, L])
+  output_tensor = tf.relu(tf.matmul(input_tensor, A_u) +
+                            tf.matmul(state_tensor, B_u))
+  new_state_tensor = tf.tanh(tf.matmul(input_tensor, A_v) + \
+                                  tf.matmul(state_tensor, B_v))
+  return output_tensor, new_state_tensor
+\end{minted}
+
+\subsection{Long Short-term Memory Cell} There are a few problems with the architecture described above.
+First, it suffers greatly from the vanishing gradient problem, since there is no way to prevent gradients from becoming very small.
+Secondly, simple recurrent networks have a hard time remembering facts about the input sequence.
+To remedy this situation, long short-term memory cells were introduced.
+The overall structure of an LSTM cell is almost the same as the basic RNN cell.
+This time we take the previous output into account, which gives the structure as $f:\R^n\times \R^m\times \R^{\ell}\to \R^m\times \R^{\ell}$ which can be divided as two function $u:\R^{n}\times\R^m\times \R^{\ell}\to \R^m$ and $v:\R^n\times \R^m\times\R^{\ell}\to \R^{\ell}$.
+To define the functione $u$ and $v$, we use auxiliary functions $F$, $I$, and $O$ defined as follows
+\begin{eqnarray*}
+  F(x, y) &=& \sigma(A_Fx + B_Fy + b_F)\\
+  I(x, y) &=& \sigma(A_Ix + B_Iy + b_I)\\
+  O(x, y) &=& \sigma(A_Ox + B_Oy + b_O)\\
+  S(x, y) &=& \sigma(A_Ox + B_Oy + b_O)\\
+\end{eqnarray*}
+The state update is given by $v(x, y, s) = F(x, y)\circ s + I(x, y)\circ \sigma_v(A_vx + B_vy + b_v)$, where $\circ$ denotes pointwise multiplication of vectors, and finally, the output can be defined as $u(x, y, s)=O(x, y)\circ \sigma_u(v(x, y, s))$.
+
+\begin{minted}{python}
+def lstm_gate(input_tensor, previous_output, port_op):
+  A = tf.Variable(shape=[N, L])
+  B = tf.Variable(shape=[L, L])
+  b = tf.Variable(shape=[L, L])
+  x = tf.matmul(input_tensor, A)+ tf.matmul(previous_output, B) + b
+  return post_op(x)
+
+def lstm_cell(input_tensor, output, state):
+  F = lstm_gate(input_tensor, output, tf.sigmoid)
+  I = lstm_gate(input_tensor, output, tf.sigmoid)
+  O = lstm_gate(input_tensor, output, tf.sigmoid)
+  S = lstm_gate(input_tensorm output, tf.tanh)
+  new_state = tf.mul(output, F) + tf.mul(I, S)
+  output = tf.mul(O, tf.tanh(new_state))
+  return output, new_state
+\end{minted}
+
+%The function $u$ is defined by $u(x, o, s)=o()$
+
+%: a function $f:\R^n\times \R^{\ell}\to \R^m\times \R^{\ell}$, which can be divided as two function $u:\R^{n}\times \R^{\ell}\to \R^m$ and $v:\R^n\times\R^{\ell}\to \R^{\ell_1}$.
+We can se the function $u$, like before, as providing the output.
+The function $v$ can be seen as the state update function.
+The function $u$ is defined by $u(x, s)=\sigma(A_U x+)$
+
+% above, but the state space is further divided as $\R^\ell=\R^{\ell_1}\times\R^{\ell_2}$, which gives the cell the structure of a function $f:\R^n\times \R^{\ell_1}\times\R^{\ell_2}\to \R^m\times \R^{\ell_1}\times\R^{\ell_2}$, which can be divided as three function $u:\R^{n}\times \R^{\ell_1}\times\R^{\ell_2}\to \R^m$,  $v:\R^n\times\R^{\ell_1}\times\R^{\ell_2} \to \R^{\ell_1}$ and $w:\R^n\times\R^{\ell_1}\times\R^{\ell_2} \to \R^{\ell_2}$.
+We can se the function $u$, like before, as providing the output.
+The function $v$ can be seen as the state update function, and the function $w$ controls a memory tape.
+
+\subsection{Gated Recurrent Unit Cell} A common variant of the long short term memory cell is the {\em gated recurrent unit cell}, more commonly known as GRU cells.
+The philosophy behind their design is similar to the long short term memory.
+Once again, each step of the computation takes into account a state vector and the output of the previous iteration.
+GRU's are functions  $f:\R^n\times \R^m\times \R^{\ell}\to \R^m\times \R^{\ell}$,  which can be divided as two function $u:\R^{n}\times\R^m\times \R^{\ell}\to \R^m$ and $v:\R^n\times \R^m\times\R^{\ell}\to \R^{\ell}$.
+To define the functione $u$ and $v$, we use auxiliary functions $U$ and  $R$ defined as follows
+\begin{eqnarray*}
+  U(x, y) &=& \sigma(A_Ux + B_Uy + b_U)\\
+  R(x, y) &=& \sigma(A_Rx + B_Ry + b_R)\\
+\end{eqnarray*}
+The state update is given by $v(x, y, s) = U(x, y)\circ s + (1-s)\circ \sigma_h(A_v x + B_v(R(x, y)\circ y) + b_v)$.
+
+\begin{minted}{python}
+def gru_gate(input_tensor, previous_output, port_op):
+  A = tf.Variable(shape=[N, L])
+  B = tf.Variable(shape=[L, L])
+  b = tf.Variable(shape=[L, L])
+  x = tf.matmul(input_tensor, A)+ tf.matmul(previous_output, B) + b
+  return post_op(x)
+
+def gru_cell(input_tensor, output, state):
+  U = gru_gate(input_tensor, output, tf.sigmoid)
+  R = gru_gate(input_tensor, output, tf.sigmoid)
+  O = gru_gate(input, tf.mul(R, output))
+  return tf.mul(R, output) + tf.mul((1-R)O)
+\end{minted}
+
+%$$
+%$$
+%
+%I(x, y)\circ \sigma_v(A_vx + B_vy + b_v)$, where $\circ$ denotes pointwise multiplication of vectors, and finally, the output can be defined as $u(x, y, s)=O(x, y)\circ \sigma_u(v(x, y, s))$.
 
@@
 
-The Buzzometer sentiment analysis tool uses a bi-directional GRU model.
+The Buzzometer sentiment analysis tool uses a bi-directional GRU model, in which the output a 4-layer bidirectional recurrent network is fed into a two-layer fully connected network which separates the input into four classes, corresponding to neutral, positive, negative and irrelevant messages.
+The only non-linearities in the network are inside the GRU cells.
+Graphically, the network can be represented as:
+
+%(GRAPHICAL REPRESENTATION OF NETWORK)
+\begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
+    \tikzstyle{arrow} = [thick,->,>=stealth]
+    \tikzstyle{startstop} = [rectangle, minimum width=1.5cm, minimum height=.6cm, text centered, draw=black, fill=black!20]
+    \node (input) [] at (0cm,1cm) {};
+    \node (output) [] at (0cm,-3cm) {};
+    \node (s-input-A) [] at (-2cm,0cm) {};
+    \node (s-output-A) [] at (10cm,0cm) {};
+    \node (s-input-B) [] at (-2cm,-1cm) {};
+    \node (s-output-B) [] at (10cm,-1cm) {};
+    \node (s-input-C) [] at (-2cm,-2cm) {};
+    \node (s-output-C) [] at (10cm,-2cm) {};
+    \path[] node[startstop] (A-0) at (0cm,0cm) {};
+    \path[] node[startstop] (B-0) at (0cm,-1cm) {};
+    \path[] node[startstop] (C-0) at (0cm,-2cm) {};
+    \draw [arrow] (input) -- (A);
+    \draw [arrow] (s-input-A) -- (A-0);
+    \draw [arrow] (s-input-B) -- (B-0);
+    \draw [arrow] (s-input-C) -- (C-0);
+    \draw [arrow] (A) -- (B);
+    \draw [arrow] (B) -- (C);
+    \draw [arrow] (C) -- (output);
+
+     \foreach \i/\j in {1/0,2/1,3/2,4/3}
+     {
+       \node (input-\i) [] at (2*\i cm,1cm) {};
+       \node (output-\i) [] at (2*\i cm,-3cm) {};
+       \path[] node[startstop] (A-\i) at (2*\i cm,0cm) {};
+       \path[] node[startstop] (B-\i) at (2*\i cm,-1cm) {};
+       \path[] node[startstop] (C-\i) at (2*\i cm,-2cm) {};
+       \draw [arrow] (A-\i) -- (B-\i);
+       \draw [arrow] (B-\i) -- (C-\i);
+       \draw [arrow] (input-\i) -- (A-\i);
+       \draw [arrow] (A-\j) -- (A-\i);
+       \draw [arrow] (B-\j) -- (B-\i);
+       \draw [arrow] (C-\j) -- (C-\i);
+       \draw [arrow] (C-\i) -- (output-\i);
+     }
+     \draw [arrow] (A-4) -- (s-output-A);
+     \draw [arrow] (B-4) -- (s-output-B);
+     \draw [arrow] (C-4) -- (s-output-C);
+
+\end{tikzpicture}
+
+\begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
+    \tikzstyle{every pin edge}=[<-,shorten <=1pt]
+    \tikzstyle{neuron}=[circle,fill=black!50,minimum size=10pt,inner sep=0pt]
+
+    % Draw the hidden layer nodes
+    \foreach \name / \x in {1,...,5}
+        \path[yshift=0.5cm]
+            node[neuron] (I-\name) at (\x,-0cm) {};
+
+    \foreach \name / \x in {1,...,5}
+        \path[yshift=0.5cm]
+            node[neuron] (H-\name) at (\x,-2cm) {};
+
+    \foreach \source in {1,...,5}
+        \foreach \dest in {1,...,5}
+            \path (I-\source) edge (H-\dest);
+
+    \foreach \name / \x in {1,...,5}
+        \path[yshift=0.5cm]
+            node[neuron] (K-\name) at (\x,-4 cm) {};
+
+    \foreach \source in {1,...,5}
+        \foreach \dest in {1,...,5}
+            \path (H-\source) edge (K-\dest);
+\end{tikzpicture}
+
+The base input ofor this network is a (unicode) character, which is one-hot encoded before being fed into the recurrent layers.
+For training, the network is unrolled to a length of 256 characters, which is about the length of an average
+message in our database.  All strings are padded or truncated to a length of 256 characters.
+
+\begin{minted}{python}
+  SEQ_LENGTH = 256
+  E_DIM = 128
+  STATE_DIM = 512
+  NUM_CLASSES = 4
+  def inference():
+      model_input = tf.placeholder('uint8', shape=[None, SEQ_LENGTH])
+      _ = tf.one_hot(Globals.model_input, depth=E_DIM, axis=-1)
+      _ = tf.reshape(_, [-1, SEQ_LENGTH, E_DIM])
+      fw = multi_layer_rnn(N_LAYERS, STATE_DIM)
+      bw = multi_layer_rnn(N_LAYERS, STATE_DIM)
+      output, _ = tf.nn.bidirectional_dynamic_rnn(fw, bw, _, dtype=tf.float32)
+      fw_output = tf.reshape(output[0][:, -1:], [-1, STATE_DIM])
+      bw_output = tf.reshape(output[1][:, :1], [-1, STATE_DIM])
+      f = project(fw_output, E_DIM)
+      b = project(bw_output, E_DIM)
+      e = tf.add(f, b)
+      Globals.model_output = project(e, NUM_CLASSES)
+      Globals.prediction = tf.cast(tf.argmax(Globals.model_output, 1), tf.uint8)
+      return Globals.model_input, Globals.model_output
+\end{minted}
+
+first hour:
+Loss: 1.331 Accuracy: 45.50%
+0	Negative	Neutral	Positive	Irrelevant
+Negative	291	113	50	127
+Neutral	108	164	62	85
+Positive	0	0	0	0
+Irrelevant	0	0	0	0
+
+Loss: 0.933 Accuracy: 60.30%
+0	Negative	Neutral	Positive	Irrelevant
+Negative	292	52	36	55
+Neutral	61	176	21	62
+Positive	5	15	65	15
+Irrelevant	33	28	14	70
+
+second hour
+Loss: 0.729 Accuracy: 71.80%
+0	Negative	Neutral	Positive	Irrelevant
+Negative	340	67	18	65
+Neutral	22	182	18	36
+Positive	10	7	91	10
+Irrelevant	8	13	8	105
+
+third hour
+Loss: 0.611 Accuracy: 76.70%
+0	Negative	Neutral	Positive	Irrelevant
+Negative	319	38	11	43
+Neutral	20	188	2	20
+Positive	10	18	91	10
+Irrelevant	27	27	7	169
+
+Loss: 0.468 Accuracy: 85.20%
+0	Negative	Neutral	Positive	Irrelevant
+Negative	334	20	3	18
+Neutral	14	242	0	9
+Positive	20	10	125	17
+Irrelevant	26	10	1	151
+
+0	Negative	Neutral	Positive	Irrelevant
+Negative	334	4	4	9
+Neutral	3	283	2	7
+Positive	1	0	125	2
+Irrelevant	1	0	4	221
+
+Loss: 0.530 Accuracy: 84.60%
+0	Negative	Neutral	Positive	Irrelevant
+Negative	361	18	13	29
+Neutral	22	246	10	18
+Positive	6	6	95	6
+Irrelevant	11	5	10	144
+
+Loss: 0.275 Accuracy: 90.00%
+0	Negative	Neutral	Positive	Irrelevant
+Negative	321	16	3	20
+Neutral	8	286	4	6
+Positive	4	2	125	16
+Irrelevant	11	9	1	168
 
@@
 
-Text generator
+\begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
+    \tikzstyle{arrow} = [thick,->,>=stealth]
+    \tikzstyle{startstop} = [rectangle, minimum width=1.5cm, minimum height=.6cm, text centered, draw=black, fill=black!20]
+    \node (input) [] at (0cm,1cm) {};
+    \node (output) [] at (0cm,-3cm) {};
+    \node (s-input-A) [] at (-2cm,0cm) {};
+    \node (s-output-A) [] at (10cm,0cm) {};
+    \node (s-input-B) [] at (-2cm,-1cm) {};
+    \node (s-output-B) [] at (10cm,-1cm) {};
+    \node (s-input-C) [] at (-2cm,-2cm) {};
+    \node (s-output-C) [] at (10cm,-2cm) {};
+    \path[] node[startstop] (A-0) at (0cm,0cm) {};
+    \path[] node[startstop] (B-0) at (0cm,-1cm) {};
+    \path[] node[startstop] (C-0) at (0cm,-2cm) {};
+    \draw [arrow] (input) -- (A);
+    \draw [arrow] (s-input-A) -- (A-0);
+    \draw [arrow] (s-input-B) -- (B-0);
+    \draw [arrow] (s-input-C) -- (C-0);
+    \draw [arrow] (A) -- (B);
+    \draw [arrow] (B) -- (C);
+    \draw [arrow] (C) -- (output);
+
+     \foreach \i/\j in {1/0,2/1,3/2,4/3}
+     {
+       \node (input-\i) [] at (2*\i cm,1cm) {};
+       \node (output-\i) [] at (2*\i cm,-3cm) {};
+       \path[] node[startstop] (A-\i) at (2*\i cm,0cm) {};
+       \path[] node[startstop] (B-\i) at (2*\i cm,-1cm) {};
+       \path[] node[startstop] (C-\i) at (2*\i cm,-2cm) {};
+       \draw [arrow] (A-\i) -- (B-\i);
+       \draw [arrow] (B-\i) -- (C-\i);
+       \draw [arrow] (input-\i) -- (A-\i);
+       \draw [arrow] (A-\j) -- (A-\i);
+       \draw [arrow] (B-\j) -- (B-\i);
+       \draw [arrow] (C-\j) -- (C-\i);
+       \draw [arrow] (C-\i) -- (output-\i);
+     }
+     \draw [arrow] (A-4) -- (s-output-A);
+     \draw [arrow] (B-4) -- (s-output-B);
+     \draw [arrow] (C-4) -- (s-output-C);
+
+\end{tikzpicture}
+
+\begin{minted}{python}
+  SEQ_LENGTH = 256
+  E_DIM = 128
+  STATE_DIM = 512
+  N_LAYERS = 3
+
+  def inference():
+      model_input = tf.placeholder('uint8', shape=[None, SEQ_LENGTH])
+      _ = tf.one_hot(Globals.model_input, depth=E_DIM, axis=-1)
+      encode = multi_layer_rnn(N_LAYERS, STATE_DIM)
+      state_tuple = tuple(tf.unstack(Globals.initial_state, axis=0))
+      output, state = tf.nn.dynamic_rnn(encode, _,
+                                        dtype=tf.float32,
+                                        initial_state=state_tuple)
+      output = tf.reshape(output, [-1, STATE_DIM])
+      output = project(output, E_DIM)
+      out = tf.cast(tf.argmax(output, 1), tf.uint8)
+      out = tf.reshape(out, [-1, SEQ_LENGTH])
+      Globals.generated_sequence = out
+      Globals.generated_characters = tf.nn.softmax(output)
+      Globals.model_output = output
+      Globals.state = state
+\end{minted}
+
+\begin{minted}{python}
+def generate_text(length, session=None):
+    generated_text = ''
+    character = [[ord(' ')]]
+    istate = np.zeros([N_LAYERS, 1, STATE_DIM])
+    while len(generated_text) < length:
+        feed_dict = {Globals.model_input: character,
+                     Globals.initial_state: istate}
+        next_char, state = session.run([Globals.generated_characters,
+                                        Globals.state],
+                                       feed_dict=feed_dict)
+        next_char = np.asarray(next_char).astype('float64')
+        next_char = next_char / next_char.sum()
+        op = np.random.multinomial
+        next_char_id = op(1, next_char.squeeze(), 1).argmax()
+        next_char_id = next_char_id if chr(next_char_id) in \
+                          string.printable else ord(" ")
+        generated_text += chr(next_char_id)
+        character = [[next_char_id]]
+        istate = state
+    return generated_text
+\end{minted}
 
@@
 
-% \section{The basic problem}
-%
-% Let $U=\R^n$ and $V=\R^m$ be vector spaces, and consider a function $f:U\to W$.
-% In theory, $f$ here can be any set theoretic function, but for our purpose, we
-% will assume that $f$ is at the very least square integrable. In general, we
-% consider a class $\C$ of functions $U\to V$, and try to find an element $g\in\C$
-% which is as close as possible to $f$.  In order to do so, we need a notion of distance
-% between functions.  There are many possible choices, but some of the most popular are
-% the $L^p$ norms and the $L^\infty$ norm.
-%
-% While the class $\C$ van be virtually any class of functions, we usually choose a class
-% that is somewhat definable, in the sense that there is a (differentiable) function $F:U\times\R^k\to V$
-% such that $\C=\{F(x, w): w\in \R^k\}$.  Such classes of functions have the advantage of
-% being equicontinuous.
-%
-%
-% %\begin{example}
-% Let $f(x)=x(x-1)(x+1)$.  As an example, we will try to fit a linear function
-% $\ell(x)=ux+v$ to $f$.  The class $\C$ is defined by $\C=\{ux+v:u,v\in R\}$.
-% For an element $g(x, u, v)\in\C$, we have $$f(x)-g(x,u,v) = x^3-x - ux+v$$
-% From this we get
-% \begin{eqnarray*}
-%   (f(x)-g(x,u,v))^2 &=& (x^3 - (1-u)x + v)^2 \\
-%   {} &=& x^6 - 2(1-u)x^4 + 2vx^3 + (1-u)^2x^2 - \\
-%   {} & & 2(1-u)vx + v^2
-% \end{eqnarray*}
-% If we compute $\int_{-1}^1 (f(x)-g(x,u,v))^2$, we get
-%
-% \begin{eqnarray*}
-%   \int_{-1}^1(f(x)-g(x; u,v))^2 &=& \int_{-1}^1(x^3 - (1-u)x + v)^2 \\
-%   {} &=& \int_{-1}^1x^6 - 2(1-u)x^4 + 2vx^3 + (1-u)^2x^2 - 2(1-u)vx + v^2\\
-%   {} &=& \frac{x^7}{7}-2v\frac{x^4}{4} + (1-u)^2\frac{x^3}{3} - 2(1-u)v\frac{x^2}{2}+v^2x\\
-%   {} &=& \frac{1}{7}-2v\frac{1}{4} + (1-u)^2\frac{1}{3} - 2(1-u)v\frac{1}{2}+v^2\\
-%   {} & & +\frac{1}{7}-2v\frac{1}{4} + (1-u)^2\frac{-1}{3} - 2(1-u)v\frac{1}{2}-v^2\\
-%   {} &=& \frac{2}{7} - (2-u)v+v^2 = \ell(u,v)
-% \end{eqnarray*}
-% and note that the last line above is a function of $u$ and $v$ alone, and we have to minimize it.
-% %\end{example}
-%
-%
-% \section{Dealing with sequences}
+\begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
+    \tikzstyle{arrow} = [thick,->,>=stealth]
+    \tikzstyle{startstop} = [rectangle, minimum width=1.5cm, minimum height=.6cm, text centered, draw=black, fill=black!20]
+    \node (input) [] at (0cm,1cm) {};
+    \node (output) [] at (0cm,-3cm) {};
+    \node (s-input-A) [] at (-2cm,0cm) {};
+    \node (s-output-A) [] at (10cm,0cm) {};
+    \node (s-input-B) [] at (-2cm,-1cm) {};
+    \node (s-output-B) [] at (10cm,-1cm) {};
+    \node (s-input-C) [] at (-2cm,-2cm) {};
+    \node (s-output-C) [] at (10cm,-2cm) {};
+    \path[] node[startstop] (A-0) at (0cm,0cm) {};
+    \path[] node[startstop] (B-0) at (0cm,-1cm) {};
+    \path[] node[startstop] (C-0) at (0cm,-2cm) {};
+    \draw [arrow] (input) -- (A);
+    \draw [arrow] (s-input-A) -- (A-0);
+    \draw [arrow] (s-input-B) -- (B-0);
+    \draw [arrow] (s-input-C) -- (C-0);
+    \draw [arrow] (A) -- (B);
+    \draw [arrow] (B) -- (C);
+    \draw [arrow] (C) -- (output);
+
+     \foreach \i/\j in {1/0,2/1,3/2,4/3}
+     {
+       \node (input-\i) [] at (2*\i cm,1cm) {};
+       \node (output-\i) [] at (2*\i cm,-3cm) {};
+       \path[] node[startstop] (A-\i) at (2*\i cm,0cm) {};
+       \path[] node[startstop] (B-\i) at (2*\i cm,-1cm) {};
+       \path[] node[startstop] (C-\i) at (2*\i cm,-2cm) {};
+       \draw [arrow] (A-\i) -- (B-\i);
+       \draw [arrow] (B-\i) -- (C-\i);
+       \draw [arrow] (input-\i) -- (A-\i);
+       \draw [arrow] (A-\j) -- (A-\i);
+       \draw [arrow] (B-\j) -- (B-\i);
+       \draw [arrow] (C-\j) -- (C-\i);
+       \draw [arrow] (C-\i) -- (output-\i);
+     }
+     \draw [arrow] (A-4) -- (s-output-A);
+     \draw [arrow] (B-4) -- (s-output-B);
+     \draw [arrow] (C-4) -- (s-output-C);
+
+\end{tikzpicture}
+
+\begin{minted}{python}
+  SEQ_LENGTH = 256
+  E_DIM = 128
+  STATE_DIM = 512
+  N_LAYERS = 4
+
+  def inference():
+      model_input = tf.placeholder('uint8', shape=[None, SEQ_LENGTH])
+      _ = tf.one_hot(Globals.model_input, depth=E_DIM, axis=-1)
+      _ = tf.reshape(_, [-1, SEQ_LENGTH, E_DIM])
+      encode = multi_layer_rnn(N_LAYERS, STATE_DIM)
+      encoded_input, state = tf.nn.dynamic_rnn(encode,
+                                               _,
+                                               dtype=tf.float32)
+      Globals.encoder_output = state
+      with tf.variable_scope('decoder'):
+          training_decoder_input = tf.zeros_like(Globals.model_input)
+          _ = tf.one_hot(training_decoder_input, depth=E_DIM, axis=-1)
+          _ = tf.reshape(_, [-1, SEQ_LENGTH, E_DIM])
+          decode = multi_layer_rnn(N_LAYERS, STATE_DIM)
+          decoded_output, state = tf.nn.dynamic_rnn(decode, _,
+                                                    dtype=tf.float32,
+                                                    initial_state=state)
+          decoded_output = tf.reshape(decoded_output, [-1, STATE_DIM])
+          output = project(decoded_output, E_DIM)
+          out = tf.cast(tf.argmax(output, 1), tf.uint8)
+          out = tf.reshape(out, [-1, SEQ_LENGTH])
+          Globals.training_decoder_input = training_decoder_input
+          Globals.model_output = output
+          Globals.prediction = out
+          Globals.decoder = decode
+          Globals.decoder_input = _
+\end{minted}
 

### 2017-08-21 deeplearn/presentation/analyse_gru_sort_results.py

@@ -1,0 +1,31 @@
+import os
+
+
+if __name__ == '__main__':
+    path = os.path.join(os.path.dirname(__file__), 'gru_sort_outputs')
+
+    files = {}
+    for file_name in os.listdir(path):
+        _, _, i = file_name.split('-')
+        i = int(i)
+        files[i] = os.path.join(path, file_name)
+    files_in_order = sorted(files.keys())
+    accuracies = []
+    for index in files_in_order:
+        file_name = files[index]
+        print(file_name, end="")
+        with open(file_name) as data_file:
+            lines = data_file.readlines()
+            accs = 0
+            for line in lines:
+                input_, output_ = line.split(';')
+                input_ = [int(x) for x in input_.split(',')]
+                output_ = [int(x) for x in output_.split(',')]
+                N = len(input_)
+                I = 0
+                for x, y in zip(sorted(input_), output_):
+                    if x == y:
+                        I += 1
+                accs += float(I) / N
+            accuracies.append(accs / len(lines))
+            print("  ", accs / len(lines))

### 2017-08-21 deeplearn/twitter_sentiment/models/text_generator/text_generator.py
presentation

@@ -187,6 +187,7 @@ def rnn_minibatch_sequencer(raw_data, batch_size, sequence_size, nb_epochs):
 def generate_text(length, session=None):
     generated_text = ''
     character = [[ord(' ')]]
+    print('Generating text')
     istate = np.zeros([Hyperparameters.n_layers, 1, Hyperparameters.hidden_states])
     while len(generated_text) < length:
         feed_dict = {Globals.model_input: character, Globals.initial_state: istate}
@@ -208,18 +209,21 @@ if __name__ == '__main__':
     chkpt = os.path.join(os.path.expanduser('~'), '.text_gen', 'model_chkpt')
     if not os.path.exists(save):
         os.makedirs(save)
-    if not os.path.exists(chkpt):
-        os.makedirs(chkpt)
-
     data = read_data_files_from_folder(os.path.join(root, 'harry_potter/*.txt'))
-
     inference()
     train()
     index = 0
-    istate = np.zeros([Hyperparameters.n_layers, 128, Hyperparameters.hidden_states])
+    istate = np.zeros([Hyperparameters.n_layers, 32, Hyperparameters.hidden_states])
     with tf.Session() as session:
         session.run(tf.global_variables_initializer())
-        for batch in rnn_minibatch_sequencer(data, 128, Hyperparameters.sequence_length, nb_epochs=100):
+        if not os.path.exists(chkpt):
+            os.makedirs(chkpt)
+        else:
+            s = tf.train.Saver()
+            s.save(session, os.path.join(chkpt, "t_gen_weights"))
+            print('Model restored')
+
+        for batch in rnn_minibatch_sequencer(data, 32, Hyperparameters.sequence_length, nb_epochs=100):
             x = batch['train_x']
             y = batch['train_y']
             feed_dict = {Globals.model_input: x,
@@ -228,8 +232,9 @@ if __name__ == '__main__':
             _, ostate = session.run([Globals.minimize, Globals.state], feed_dict=feed_dict)
             print('x', end='', flush=True)
             istate = ostate
-            if index % batch['epoch_number']*150 == 0:
-                file_name = "sort-test-{}".format(index)
+            #print(index, batch['epoch_number'])
+            if index % (batch['epoch_number']*150) == 0:
+                file_name = "text_gen-{}".format(index)
                 file_name = os.path.join(save, file_name)
                 with open(file_name, 'w') as test_file:
                     test_file.write(generate_text(20000, session))

### 2017-08-22 deeplearn/presentation/paper_2.tex
presentation

@@
 \usepackage{tikz}
+\usepackage{booktabs}
 \usetikzlibrary{shapes.geometric, arrows}
@@
 \newcommand{\C}{\mathcal{C}}
+\renewcommand{\u}{\mathbf{u}}
+\renewcommand{\v}{\mathbf{v}}
+\newcommand{\w}{\mathbf{w}}
+\newcommand{\x}{\mathbf{x}}
+\newcommand{\y}{\mathbf{y}}
+\renewcommand{\b}{\mathbf{b}}
+\newcommand{\e}{\mathbf{e}}
+
+\newcommand{\A}{\mathbf{A}}
+\newcommand{\B}{\mathbf{B}}
+\newcommand{\W}{\mathbf{W}}
 
@@
 
-%\def\layersep{2.5cm}
-
-% \newcommand{\nodedef}{\tikzstyle{arrow} = [thick,->,>=stealth]
-%     \node (input) [] at (0cm,1cm) {};
-%     \node (output) [] at (0cm,-1cm) {};
-%     \node (s-input) [] at (-2cm,0cm) {};
-%     \node (s-output) [] at (2cm,0cm) {};
-%     \path[yshift=0.5cm]
-%         node[startstop] (N) at (0cm,-.5cm) {};
-%
-%     \draw [arrow] (input) -- (N);
-%     \draw [arrow] (s-input) -- (N);
-%     \draw [arrow] (N) -- (output);
-%     \draw [arrow] (N) -- (s-output);}
-%
-%
-% \newcommand{\nodedefinner}{\tikzstyle{arrow} = [thick,->,>=stealth]
-%     \node (input) [] at (0cm,1cm) {};
-%     \node (output) [] at (0cm,-1cm) {};
-%     \node (s-input) [] at (-2cm,0cm) {};
-%     \node (s-output) [] at (2cm,0cm) {};
-%     \path[yshift=0.5cm]
-%         node[startstop] (N) at (0cm,-.5cm) {};
-%
-%     %\draw [arrow] (input) -- (N);
-%     %\draw [arrow] (s-input) -- (N);
-%     \draw [arrow] (N) -- (output);
-%     \draw [arrow] (N) -- (s-output);}
-
-% \begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
-%     \tikzstyle{arrow} = [thick,->,>=stealth]
-%     \tikzstyle{startstop} = [rectangle, minimum width=2cm, minimum height=.75cm, text centered, draw=black, fill=black!20]
-%     \node (input) [] at (0cm,1cm) {};
-%     \node (output) [] at (0cm,-1cm) {};
-%     \node (s-input) [] at (-2cm,0cm) {};
-%     \node (s-output) [] at (2cm,0cm) {};
-%     \path[] node[startstop] (N) at (0cm,0cm) {};
-%     \draw [arrow] (input) -- (N);
-%     \draw [arrow] (s-input) -- (N);
-%     \draw [arrow] (N) -- (output);
-%     \draw [arrow] (N) -- (s-output);
-% \end{tikzpicture}
-%
-%
-
  \section{Introduction}
 
- Let $U=\R^n$ and $V=\R^m$ be vector spaces, and consider a function $f:U\to W$.
- In theory, $f$ here can be any set theoretic function, but for our purpose, we
- will assume that $f$ is at the very least square integrable. In general, we
- consider a class $\C$ of functions $U\to V$, and try to find an element $g\in\C$
- which is as close as possible to $f$.  In order to do so, we need a notion of distance
- between functions.  There are many possible choices, but some of the most popular are
- the $L^p$ norms and the $L^\infty$ norm.
+Throughout this handout, we use bold face lower case letters to represent vectors, and bold face capital letters to represent matrices.
+Most of the time, we will blur the dustinction between matrices and vectors, except when it makes the exposition clearer.
 
- While the class $\C$ van be virtually any class of functions, we usually choose a class
- that is somewhat definable, in the sense that there is a (differentiable) function $F:U\times\R^k\to V$
- such that $\C=\{F(x, w): w\in \R^k\}$.  Such classes of functions have the advantage of
- being equicontinuous.
+ Let $f:\R^n\to \R^m$ be a function, and consider a class $\C$ of functions $g: \R^n\to \R^m$, which we assume to be parametrized by $\R^\ell$, in the sense that there is a function $F(\x,\w):\R^n\times \R^\ell\to\R^n$ such that $\C=\{\x\mapsto F(\x,\w):\w\in\R^n\}$.
+The generic problem of machine learning is to find a vector $\w\in\R^{\ell}$ such that $D(f(\x), F(\x, \w))$ is as small as possible, where $D$ represents a notion of distance between functions, for example, $$D(f(\x), F(\x, \w)) = \min\{\|f(\x)-F(\x, \w)\|:\x\in\R^n\}$$ Note that since the minimum is computed over all possible vlues of $\x$, the value of $D$ in te equation above depends on $\w$.
+Define $$\ell(\w)=D(f(\x), F(\x, \w))$$ Given the definition of $D$, we see that $\ell(\x)\geq 0$, and therefore $\ell$ has a global minimum (which is not necessarily unique).
+In general, the global minimum of $\ell$ is not attained, in the sense that if $\epsilon=\min\{\ell(\w):\x\in\R^n\}$, then there does not necessarily exist a value $\w_0\in\R^\ell$ such that $\epsilon=\ell(\w_0)$.
+However, it can be approximated, that is to say there is a sequence $\w_i$ such that $\ell(\w_i)\to\epsilon$ as $i\to\infty$.
+We may not be able to achieve the minimum with a specific value of $\w$, but we can get as close to it as we want: for every $\varepsilon>0$, there is some natural number $N_\varepsilon>0$ such that $\|\epsilon-\ell(\w_i)\|<\varepsilon$ for every $i>N_\varepsilon$.
+Since we can choose $\varepsilon$ to be as small as we want, if we choose it to be smaller than the smallest number which can be represented with a {\tt float32} (say), then from the point of view of any program, $\ell(\w_i)=\epsilon$ when $i>N_\varepsilon$.
 
- \begin{example}
- Let $f(x)=x(x-1)(x+1)$.  As an example, we will try to fit a linear function
- $\ell(x)=ux+v$ to $f$.  The class $\C$ is defined by $\C=\{ux+v:u,v\in R\}$.
- For an element $g(x, u, v)\in\C$, we have $$f(x)-g(x,u,v) = x^3-x - ux+v$$
- From this we get
- \begin{eqnarray*}
-   (f(x)-g(x,u,v))^2 &=& (x^3 - (1-u)x + v)^2 \\
-   {} &=& x^6 - 2(1-u)x^4 + 2vx^3 + (1-u)^2x^2 - \\
-   {} & & 2(1-u)vx + v^2
- \end{eqnarray*}
- If we compute $\int_{-1}^1 (f(x)-g(x,u,v))^2$, we get
+ The function $\ell$ we defined above is called a {\em loss function}.
+There are many functions which are suitable for $\ell$.
+In fact, $\ell$ does not need to represent a distance at all.
+In order to draw a meaningful conclusion from the minimum value of $\ell$, all we need to know is that $\ell(\w)=0$ implies $f(\x)=F(\x,\w)$ for every $\x$, or at the very least $\|f(\x)-F(\x,\w)\|\leq\varepsilon$ for arbitrary $\varepsilon$.
+In every neural network we define, we specify a class of functions $F(\x, \w)$, and a definition for $\ell(\w)$ based on $F$.
+In tensorflow, the function we are trying to approximate is given in the form of a relation $f(\x)=\y$.
+We provide {\em placeholders} for the values of $\x$ and $\y$.
+The function $F(\x,\w)$ is defined by a computation graph.
+For example, here is a linear regression, in which we try to approximate a function $f:\R^{10}\to\R^{15}$ using a function of the form $F(\x, \W\b)=\W\x+\b$.
+Here $\W$ is an $15\times 10$ matrix, and $\b\in\R^{15}$
 
- \begin{eqnarray*}
-   \int_{-1}^1(f(x)-g(x; u,v))^2 &=& \int_{-1}^1(x^3 - (1-u)x + v)^2 \\
-   {} &=& \int_{-1}^1x^6 - 2(1-u)x^4 + 2vx^3 + (1-u)^2x^2 - 2(1-u)vx + v^2\\
-   {} &=& \frac{x^7}{7}-2v\frac{x^4}{4} + (1-u)^2\frac{x^3}{3} - 2(1-u)v\frac{x^2}{2}+v^2x\\
-   {} &=& \frac{1}{7}-2v\frac{1}{4} + (1-u)^2\frac{1}{3} - 2(1-u)v\frac{1}{2}+v^2\\
-   {} & & +\frac{1}{7}-2v\frac{1}{4} + (1-u)^2\frac{-1}{3} - 2(1-u)v\frac{1}{2}-v^2\\
-   {} &=& \frac{2}{7} - (2-u)v+v^2 = \ell(u,v)
- \end{eqnarray*}
- and note that the last line above is a function of $u$ and $v$ alone, and we have to minimize it.
- \end{example}
+ \begin{minted}{python}
+ x = tf.placeholder(shape=[None, 10])
+ y = tf.placeholder(shape=[None, 15])
+ W = tf.Variable(shape=[None, 10, 15])
+ b = tf.Variable(shape=[None, 15])
+ F = tf.matmul(x, W) + b
+ loss = tf.mean_square_error(y, F)
+ \end{minted}
 
-\section{Back Propagation}
+We are mainly interested in classification problems, or more generally problems which can be expressed as classification problems.
+Let $A$ and $B$ be finite sets, and let $f:A\to B$ be any function.
+In order to use neural networks, which represent functions between topological vector spaces, to represent  function between ordinary sets, we must first translate the sets $A$ and $B$ into vector spaces.
+The most straightforward (and freest) way to interpret a set $A$ as a vector space is to use $A$ as the basis for a vector space.
+Write $A=\{a_1,...,a_n\}$, and let $\{\e_1,...,\e_n\}$ represent the standard basis for $\R^n$.
+Let $e:A\to\R^n$ be defined by $e(a_i)=\e_i$.
+Similarly, we can write $B=\{b_1,...,b_m\}$, and define $e:B\to \R^m$ by $e(b_i)=\e_i$.
+This specific function is called {\em one-hot encoding}, and is present in tensorflow as tf.onehot.
+Since we are mapping a finite set to an infinte set, the function $e$ has no inverse, but it has a "pseudo-inverse" given by the {\em argmax()} function defined by $argmax(\v)=i$ where $i$ is smallest such that $v_i\geq v_j$ for every $j\not=i$.
+In other words, $argmax$ computes the index of the largest component of a vector.
 
-In general, in order to minimize a function $f(w)$ of one variable, we make use of the fact that
-the extreme values of $f$ all have $f'(x)=0$.  If $f$ is a function of two or more variables, then
-if $x$ is such that $f(x)$ is a local max, or minimum, then all partial derivatives
-$\frac{\partial f}{\partial w_i}(x)=0$.  If $f:\R^n\to\R$ is a function, then the
-{\em gradient} of $f$ is the vector
-$\nabla f(x) = \left(\frac{\partial f}{\partial w_1}(x),...,\frac{\partial f}{\partial w_n}(x)\right)$.
-One interesting property of $\nabla f(x)$ is that, as a vector, it ``points'' in the direction in which
-$f$ decreases the most.  That is to say, there is a good chance that $f(x-\alpha\nabla f(x))\leq f(x)$.
-The chance is better the smaller $\alpha$ is.  We see that if we iterate that process, we get a
-sequence $x_1,...,x_n,...$ such that $f(x_{i+1})\leq f(x_i)$ for every $i$.  Since $m\leq f(x_i)\leq f(x_0)$
-for every $i$, we get a bounded decreasing sequence of real number, which converges.
+Given $\v\in\R^m$ is the output of $F(\x,\w)$, then if $\v$ has the property that $\sum v_i=1$, and $v_i\geq 0$ for every $i$, then we can interpret $\v$ as a probability distribution, where each component $v_i$ represents the probability that $\v=\e_i$.
+The code, in tensorflow, which approximates a function $A\to B$ can be written as:
 
-Caveat: this ideal situation doesn't allways happen.
+\begin{minted}{python}
+x_in = tf.placeholder(shape=[None, 1])
+x = tf.one_hot(x_in, 10)
+y_in = tf.placeholder(shape=[None, 1])
+y = tf.one_hot(y_in, 15)
+W = tf.Variable(shape=[None, 10, 15])
+b = tf.Variable(shape=[None, 15])
+F = tf.softmax(tf.matmul(x, W) + b)
+loss = tf.cross_entropy(y, F)
+\end{minted}
 
-This iterative process is {\em (Stocastic) gradient descent}.
+The softmax function in the penultimate line is defined by $\sigma(\v) = \w$  where $$w_i= \frac{e^{v_i}}{\sum e^{v_j}}$$ Note that $\sum w_j=1$, which means that the output of the softmax function can be interpreted as a probability distribution.
+This makes it a popular choice for classification problems with many classes.
+If the target space is $\R^2$, then this reduces to a binomial distribution.
+The cross-entropy function in the last line is defined by $C(\v, \w) = -\sum(v_i\log(w_i))$.
+Intuitively, $C$ measures the number of vector components that are necessary to differentiate between the distribution $\v$ and the distribution $\w$.
+Therefore, the optimal value for $C$ to have is $1$: measuring any bit should allow us to differentiate between $\v$ and $\w$.
+Note that $C(\v, \w)=1$ if and only if $\log C(\v, \w)=0$, so that the logarithm of the cross entropy function is a candidate for a distance function.
+The vector $\w$ here should be the output of the model, whereas $\v$ should be the true value.
+By definition, $w_i= \frac{e^{v_i}}{\sum e^{v_j}}$, so that $\log(w_i) = \log e^{w_i} - \log(\sum e^{v_j}) = w_i - \log(\sum e^{v_j})$.
+By definition again, $v_i=1$ and $v_j=0$ if $i\not=j$.
+Therefore, $C(\v, \w) = w_i + \log(\sum e^{w_j})$
 
-Suppose $f$ is a complsition of two functions, and that we can write $f(x,w)= g(h(x, w_1), w_2)$.
-Then we can write $\nabla_w f(x, w)=\nabla g(h(x, w_1), w_2)\cdot \nabla h(x, w_1)$. Each level
-of composition corresponds to a layer of neural network.  The chain rule transforms function composition
-into a product. If a network becomes deep, then the derivative of the loss function becomes a long
-product.  When gradients become small in norm (like they do when we approach a min), then the gradient
-becomes very close to $0$, and the update rule for gradient descent stops changing the weights.
-
-This is the vanishing gradient problem for very deep networks, and makes convergence slow.
+ % \begin{example}
+ % Let $f(x)=x(x-1)(x+1)$.  As an example, we will try to fit a linear function
+ % $\ell(x)=ux+v$ to $f$.  The class $\C$ is defined by $\C=\{ux+v:u,v\in R\}$.
+ % For an element $g(x, u, v)\in\C$, we have $$f(x)-g(x,u,v) = x^3-x - ux+v$$
+ % From this we get
+ % \begin{eqnarray*}
+ %   (f(x)-g(x,u,v))^2 &=& (x^3 - (1-u)x + v)^2 \\
+ %   {} &=& x^6 - 2(1-u)x^4 + 2vx^3 + (1-u)^2x^2 - \\
+ %   {} & & 2(1-u)vx + v^2
+ % \end{eqnarray*}
+ % If we compute $\int_{-1}^1 (f(x)-g(x,u,v))^2$, we get
+ %
+ % \begin{eqnarray*}
+ %   \int_{-1}^1(f(x)-g(x; u,v))^2 &=& \int_{-1}^1(x^3 - (1-u)x + v)^2 \\
+ %   {} &=& \int_{-1}^1x^6 - 2(1-u)x^4 + 2vx^3 + (1-u)^2x^2 - 2(1-u)vx + v^2\\
+ %   {} &=& \frac{x^7}{7}-2v\frac{x^4}{4} + (1-u)^2\frac{x^3}{3} - 2(1-u)v\frac{x^2}{2}+v^2x\\
+ %   {} &=& \frac{1}{7}-2v\frac{1}{4} + (1-u)^2\frac{1}{3} - 2(1-u)v\frac{1}{2}+v^2\\
+ %   {} & & +\frac{1}{7}-2v\frac{1}{4} + (1-u)^2\frac{-1}{3} - 2(1-u)v\frac{1}{2}-v^2\\
+ %   {} &=& \frac{2}{7} - (2-u)v+v^2 = \ell(u,v)
+ % \end{eqnarray*}
+ % and note that the last line above is a function of $u$ and $v$ alone, and we have to minimize it.
+ % \end{example}
 
@@
 \end{tikzpicture}
+
+\section{Back Propagation}
+
+In general, in order to minimize a function $f(w)$ of one variable, we make use of the fact that the extreme values of $f$ all have $f'(x)=0$.
+If $f$ is a function of two or more variables, then if $x$ is such that $f(x)$ is a local max, or minimum, then all partial derivatives $\frac{\partial f}{\partial w_i}(x)=0$.
+If $f:\R^n\to\R$ is a function, then the {\em gradient} of $f$ is the vector $\nabla f(x) = \left(\frac{\partial f}{\partial w_1}(x),...,\frac{\partial f}{\partial w_n}(x)\right)$.
+One interesting property of $\nabla f(x)$ is that, as a vector, it ``points'' in the direction in which $f$ decreases the most.
+That is to say, there is a good chance that $f(x-\alpha\nabla f(x))\leq f(x)$.
+The chance is better the smaller $\alpha$ is.
+We see that if we iterate that process, we get a sequence $x_1,...,x_n,...$ such that $f(x_{i+1})\leq f(x_i)$ for every $i$.
+Since $m\leq f(x_i)\leq f(x_0)$ for every $i$, we get a bounded decreasing sequence of real number, which converges.
+
+Caveat: this ideal situation doesn't allways happen.
+
+This iterative process is {\em (Stocastic) gradient descent}.
+
+Suppose $f$ is a complsition of two functions, and that we can write $f(x,w)= g(h(x, w_1), w_2)$.
+Then we can write $\nabla_w f(x, w)=\nabla g(h(x, w_1), w_2)\cdot \nabla h(x, w_1)$. Each level
+of composition corresponds to a layer of neural network.  The chain rule transforms function composition
+into a product. If a network becomes deep, then the derivative of the loss function becomes a long
+product.  When gradients become small in norm (like they do when we approach a min), then the gradient
+becomes very close to $0$, and the update rule for gradient descent stops changing the weights.
+
+This is the vanishing gradient problem for very deep networks, and makes convergence slow.
 
@@
 
-The Buzzometer sentiment analysis tool uses a bi-directional GRU model, in which the output a 4-layer bidirectional recurrent network is fed into a two-layer fully connected network which separates the input into four classes, corresponding to neutral, positive, negative and irrelevant messages.
-The only non-linearities in the network are inside the GRU cells.
-Graphically, the network can be represented as:
+The Buzzometer sentiment analysis tool uses a bi-directional GRU model, in which the output a 4-layer
+bidirectional recurrent network is fed into a two-layer fully connected network which separates the input into
+four classes, corresponding to neutral, positive, negative and irrelevant messages. The only non-linearities
+ in the network are inside the GRU cells.  Graphically, the network can be represented as:
 
@@
 
-The base input ofor this network is a (unicode) character, which is one-hot encoded before being fed into the recurrent layers.
-For training, the network is unrolled to a length of 256 characters, which is about the length of an average
-message in our database.  All strings are padded or truncated to a length of 256 characters.
+The base input ofor this network is a (unicode) character, which is one-hot encoded before being fed into the
+ recurrent layers.  For training, the network is unrolled to a length of 256 characters,
+ which is about twice the length  of an average message in our database, and all strings are padded or truncated
+ to a length of 256 characters.
 
@@
 
-first hour:
-Loss: 1.331 Accuracy: 45.50%
-0	Negative	Neutral	Positive	Irrelevant
-Negative	291	113	50	127
-Neutral	108	164	62	85
-Positive	0	0	0	0
-Irrelevant	0	0	0	0
+%first hour:
 
-Loss: 0.933 Accuracy: 60.30%
-0	Negative	Neutral	Positive	Irrelevant
-Negative	292	52	36	55
-Neutral	61	176	21	62
-Positive	5	15	65	15
-Irrelevant	33	28	14	70
+%Loss: 1.331 Accuracy: 45.50%
+%0	Negative	Neutral	Positive	Irrelevant
+%Negative	291	113	50	127
+%Neutral	108	164	62	85
+%Positive	0	0	0	0
+%Irrelevant	0	0	0	0
+\begin{tabular}{lcccc}
+  {} & {\bf Negative} & {\bf Neutral} & {\bf Positive} & {\bf Irrelevant}\\
+  {\bf Negative}   & {291} & {113} & {50} & {127}\\
+  {\bf Neutral}    & {108} & {113} & {50} & {85}\\
+  {\bf Positive}   & {0} & {0} & {0} & {0}\\
+  {\bf Irrelevant} & {0} & {0} & {0} & {0}\\
+\end{tabular}
 
-second hour
-Loss: 0.729 Accuracy: 71.80%
-0	Negative	Neutral	Positive	Irrelevant
-Negative	340	67	18	65
-Neutral	22	182	18	36
-Positive	10	7	91	10
-Irrelevant	8	13	8	105
+%Loss: 0.933 Accuracy: 60.30%
+%0	Negative	Neutral	Positive	Irrelevant
+%Negative	292	52	36	55
+%Neutral	61	176	21	62
+%Positive	5	15	65	15
+%Irrelevant	33	28	14	70
+\begin{tabular}{lcccc}
+  {} & {\bf Negative} & {\bf Neutral} & {\bf Positive} & {\bf Irrelevant}\\
+  {\bf Negative}   & {292} & {52} & {36} & {55}\\
+  {\bf Neutral}    & {61} & {176} & {21} & {62}\\
+  {\bf Positive}   & {5} & {15} & {65} & {15}\\
+  {\bf Irrelevant} & {33} & {28} & {14} & {70}\\
+\end{tabular}
 
-third hour
-Loss: 0.611 Accuracy: 76.70%
-0	Negative	Neutral	Positive	Irrelevant
-Negative	319	38	11	43
-Neutral	20	188	2	20
-Positive	10	18	91	10
-Irrelevant	27	27	7	169
+% second hour
+% Loss: 0.729 Accuracy: 71.80%
+% 0	Negative	Neutral	Positive	Irrelevant
+% Negative	340	67	18	65
+% Neutral	22	182	18	36
+% Positive	10	7	91	10
+% Irrelevant	8	13	8	105
+% \begin{tabular}{lcccc}
+%   {} & {\bf Negative} & {\bf Neutral} & {\bf Positive} & {\bf Irrelevant}\\
+%   {\bf Negative}   & {} & {} & {} & {}\\
+%   {\bf Neutral}    & {} & {} & {} & {}\\
+%   {\bf Positive}   & {} & {} & {} & {}\\
+%   {\bf Irrelevant} & {} & {} & {} & {}\\
+% \end{tabular}
 
-Loss: 0.468 Accuracy: 85.20%
-0	Negative	Neutral	Positive	Irrelevant
-Negative	334	20	3	18
-Neutral	14	242	0	9
-Positive	20	10	125	17
-Irrelevant	26	10	1	151
+%third hour
+%Loss: 0.611 Accuracy: 76.70%
+%0	Negative	Neutral	Positive	Irrelevant
+%Negative	319	38	11	43
+%Neutral	20	188	2	20
+%Positive	10	18	91	10
+%Irrelevant	27	27	7	169
+\begin{tabular}{lcccc}
+  {} & {\bf Negative} & {\bf Neutral} & {\bf Positive} & {\bf Irrelevant}\\
+  {\bf Negative}   & {319} & {38} & {11} & {43}\\
+  {\bf Neutral}    & {20} & {188} & {2} & {20}\\
+  {\bf Positive}   & {10} & {18} & {91} & {10}\\
+  {\bf Irrelevant} & {27} & {27} & {7} & {169}\\
+\end{tabular}
 
-0	Negative	Neutral	Positive	Irrelevant
-Negative	334	4	4	9
-Neutral	3	283	2	7
-Positive	1	0	125	2
-Irrelevant	1	0	4	221
+% Loss: 0.468 Accuracy: 85.20%
+% 0	Negative	Neutral	Positive	Irrelevant
+% Negative	334	20	3	18
+% Neutral	14	242	0	9
+% Positive	20	10	125	17
+% Irrelevant	26	10	1	151
+% \begin{tabular}{lcccc}
+%   {} & {\bf Negative} & {\bf Neutral} & {\bf Positive} & {\bf Irrelevant}\\
+%   {\bf Negative}   & {} & {} & {} & {}\\
+%   {\bf Neutral}    & {} & {} & {} & {}\\
+%   {\bf Positive}   & {} & {} & {} & {}\\
+%   {\bf Irrelevant} & {} & {} & {} & {}\\
+% \end{tabular}
+%
 
-Loss: 0.530 Accuracy: 84.60%
-0	Negative	Neutral	Positive	Irrelevant
-Negative	361	18	13	29
-Neutral	22	246	10	18
-Positive	6	6	95	6
-Irrelevant	11	5	10	144
+%0	Negative	Neutral	Positive	Irrelevant
+%Negative	334	4	4	9
+%Neutral	3	283	2	7
+%Positive	1	0	125	2
+%Irrelevant	1	0	4	221
+\begin{tabular}{lcccc}
+  {} & {\bf Negative} & {\bf Neutral} & {\bf Positive} & {\bf Irrelevant}\\
+  {\bf Negative}   & {334} & {4} & {4} & {9}\\
+  {\bf Neutral}    & {3} & {283} & {2} & {7}\\
+  {\bf Positive}   & {1} & {0} & {125} & {2}\\
+  {\bf Irrelevant} & {1} & {0} & {4} & {221}\\
+\end{tabular}
 
-Loss: 0.275 Accuracy: 90.00%
-0	Negative	Neutral	Positive	Irrelevant
-Negative	321	16	3	20
-Neutral	8	286	4	6
-Positive	4	2	125	16
-Irrelevant	11	9	1	168
+% Loss: 0.530 Accuracy: 84.60%
+% 0	Negative	Neutral	Positive	Irrelevant
+% Negative	361	18	13	29
+% Neutral	22	246	10	18
+% Positive	6	6	95	6
+% Irrelevant	11	5	10	144
+% \begin{tabular}{lcccc}
+%   {} & {\bf Negative} & {\bf Neutral} & {\bf Positive} & {\bf Irrelevant}\\
+%   {\bf Negative}   & {} & {} & {} & {}\\
+%   {\bf Neutral}    & {} & {} & {} & {}\\
+%   {\bf Positive}   & {} & {} & {} & {}\\
+%   {\bf Irrelevant} & {} & {} & {} & {}\\
+% \end{tabular}
+
+%Loss: 0.275 Accuracy: 90.00%
+%0	Negative	Neutral	Positive	Irrelevant
+%Negative	321	16	3	20
+%Neutral	8	286	4	6
+%Positive	4	2	125	16
+%Irrelevant	11	9	1	168
+\begin{tabular}{lcccc}
+  {} & {\bf Negative} & {\bf Neutral} & {\bf Positive} & {\bf Irrelevant}\\
+  {\bf Negative}   & {321} & {16} & {3} & {20}\\
+  {\bf Neutral}    & {8} & {286} & {4} & {6}\\
+  {\bf Positive}   & {4} & {2} & {125} & {16}\\
+  {\bf Irrelevant} & {11} & {9} & {1} & {168}\\
+\end{tabular}
 

### 2017-08-23 deeplearn/presentation/paper_2.tex
presentation update

@@
 Intuitively, $C$ measures the number of vector components that are necessary to differentiate between the distribution $\v$ and the distribution $\w$.
-Therefore, the optimal value for $C$ to have is $1$: measuring any bit should allow us to differentiate between $\v$ and $\w$.
+Therefore, the optimal value for $C$ to have is $N$: measuring any bit should allow us to differentiate between $\v$ and $\w$.
 Note that $C(\v, \w)=1$ if and only if $\log C(\v, \w)=0$, so that the logarithm of the cross entropy function is a candidate for a distance function.
@@
 
- % \begin{example}
- % Let $f(x)=x(x-1)(x+1)$.  As an example, we will try to fit a linear function
- % $\ell(x)=ux+v$ to $f$.  The class $\C$ is defined by $\C=\{ux+v:u,v\in R\}$.
- % For an element $g(x, u, v)\in\C$, we have $$f(x)-g(x,u,v) = x^3-x - ux+v$$
- % From this we get
- % \begin{eqnarray*}
- %   (f(x)-g(x,u,v))^2 &=& (x^3 - (1-u)x + v)^2 \\
- %   {} &=& x^6 - 2(1-u)x^4 + 2vx^3 + (1-u)^2x^2 - \\
- %   {} & & 2(1-u)vx + v^2
- % \end{eqnarray*}
- % If we compute $\int_{-1}^1 (f(x)-g(x,u,v))^2$, we get
- %
- % \begin{eqnarray*}
- %   \int_{-1}^1(f(x)-g(x; u,v))^2 &=& \int_{-1}^1(x^3 - (1-u)x + v)^2 \\
- %   {} &=& \int_{-1}^1x^6 - 2(1-u)x^4 + 2vx^3 + (1-u)^2x^2 - 2(1-u)vx + v^2\\
- %   {} &=& \frac{x^7}{7}-2v\frac{x^4}{4} + (1-u)^2\frac{x^3}{3} - 2(1-u)v\frac{x^2}{2}+v^2x\\
- %   {} &=& \frac{1}{7}-2v\frac{1}{4} + (1-u)^2\frac{1}{3} - 2(1-u)v\frac{1}{2}+v^2\\
- %   {} & & +\frac{1}{7}-2v\frac{1}{4} + (1-u)^2\frac{-1}{3} - 2(1-u)v\frac{1}{2}-v^2\\
- %   {} &=& \frac{2}{7} - (2-u)v+v^2 = \ell(u,v)
- % \end{eqnarray*}
- % and note that the last line above is a function of $u$ and $v$ alone, and we have to minimize it.
- % \end{example}
-
 \section{Dealing with sequences}
 
+\subsection{Basic theory} Classical neural networks are great for classification problems involving either fixed (finite) sets $f:A\to B$, where $A$ represents the set to be classified, and $B$ is the set of labels, or functions $f:\R^n\to B$.
+An example of the latter is image classification, since an $n\times m$ image can be directly represented as a vector in $\R^{\ell mn}$.
+In many cases, however, the data to be classified is better represented as a (finite) sequence.
+For example, sentences are sequences of words, words are sequences of letters, movies are sequences of images, and sound files are sequences of amplitudes.
+It is possible to deal with sequences with ordinary neural networks, but we run into difficulties when we try to model dependency between different elements of a sequence.
 Let $A$ be a finite set.
-We define the set $A^*$ of all finite sequences of elements of $A$ to be the smallest set $X$ with the property that the empty sequence $\epsilon$ is in $X$, and whenever $w\in X$ and $a\in A$, the sequence $aw\in X$.
-If $A$ and $B$ are two finite sets, and $f:A\to B$ is any function, then we can extend $f$ extends uniquely to a function $f^*:A^*\to B^*$, which is defined by $f(a_1a_2,...,a_n)=f(a_1)f(a_2)\cdots f(a_n)$.
-This $f^*$ is the only function extending $f$ which satisfies $f(vw)=f(v)f(w)$.
+A finite sequence of elements of $A$ is a tuple $(a_1,...,a_n)$, where $a_i\in A$ for every $i$.
+Here the number $n$ is allowed to change.
+The set of all finite sequences of elements of $A$ is denoted $A^*$.
+From the definition we just gave, we have $A^*=\bigcup_{n\geq 0} A^n$, which is a good enough definition for our purpose.
+If we are given two finite sets $A$ and $B$, there are many situations we can consider if we also have access to $A^*$ and $B^*$.
+\begin{enumerate}
+  \item A function $f:A\to B^*$, which given an element of $A$ outputs a sequence of elements of $B$.
+This is called a {\em one-to-many} function.
+We use this architecture in a sequence sorting neural network.
+The Tweet2Vec autoencoder also uses a one-to-many neural network.
+  \item A function $f:A^*\to B$, which is given a sequence of elments of $A$ and produces a single element of $B$.
+This is called a {\em many-to-one} function.
+A good example of sich a function is the Buzzometer sentiment analysis tool which assigns a single sentiment value to messages which can vary in length.
+  \item A function $f:A^*\to B^*$ which is given a seuqence of elements of $A$< and outputs a sequence of elements of $B$.
+This is called a {\em many-to-many} function.
+As an example of such a function we will show the implementation of a simple text generator.
+Translation, video captioning, part-of-speech tagging and speech recognition are all examples of many-to-many functions.
+\end{enumerate}
 
-Let $S$ be another finite set with a distinguished element $\perp\in S$, and let $f:A\times S\to B\times S$ be a function.
-We can use $f$ to define a function $f^*:A^*\to B^*$ in several ways.
-Consider an auxiliary function $f^*:A^*\to B^*\times S$ by defining First we define $f^*(\epsilon)=(\epsilon, \perp)$.
-If $a\in A$ then $f^*(a)=f(a, \perp)$.
-If $v=aw$, then by induction we can compute $f^*(w)=(u, s)$.
-Write $f(a, s)=(b, s')$, and define $f^*(aw)=(bu, s')$.
-We can produce a function $f^*:A^*\to B^*$ by taking $f^*(w)=\pi_2 g(w)$.
+We can of course define functions $A\to B^*$, $A^*\to B$ and $A^*\to B^*$ directly (they are just sets, after all), but it is more interesting and instructive to make use of the structure of $A^*$ and $B^*$ as sets of sequences of $A$ and $B$, and see how one can use a function (or class of functions) $A\to B$ to define a function $A^*\to B^*$.
+In particular, we will me making use of the iterative nature of the construction $A^*$ and $B^*$.
+Here is the simplest example: if $A=B$, and $f:A\to A$ is a function, then we can iterate $f$: $f^0(a)=a$ for every $a\in A$, and $f^{n+1}(a) = f(f^n(a))$.
+We terminate the iteration at the first value of $n$ for which $f^n(a)=f^{n-1}(a)$ (note that there is no guarantee that this will ever happen).
+This gives a function $f:A\to A^\omega$.
 
-We can use $f$ to define a function $f^*:A^*\to B$.
-Consider an auxiliary function $f^*:A^*\to B\times S$ by defining First we define $f^*(\epsilon)=(b_0, \perp)$.
-If $a\in A$ then $f^*(a)=f(a, \perp)$.
-If $v=aw$, then by induction we can compute $f^*(w)=(u, s)$.
-Write $f(a, s)=(b, s')$, and define $f^*(aw)=(b, s')$.
-We can produce a function $f^*:A^*\to B$ by taking $f^*(w)=\pi_2 g(w)$.
+We describe a more general situation.
+Let $S$ be a (finite) set of {\em states} with a distinguished element $\perp\in S$, and consider a function $f:A\times S\to B\times S$.
+Let $a_1,...,a_n$ be a finite sequence of elements of $A$.
+We define two sequences $b_1,...,b_n$ and $s_1,...,s_n$ of elements of $B$ and $S$ respectively as follows.
+We first define $b_1, s_1 = f(a_1, \perp)$.
+Suppose that we have defined the elements $b_1,...,b_n$ and $s_1,...,s_n$.
+We define $b_{n+1}$ and $s_{n+1}$ via $b_{n+1}, s_{n+1} = f(a_{n+1}, s_n)$.
+We define $f^*(a_1...a_n)=b_1...b_n$, with $b_1,...,b_n$ defined as above.
+This is a function $f^*:A^*\to B^*$, and it has the property that the length of the output sequence is the same as that of the input sequence.
+An interesting special case of this is when $S=B$.
+In this case, we can have $f:A\times B\to B$.
+We can define b_1 = f(a_1, \perp)$, and $b_{n+1}=f(a_{n+1}, b_n)$.
+This is in fact the main type of function we use in our examples later.
 
-Let $S$ be another finite set with a distinguished element $\perp\in S$, and let $f:A\times S\to B\times S$ be a function.
-We can use $f$ to define a function $f^*:A\to B^N$ for every $N$ in several ways.
-Consider an auxiliary function $f^*:A\to B^N\times S$ by defining.
-If $a\in A$ then $f^*(a)_1=f(a, \perp)$.
-Suppose $f^*(a)_n=(w, s)$, and  $f(a, s) = (b, s')$, then $f^*(a)_{n+1}=(bw, s')$.
-This is an example of a one-to-many function.
-If $A$ has a distinguished element $a_0$, then we can define $f^*(a)_1=f(a, \perp)$, and $f^*(a)_n=(w, s)$, and  $f(a_0, s) = (b, s')$, then $f^*(a)_{n+1}=(bw, s')$.
-This is an example of a one-to-many function.
-Note that this is a special case of the many to many function, where $f^*(a)$ is just defined as $f^*(aaa\cdots a)$, or $f^*(aa_0\cdots a_0)$.
+We use the exact same method to define a many-to-one function.
+If $f:A\times S\to B\times S$ is a function, and  $a_1,...,a_n$, $b_1,...,b_n$ and $s_1,...,s_n$ are defined as above, then we define $f^*(a_1,...,a_n)=b_n$.
+In other words, we discard all elements of $B$ produced during the iteration, and keep only the last one.
 
-%If $v=aw$, then by induction we can compute $f^*(w)=(u, s)$. Write $f(a, s)=(b, s')$, and define $f^*(aw)=(bu, %s')$.  We can produce a function $f^*:A^*\to B^*$ by taking $f^*(w)=\pi_2 g(w)$.
-%  We can then define $f^*(v)=f^*(aw)=f(a, s) = (w', s')$, or mode formally, $f^*(aw)=f(a, \pi_2f^*(w)$
-%
-%$f^*(v)=f^*()
-%
-%There is more than one way to extend a recurrent function to a set of sequences
+Finally, we tackle the case of a one-to-many function.
+Again we let $f:A\times S\to B\times S$ be any function.
+We define sequences $b_1,...,b_i,...$ and $s_1,...,s_i,...$ as follows.
+$b_1,s_1=f(a,\perp)$, and for every $n$, $b_{n+1}, s_{n+1}=f(a, s_n)$.
+In other words, we keep iterating $f$, feeding $a$ as input at every step.
+An imoprtant variant of this is when $a$ itself has a distinguished element $\perp_A$, in which case, after feeding $a$ as an input to $f$ in the first iteration, we subsequently feed it $\perp_A$, and get $b_{n+1}, s_{n+1}=f(\perp_A, s_n)$.
+We stop the iteration whenever $s_n=\perp$ (or any other predetermined element of $S$).
+Note that this may never happen, i.e.
+the sequence $b_n$ may go on forever.
+In fact, there is no way to give a general pattern of definition for $f:A\times S\to B\times S$ which will guarantee that $f^*(a)$ is finite for every $a$.
 
-\section{The base of recurrent networks}
+\subsection{The base of recurrent networks}
 
@@
 
-In general, in order to minimize a function $f(w)$ of one variable, we make use of the fact that the extreme values of $f$ all have $f'(x)=0$.
+\subsection{The standard version} In general, in order to minimize a function $f(w)$ of one variable, we make use of the fact that the extreme values of $f$ all have $f'(x)=0$.
 If $f$ is a function of two or more variables, then if $x$ is such that $f(x)$ is a local max, or minimum, then all partial derivatives $\frac{\partial f}{\partial w_i}(x)=0$.
@@
 
-\section{Backward Propagation Through Time}
-
+\subsection{Backward Propagation Through Time}
 How do we train recurrent neural networks?

### 2017-08-24 deeplearn/presentation/paper_2.tex
presentation and text generator

@@
 \newcommand{\e}{\mathbf{e}}
+\newcommand{\s}{\mathbf{s}}
 
@@
 This is in fact the main type of function we use in our examples later.
+Note that this is a state machine, close to a deterministic finite automaton, except that here the state space $S$ is allowed to be infinite.
 
@@
 
-\subsection{The base of recurrent networks}
+\subsection{Recurrent networks}
 
-Consider a function $f:U\times S\to V\times S$, where $U$, $V$ and $S$ are finite dimensional vector spaces.
-Note that every vector space has a a distinguished element $0$, the zero-vector.  Let $u_1,...,u_n$ be a finite
-sequence of vectors in $U$.  We construct a sequence of vectors in $V$ as follows. Write $f(u_1, 0)=(v_1, s_1)$,
-and for every $i$, if $f(u_{i}, s_{i-1})=(v_i, s_i)$. Note that this is a state machine, close to a
-deterministic finite automaton, except that here the state space $S$ is infinite.
-
-The set $\{0,1\}^*$ of all finite sequences of $0$'s and $1$'s is countable, and $S$, being a real vector space,
+In general, a recurrent neural network is just a presceiption for a class of functions of the form $f:\R^n\times \R^\ell\times\R^m\to \R^m\times \R^\ell$, where $\R^m$ is the space of parameters.
+The set $\{0,1\}^*$ of all finite sequences of $0$'s and $1$'s is countable, and $\R^\ell$, being a real vector space,
 is uncountable.  Therefore, we can encode every element $w\in\{0,1\}^*$ as a vector in $S$.  Informally,
-$S$ is enough to encode any finite state space, and every possible value for the content of the tape of a
+$\R^\ell$ is enough to encode any finite state space, and every possible value for the content of the tape of a
 Turing machine. We get:
@@
 
-which explains why recurrent neural networks seem to be able to produce results that other networks can't.
+This explains why recurrent neural networks seem to be able to produce results that other networks can't.
 Training a recurrent network is the same as producing a Turing machine.
+A common way to represent a recurrent node is to use a box, the inside of which represent the definition of $f$.
+The vertical arrows represent the input and onput of the node, and the horisontal arrows represent the input and output state of the node.
 
+\begin{center}
   \begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
@@
 \end{tikzpicture}
+\end{center}
 
+Just like any regular neural network layer, recurrent nodes can be composed.
+The composition of two recurrent nodes is done by feeding the output of one node into the input of the other, and concatenating their state space.
+Formally, if $f:\R^n\times \R^\ell\times\to \R^m\times \R^\ell$ and $g:\R^n\times \R^\ell\times\to \R^m\times \R^\ell$ are two recurrent nodes, then we have the node $(f;g):\R^m\times\R^\ell\times\R^L\to\R^k\times\R^\ell\times\R^L$ defined by $(f;g)(\x, \s_1, \s_2) = (\y, \s'_1, \s'_2)$ where $f(\x, \s_1) = \x', \s'_1$ and $g(\x', \s_2) = \y, \s'_2$.
+Graphically, we can represent this by stacking the boxes representing $f$ and $g$ on top of one another
+
+\begin{center}
 \begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
@@
 \end{tikzpicture}
+\end{center}
 

### 2017-08-24 deeplearn/twitter_sentiment/models/text_generator/text_generator.py

@@ -223,7 +223,7 @@ if __name__ == '__main__':
             s.save(session, os.path.join(chkpt, "t_gen_weights"))
             print('Model restored')
 
-        for batch in rnn_minibatch_sequencer(data, 32, Hyperparameters.sequence_length, nb_epochs=100):
+        for batch in rnn_minibatch_sequencer(data, 32, Hyperparameters.sequence_length, nb_epochs=500):
             x = batch['train_x']
             y = batch['train_y']
             feed_dict = {Globals.model_input: x,
@@ -237,7 +237,7 @@ if __name__ == '__main__':
                 file_name = "text_gen-{}".format(index)
                 file_name = os.path.join(save, file_name)
                 with open(file_name, 'w') as test_file:
-                    test_file.write(generate_text(20000, session))
+                    test_file.write(generate_text(2000, session))
             if index % 250 == 0:
                 s = tf.train.Saver()
                 s.save(session, os.path.join(chkpt, "t_gen_weights"))

### 2017-08-24 deeplearn/presentation/paper_2.tex
presentation

@@
 Formally, if $f:\R^n\times \R^\ell\times\to \R^m\times \R^\ell$ and $g:\R^n\times \R^\ell\times\to \R^m\times \R^\ell$ are two recurrent nodes, then we have the node $(f;g):\R^m\times\R^\ell\times\R^L\to\R^k\times\R^\ell\times\R^L$ defined by $(f;g)(\x, \s_1, \s_2) = (\y, \s'_1, \s'_2)$ where $f(\x, \s_1) = \x', \s'_1$ and $g(\x', \s_2) = \y, \s'_2$.
-Graphically, we can represent this by stacking the boxes representing $f$ and $g$ on top of one another
+Graphically, we can represent this by stacking the boxes representing $f$ and $g$ on top of one another:
 
@@
 
-\section{Back Propagation}
+The whole stack can then be seen as one single recurrent node.
+
+\section{Training}
 
@@
 
-There are three main types of recurrent cells which are used to build recurrent neural networks.
+In this section we descripbe three of the main types of recurrent cells currently in use.
+They are the basic recurrent cell, which is a straightforward implementation of a function $f:\R^n\times\R^\ell\to\R^\ell$, then long short term memory cell (LSTM), which is more suitable to model long term dependencies in sequences, and the gated recurrent unit cell, which is a simplification of the LSTM cell, but which performs almost equally well at long term dependencies.
+We give a formal definition of each oc these cells along with a na\"ive implementation in {\tt tensorflow}.
+The implementation we provide is only expositional though, and would not work out of the box in an actual production model.
+%There are three main types of recurrent cells which are used to build recurrent neural networks.
 Let us describe them, and give some of their basic properties, along with a na\"ive implementation in tensorflow.
@@
 \begin{minted}{python}
-def basic_rnn_cell(input_tensor, state_tensor):
-  A_u = tf.Variable(shape=[N, L])
-  B_u = tf.Variable(shape=[L, L])
-  A_v = tf.Variable(shape=[N, L])
-  B_v = tf.Variable(shape=[L, L])
-  output_tensor = tf.relu(tf.matmul(input_tensor, A_u) +
+def basic_rnn_cell(input_tensor, state_tensor, output_dimension):
+  input_dimension = input_tensor.get_shape()[1]
+  state_dimension = input_tensor.get_shape()[1]
+  A_u = tf.Variable(shape=[input_dimension, output_dimension])
+  B_u = tf.Variable(shape=[state_dimension, output_dimension])
+  A_v = tf.Variable(shape=[input_dimension, state_dimension])
+  B_v = tf.Variable(shape=[state_dimension, state_dimension])
+  output_tensor = tf.relu(tf.matmul(input_tensor, A_u) + \
                             tf.matmul(state_tensor, B_u))
@@
 \end{minted}
+
+A variant of this cell uses the output as a state vector, thus computing every step as a function of the cirrent ouput, and the previous output.
 
@@
 
+As an example of a many-to-many recurrent neural network, we propose an architecture that generates text in the style of a particular author, or body of text.
+The architecture is very simple, and consists of three stacked GRU cells.
+This model will also expose some of the challenges in training recurrent networks.
+As an example, we trained the model on the text of the Harry Potter series, for a total of 500 epochs.
+Every once in a while we paused the training to let it generate a sequence of about 2000 characters.
+For training we network was unrolled to 30 characters, and the training was done in batches of 32 strings.
+The main challenge for training is that the strings should continue from batch to batch.
+For example, consider training a similar network unrolled to 2 characters with a batch size of 3.
+Consider the sentence {\tt The quick brown fox jumps over the lazy dog}.
+The first step is to cut the string into substrings of length 2:
+{\tt |Th|e |qu|ic|k |br|ow|n |fo|x |ju|mp|s |ov|er| t|he| l|az|y |do|g.|}. In a normal batching situation, we would then cut this liet of strings into chunks of length 3, like so
+{\tt |(Th|e |qu)|(ic|k |br)|(ow|n |fo)|(x |ju|mp)|(s |ov|er)|( t|he| l)|(az|y |do)|} but this causes a problem.
+The first sequence of the first batch is {\tt Th}, which will leave the network in a certain state $s$.
+This For this state to be updated properly, the first element of the second batch should be {\tt |e |}, and {\em not} {\tt |ic|}.
+We must therefore use a different batching strategy.
+The string will give us 7 batches in total, so we number the subsequences with a batch index from 1 to 7 in order, starting over at 1 when we run out of indices.
+The expected output, in this case, is the same as the input but shifted by one character.
+In code, the batching looks like:
+
+\begin{minted}{python}
+def rnn_minibatch_sequencer(raw_data, batch_size, sequence_size, nb_epochs):
+    data = np.array([x for x in raw_data])
+    data_len = data.shape[0]
+    # using (data_len-1) because we must provide for the sequence shifted by 1 too
+    nb_batches = (data_len - 1) // (batch_size * sequence_size)
+
+    total_num_batches = (data_len * nb_epochs) // batch_size
+    #batches_per_epoch = len(data_x) // batch_size
+
+    assert nb_batches > 0, "Not enough data, even for a single batch. Try using a smaller batch_size."
+    rounded_data_len = nb_batches * batch_size * sequence_size
+    xdata = np.reshape(data[0:rounded_data_len], [batch_size, nb_batches * sequence_size])
+    ydata = np.reshape(data[1:rounded_data_len + 1], [batch_size, nb_batches * sequence_size])
+    I = 0
+    for epoch in range(nb_epochs):
+        for batch in range(nb_batches):
+            x = xdata[:, batch * sequence_size:(batch + 1) * sequence_size]
+            y = ydata[:, batch * sequence_size:(batch + 1) * sequence_size]
+            x = np.roll(x, -epoch, axis=0)  # to continue the text from epoch to epoch (do not reset rnn state!)
+            y = np.roll(y, -epoch, axis=0)
+            I += 1
+            yield {'train_x':  x,
+                   'train_y':  y,
+                   'validate': None,
+                   'batch_number':  batch,
+                   'epoch_number':  epoch+1,
+                   'batch_index':   I,
+                   'total_batches': total_num_batches,
+                   'total_epochs':  nb_epochs}
+\end{minted}
+
 \begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]

### 2017-08-25 deeplearn/twitter_sentiment/models/text_generator/text_generator.py
text_generator update

@@ -223,7 +223,7 @@ if __name__ == '__main__':
             s.save(session, os.path.join(chkpt, "t_gen_weights"))
             print('Model restored')
 
-        for batch in rnn_minibatch_sequencer(data, 32, Hyperparameters.sequence_length, nb_epochs=500):
+        for batch in rnn_minibatch_sequencer(data, 32, Hyperparameters.sequence_length, nb_epochs=1000):
             x = batch['train_x']
             y = batch['train_y']
             feed_dict = {Globals.model_input: x,
@@ -237,7 +237,7 @@ if __name__ == '__main__':
                 file_name = "text_gen-{}".format(index)
                 file_name = os.path.join(save, file_name)
                 with open(file_name, 'w') as test_file:
-                    test_file.write(generate_text(2000, session))
+                    test_file.write(generate_text(4000, session))
             if index % 250 == 0:
                 s = tf.train.Saver()
                 s.save(session, os.path.join(chkpt, "t_gen_weights"))

### 2017-08-28 deeplearn/presentation/paper_2.tex
Text Generator bug fix

@@
 In this case, we can have $f:A\times B\to B$.
-We can define b_1 = f(a_1, \perp)$, and $b_{n+1}=f(a_{n+1}, b_n)$.
+We can define $b_1 = f(a_1, \perp)$, and $b_{n+1}=f(a_{n+1}, b_n)$.
 This is in fact the main type of function we use in our examples later.
@@
     \tikzstyle{arrow} = [thick,->,>=stealth]
+     \tikzstyle{every pin edge}=[<-,shorten <=1pt]
+     \tikzstyle{neuron}=[circle,fill=black!50,minimum size=6pt,inner sep=0pt]
+
     \tikzstyle{startstop} = [rectangle, minimum width=1.5cm, minimum height=.6cm, text centered, draw=black, fill=black!20]
     \node (input) [] at (0cm,1cm) {};
-    \node (output) [] at (0cm,-3cm) {};
+    \node (output) [] at (0cm,-2.25cm) {};
     \node (s-input-A) [] at (-2cm,0cm) {};
@@
     \draw [arrow] (B) -- (C);
-    \draw [arrow] (C) -- (output);
+    %\draw [arrow] (C) -- (output);
 
@@
        \node (input-\i) [] at (2*\i cm,1cm) {};
-       \node (output-\i) [] at (2*\i cm,-3cm) {};
+       \node (output-\i) [] at (2*\i cm,-2.25cm) {};
        \path[] node[startstop] (A-\i) at (2*\i cm,0cm) {};
@@
        \draw [arrow] (C-\j) -- (C-\i);
-       \draw [arrow] (C-\i) -- (output-\i);
+       %5\draw [arrow] (C-\i) -- (output-\i);
      }
@@
 
+     \foreach \name / \x in {1,...,15} {
+         \path[xshift=-1cm]
+             node[neuron] (I-\name) at (0.3*\x,-3.5cm) {};
+             \path (output) edge (I-\name);
+             }
+
+     \foreach \name / \x in {1,...,15} {
+         \path[xshift=4.5cm]
+             node[neuron] (J-\name) at (0.3*\x,-3.5cm) {};
+            \path (output-4) edge (J-\name); }
+
+     \foreach \name / \x in {1,...,15}{
+         \path[xshift=1.5cm]
+             node[neuron] (K-\name) at (0.3*\x,-4.5cm) {};
+        \path (I-\name) edge (K-\name);
+        \path (J-\name) edge (K-\name); }
+
+        \foreach \name / \x in {1,...,4}{
+            \path[xshift=3cm] node[neuron] (O-\name) at (0.3*\x,-5.25cm) {};
+            \foreach \j in {1,...,15}{
+           \path (K-\j) edge (O-\name);}
+           %\path (J-\name) edge (K-\name);
+           }
+
+%     \foreach \name / \x in {1,...,5}
+%         \path[yshift=0.5cm]
+%             node[neuron] (H-\name) at (\x,-2cm) {};
+
+%\end{tikzpicture}
+%
+%
+%\begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
+    %
+    % % Draw the hidden layer nodes
+    %
+    % \foreach \source in {1,...,5}
+    %     \foreach \dest in {1,...,5}
+    %         \path (I-\source) edge (H-\dest);
+    %
+    % \foreach \name / \x in {1,...,5}
+    %     \path[yshift=0.5cm]
+    %         node[neuron] (K-\name) at (\x,-4 cm) {};
+    %
+    % \foreach \source in {1,...,5}
+    %     \foreach \dest in {1,...,5}
+    %         \path (H-\source) edge (K-\dest);
 \end{tikzpicture}
 
-\begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
-    \tikzstyle{every pin edge}=[<-,shorten <=1pt]
-    \tikzstyle{neuron}=[circle,fill=black!50,minimum size=10pt,inner sep=0pt]
-
-    % Draw the hidden layer nodes
-    \foreach \name / \x in {1,...,5}
-        \path[yshift=0.5cm]
-            node[neuron] (I-\name) at (\x,-0cm) {};
-
-    \foreach \name / \x in {1,...,5}
-        \path[yshift=0.5cm]
-            node[neuron] (H-\name) at (\x,-2cm) {};
-
-    \foreach \source in {1,...,5}
-        \foreach \dest in {1,...,5}
-            \path (I-\source) edge (H-\dest);
-
-    \foreach \name / \x in {1,...,5}
-        \path[yshift=0.5cm]
-            node[neuron] (K-\name) at (\x,-4 cm) {};
-
-    \foreach \source in {1,...,5}
-        \foreach \dest in {1,...,5}
-            \path (H-\source) edge (K-\dest);
-\end{tikzpicture}
-
-The base input ofor this network is a (unicode) character, which is one-hot encoded before being fed into the
+The base input for this network is a (unicode) character, which is one-hot encoded before being fed into the
  recurrent layers.  For training, the network is unrolled to a length of 256 characters,
@@
     \tikzstyle{startstop} = [rectangle, minimum width=1.5cm, minimum height=.6cm, text centered, draw=black, fill=black!20]
+    \tikzstyle{neuron}=[circle,fill=black!50,minimum size=6pt,inner sep=0pt]
+
     \node (input) [] at (0cm,1cm) {};
@@
      \draw [arrow] (C-4) -- (s-output-C);
+
+     \foreach \name / \x in {1,...,15}{
+         \path[xshift=0cm]
+             node[neuron] (K-\name) at (0.5*\x,-3.5cm) {};
+        \path (output-2) edge (K-\name);
+        %\path (J-\name) edge (K-\name);
+        }
+
+        \foreach \name / \x in {1,...,10}{
+            \path[xshift=1cm] node[neuron] (O-\name) at (0.6*\x,-5.75cm) {};
+            \foreach \j in {1,...,15}{
+           \path (K-\j) edge (O-\name);}
+           %\path (J-\name) edge (K-\name);
+           }
 
@@
     \draw [arrow] (s-input-C) -- (C-0);
-    \draw [arrow] (A) -- (B);
-    \draw [arrow] (B) -- (C);
-    \draw [arrow] (C) -- (output);
+    \draw [arrow] (A-0) -- (B-0);
+    \draw [arrow] (B-0) -- (C-0);
+    \draw [arrow] (C-0) -- (output);
 

### 2017-08-28 deeplearn/presentation/presentation_slides.tex
Text Generator bug fix

--- 2017-08-24 deeplearn/presentation/paper_2.tex
@@
-\documentclass[10pt]{amsart}
+\documentclass[10pt]{beamer}
 \usepackage{amsmath}
@@
 \usetikzlibrary{shapes.geometric, arrows}
+\usepackage{listings}
 
@@
 
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{corollary}[theorem]{Corollary}
-
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{example}[theorem]{Example}
+%\newtheorem{theorem}{Theorem}[section]
+%\newtheorem{proposition}[theorem]{Proposition}
+%\newtheorem{lemma}[theorem]{Lemma}
+%\newtheorem{corollary}[theorem]{Corollary}
+%
+%
+%\theoremstyle{definition}
+%\newtheorem{definition}[theorem]{Definition}
+%\newtheorem{example}[theorem]{Example}
 
@@
 
- \section{Introduction}
+\begin{frame}{Training RNN's}
+\end{frame}
 
-Throughout this handout, we use bold face lower case letters to represent vectors, and bold face capital letters to represent matrices.
-Most of the time, we will blur the dustinction between matrices and vectors, except when it makes the exposition clearer.
+\begin{frame}{The Basic RNN Cell}
+\end{frame}
 
- Let $f:\R^n\to \R^m$ be a function, and consider a class $\C$ of functions $g: \R^n\to \R^m$, which we assume to be parametrized by $\R^\ell$, in the sense that there is a function $F(\x,\w):\R^n\times \R^\ell\to\R^n$ such that $\C=\{\x\mapsto F(\x,\w):\w\in\R^n\}$.
-The generic problem of machine learning is to find a vector $\w\in\R^{\ell}$ such that $D(f(\x), F(\x, \w))$ is as small as possible, where $D$ represents a notion of distance between functions, for example, $$D(f(\x), F(\x, \w)) = \min\{\|f(\x)-F(\x, \w)\|:\x\in\R^n\}$$ Note that since the minimum is computed over all possible vlues of $\x$, the value of $D$ in te equation above depends on $\w$.
-Define $$\ell(\w)=D(f(\x), F(\x, \w))$$ Given the definition of $D$, we see that $\ell(\x)\geq 0$, and therefore $\ell$ has a global minimum (which is not necessarily unique).
-In general, the global minimum of $\ell$ is not attained, in the sense that if $\epsilon=\min\{\ell(\w):\x\in\R^n\}$, then there does not necessarily exist a value $\w_0\in\R^\ell$ such that $\epsilon=\ell(\w_0)$.
-However, it can be approximated, that is to say there is a sequence $\w_i$ such that $\ell(\w_i)\to\epsilon$ as $i\to\infty$.
-We may not be able to achieve the minimum with a specific value of $\w$, but we can get as close to it as we want: for every $\varepsilon>0$, there is some natural number $N_\varepsilon>0$ such that $\|\epsilon-\ell(\w_i)\|<\varepsilon$ for every $i>N_\varepsilon$.
-Since we can choose $\varepsilon$ to be as small as we want, if we choose it to be smaller than the smallest number which can be represented with a {\tt float32} (say), then from the point of view of any program, $\ell(\w_i)=\epsilon$ when $i>N_\varepsilon$.
-
- The function $\ell$ we defined above is called a {\em loss function}.
-There are many functions which are suitable for $\ell$.
-In fact, $\ell$ does not need to represent a distance at all.
-In order to draw a meaningful conclusion from the minimum value of $\ell$, all we need to know is that $\ell(\w)=0$ implies $f(\x)=F(\x,\w)$ for every $\x$, or at the very least $\|f(\x)-F(\x,\w)\|\leq\varepsilon$ for arbitrary $\varepsilon$.
-In every neural network we define, we specify a class of functions $F(\x, \w)$, and a definition for $\ell(\w)$ based on $F$.
-In tensorflow, the function we are trying to approximate is given in the form of a relation $f(\x)=\y$.
-We provide {\em placeholders} for the values of $\x$ and $\y$.
-The function $F(\x,\w)$ is defined by a computation graph.
-For example, here is a linear regression, in which we try to approximate a function $f:\R^{10}\to\R^{15}$ using a function of the form $F(\x, \W\b)=\W\x+\b$.
-Here $\W$ is an $15\times 10$ matrix, and $\b\in\R^{15}$
-
- \begin{minted}{python}
- x = tf.placeholder(shape=[None, 10])
- y = tf.placeholder(shape=[None, 15])
- W = tf.Variable(shape=[None, 10, 15])
- b = tf.Variable(shape=[None, 15])
- F = tf.matmul(x, W) + b
- loss = tf.mean_square_error(y, F)
- \end{minted}
-
-We are mainly interested in classification problems, or more generally problems which can be expressed as classification problems.
-Let $A$ and $B$ be finite sets, and let $f:A\to B$ be any function.
-In order to use neural networks, which represent functions between topological vector spaces, to represent  function between ordinary sets, we must first translate the sets $A$ and $B$ into vector spaces.
-The most straightforward (and freest) way to interpret a set $A$ as a vector space is to use $A$ as the basis for a vector space.
-Write $A=\{a_1,...,a_n\}$, and let $\{\e_1,...,\e_n\}$ represent the standard basis for $\R^n$.
-Let $e:A\to\R^n$ be defined by $e(a_i)=\e_i$.
-Similarly, we can write $B=\{b_1,...,b_m\}$, and define $e:B\to \R^m$ by $e(b_i)=\e_i$.
-This specific function is called {\em one-hot encoding}, and is present in tensorflow as tf.onehot.
-Since we are mapping a finite set to an infinte set, the function $e$ has no inverse, but it has a "pseudo-inverse" given by the {\em argmax()} function defined by $argmax(\v)=i$ where $i$ is smallest such that $v_i\geq v_j$ for every $j\not=i$.
-In other words, $argmax$ computes the index of the largest component of a vector.
-
-Given $\v\in\R^m$ is the output of $F(\x,\w)$, then if $\v$ has the property that $\sum v_i=1$, and $v_i\geq 0$ for every $i$, then we can interpret $\v$ as a probability distribution, where each component $v_i$ represents the probability that $\v=\e_i$.
-The code, in tensorflow, which approximates a function $A\to B$ can be written as:
-
-\begin{minted}{python}
-x_in = tf.placeholder(shape=[None, 1])
-x = tf.one_hot(x_in, 10)
-y_in = tf.placeholder(shape=[None, 1])
-y = tf.one_hot(y_in, 15)
-W = tf.Variable(shape=[None, 10, 15])
-b = tf.Variable(shape=[None, 15])
-F = tf.softmax(tf.matmul(x, W) + b)
-loss = tf.cross_entropy(y, F)
-\end{minted}
-
-The softmax function in the penultimate line is defined by $\sigma(\v) = \w$  where $$w_i= \frac{e^{v_i}}{\sum e^{v_j}}$$ Note that $\sum w_j=1$, which means that the output of the softmax function can be interpreted as a probability distribution.
-This makes it a popular choice for classification problems with many classes.
-If the target space is $\R^2$, then this reduces to a binomial distribution.
-The cross-entropy function in the last line is defined by $C(\v, \w) = -\sum(v_i\log(w_i))$.
-Intuitively, $C$ measures the number of vector components that are necessary to differentiate between the distribution $\v$ and the distribution $\w$.
-Therefore, the optimal value for $C$ to have is $N$: measuring any bit should allow us to differentiate between $\v$ and $\w$.
-Note that $C(\v, \w)=1$ if and only if $\log C(\v, \w)=0$, so that the logarithm of the cross entropy function is a candidate for a distance function.
-The vector $\w$ here should be the output of the model, whereas $\v$ should be the true value.
-By definition, $w_i= \frac{e^{v_i}}{\sum e^{v_j}}$, so that $\log(w_i) = \log e^{w_i} - \log(\sum e^{v_j}) = w_i - \log(\sum e^{v_j})$.
-By definition again, $v_i=1$ and $v_j=0$ if $i\not=j$.
-Therefore, $C(\v, \w) = w_i + \log(\sum e^{w_j})$
-
-\section{Dealing with sequences}
-
-\subsection{Basic theory} Classical neural networks are great for classification problems involving either fixed (finite) sets $f:A\to B$, where $A$ represents the set to be classified, and $B$ is the set of labels, or functions $f:\R^n\to B$.
-An example of the latter is image classification, since an $n\times m$ image can be directly represented as a vector in $\R^{\ell mn}$.
-In many cases, however, the data to be classified is better represented as a (finite) sequence.
-For example, sentences are sequences of words, words are sequences of letters, movies are sequences of images, and sound files are sequences of amplitudes.
-It is possible to deal with sequences with ordinary neural networks, but we run into difficulties when we try to model dependency between different elements of a sequence.
-Let $A$ be a finite set.
-A finite sequence of elements of $A$ is a tuple $(a_1,...,a_n)$, where $a_i\in A$ for every $i$.
-Here the number $n$ is allowed to change.
-The set of all finite sequences of elements of $A$ is denoted $A^*$.
-From the definition we just gave, we have $A^*=\bigcup_{n\geq 0} A^n$, which is a good enough definition for our purpose.
-If we are given two finite sets $A$ and $B$, there are many situations we can consider if we also have access to $A^*$ and $B^*$.
-\begin{enumerate}
-  \item A function $f:A\to B^*$, which given an element of $A$ outputs a sequence of elements of $B$.
-This is called a {\em one-to-many} function.
-We use this architecture in a sequence sorting neural network.
-The Tweet2Vec autoencoder also uses a one-to-many neural network.
-  \item A function $f:A^*\to B$, which is given a sequence of elments of $A$ and produces a single element of $B$.
-This is called a {\em many-to-one} function.
-A good example of sich a function is the Buzzometer sentiment analysis tool which assigns a single sentiment value to messages which can vary in length.
-  \item A function $f:A^*\to B^*$ which is given a seuqence of elements of $A$< and outputs a sequence of elements of $B$.
-This is called a {\em many-to-many} function.
-As an example of such a function we will show the implementation of a simple text generator.
-Translation, video captioning, part-of-speech tagging and speech recognition are all examples of many-to-many functions.
-\end{enumerate}
-
-We can of course define functions $A\to B^*$, $A^*\to B$ and $A^*\to B^*$ directly (they are just sets, after all), but it is more interesting and instructive to make use of the structure of $A^*$ and $B^*$ as sets of sequences of $A$ and $B$, and see how one can use a function (or class of functions) $A\to B$ to define a function $A^*\to B^*$.
-In particular, we will me making use of the iterative nature of the construction $A^*$ and $B^*$.
-Here is the simplest example: if $A=B$, and $f:A\to A$ is a function, then we can iterate $f$: $f^0(a)=a$ for every $a\in A$, and $f^{n+1}(a) = f(f^n(a))$.
-We terminate the iteration at the first value of $n$ for which $f^n(a)=f^{n-1}(a)$ (note that there is no guarantee that this will ever happen).
-This gives a function $f:A\to A^\omega$.
-
-We describe a more general situation.
-Let $S$ be a (finite) set of {\em states} with a distinguished element $\perp\in S$, and consider a function $f:A\times S\to B\times S$.
-Let $a_1,...,a_n$ be a finite sequence of elements of $A$.
-We define two sequences $b_1,...,b_n$ and $s_1,...,s_n$ of elements of $B$ and $S$ respectively as follows.
-We first define $b_1, s_1 = f(a_1, \perp)$.
-Suppose that we have defined the elements $b_1,...,b_n$ and $s_1,...,s_n$.
-We define $b_{n+1}$ and $s_{n+1}$ via $b_{n+1}, s_{n+1} = f(a_{n+1}, s_n)$.
-We define $f^*(a_1...a_n)=b_1...b_n$, with $b_1,...,b_n$ defined as above.
-This is a function $f^*:A^*\to B^*$, and it has the property that the length of the output sequence is the same as that of the input sequence.
-An interesting special case of this is when $S=B$.
-In this case, we can have $f:A\times B\to B$.
-We can define b_1 = f(a_1, \perp)$, and $b_{n+1}=f(a_{n+1}, b_n)$.
-This is in fact the main type of function we use in our examples later.
-Note that this is a state machine, close to a deterministic finite automaton, except that here the state space $S$ is allowed to be infinite.
-
-We use the exact same method to define a many-to-one function.
-If $f:A\times S\to B\times S$ is a function, and  $a_1,...,a_n$, $b_1,...,b_n$ and $s_1,...,s_n$ are defined as above, then we define $f^*(a_1,...,a_n)=b_n$.
-In other words, we discard all elements of $B$ produced during the iteration, and keep only the last one.
-
-Finally, we tackle the case of a one-to-many function.
-Again we let $f:A\times S\to B\times S$ be any function.
-We define sequences $b_1,...,b_i,...$ and $s_1,...,s_i,...$ as follows.
-$b_1,s_1=f(a,\perp)$, and for every $n$, $b_{n+1}, s_{n+1}=f(a, s_n)$.
-In other words, we keep iterating $f$, feeding $a$ as input at every step.
-An imoprtant variant of this is when $a$ itself has a distinguished element $\perp_A$, in which case, after feeding $a$ as an input to $f$ in the first iteration, we subsequently feed it $\perp_A$, and get $b_{n+1}, s_{n+1}=f(\perp_A, s_n)$.
-We stop the iteration whenever $s_n=\perp$ (or any other predetermined element of $S$).
-Note that this may never happen, i.e.
-the sequence $b_n$ may go on forever.
-In fact, there is no way to give a general pattern of definition for $f:A\times S\to B\times S$ which will guarantee that $f^*(a)$ is finite for every $a$.
-
-\subsection{Recurrent networks}
-
-In general, a recurrent neural network is just a presceiption for a class of functions of the form $f:\R^n\times \R^\ell\times\R^m\to \R^m\times \R^\ell$, where $\R^m$ is the space of parameters.
-The set $\{0,1\}^*$ of all finite sequences of $0$'s and $1$'s is countable, and $\R^\ell$, being a real vector space,
-is uncountable.  Therefore, we can encode every element $w\in\{0,1\}^*$ as a vector in $S$.  Informally,
-$\R^\ell$ is enough to encode any finite state space, and every possible value for the content of the tape of a
-Turing machine. We get:
-
-\begin{theorem}
-Recurrent neural networks are Turing-complete.
-\end{theorem}
-
-This explains why recurrent neural networks seem to be able to produce results that other networks can't.
-Training a recurrent network is the same as producing a Turing machine.
-A common way to represent a recurrent node is to use a box, the inside of which represent the definition of $f$.
-The vertical arrows represent the input and onput of the node, and the horisontal arrows represent the input and output state of the node.
-
-\begin{center}
-  \begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
-    \tikzstyle{arrow} = [thick,->,>=stealth]
-    \tikzstyle{startstop} = [rectangle, minimum width=2cm, minimum height=.75cm, text centered, draw=black, fill=black!20]
-    \node (input) [] at (0cm,1cm) {};
-    \node (output) [] at (0cm,-1cm) {};
-    \node (s-input) [] at (-2cm,0cm) {};
-    \node (s-output) [] at (2cm,0cm) {};
-    \path[] node[startstop] (N) at (0cm,0cm) {};
-    \draw [arrow] (input) -- (N);
-    \draw [arrow] (s-input) -- (N);
-    \draw [arrow] (N) -- (output);
-    \draw [arrow] (N) -- (s-output);
-\end{tikzpicture}
-\end{center}
-
-Just like any regular neural network layer, recurrent nodes can be composed.
-The composition of two recurrent nodes is done by feeding the output of one node into the input of the other, and concatenating their state space.
-Formally, if $f:\R^n\times \R^\ell\times\to \R^m\times \R^\ell$ and $g:\R^n\times \R^\ell\times\to \R^m\times \R^\ell$ are two recurrent nodes, then we have the node $(f;g):\R^m\times\R^\ell\times\R^L\to\R^k\times\R^\ell\times\R^L$ defined by $(f;g)(\x, \s_1, \s_2) = (\y, \s'_1, \s'_2)$ where $f(\x, \s_1) = \x', \s'_1$ and $g(\x', \s_2) = \y, \s'_2$.
-Graphically, we can represent this by stacking the boxes representing $f$ and $g$ on top of one another
-
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
-    \tikzstyle{arrow} = [thick,->,>=stealth]
-    \tikzstyle{startstop} = [rectangle, minimum width=1.5cm, minimum height=.6cm, text centered, draw=black, fill=black!20]
-    \node (input) [] at (0cm,1cm) {};
-    \node (output) [] at (0cm,-3cm) {};
-    \node (s-input-A) [] at (-2cm,0cm) {};
-    \node (s-output-A) [] at (2cm,0cm) {};
-    \node (s-input-B) [] at (-2cm,-1cm) {};
-    \node (s-output-B) [] at (2cm,-1cm) {};
-    \node (s-input-C) [] at (-2cm,-2cm) {};
-    \node (s-output-C) [] at (2cm,-2cm) {};
-    \path[] node[startstop] (A) at (0cm,0cm) {};
-    \path[] node[startstop] (B) at (0cm,-1cm) {};
-    \path[] node[startstop] (C) at (0cm,-2cm) {};
-    \draw [arrow] (input) -- (A);
-    \draw [arrow] (s-input-A) -- (A);
-    \draw [arrow] (s-input-B) -- (B);
-    \draw [arrow] (s-input-C) -- (C);
-    \draw [arrow] (A) -- (s-output-A);
-    \draw [arrow] (B) -- (s-output-B);
-    \draw [arrow] (C) -- (s-output-C);
-    \draw [arrow] (A) -- (B);
-    \draw [arrow] (B) -- (C);
-    \draw [arrow] (C) -- (output);
-\end{tikzpicture}
-\end{center}
-
-\section{Back Propagation}
-
-\subsection{The standard version} In general, in order to minimize a function $f(w)$ of one variable, we make use of the fact that the extreme values of $f$ all have $f'(x)=0$.
-If $f$ is a function of two or more variables, then if $x$ is such that $f(x)$ is a local max, or minimum, then all partial derivatives $\frac{\partial f}{\partial w_i}(x)=0$.
-If $f:\R^n\to\R$ is a function, then the {\em gradient} of $f$ is the vector $\nabla f(x) = \left(\frac{\partial f}{\partial w_1}(x),...,\frac{\partial f}{\partial w_n}(x)\right)$.
-One interesting property of $\nabla f(x)$ is that, as a vector, it ``points'' in the direction in which $f$ decreases the most.
-That is to say, there is a good chance that $f(x-\alpha\nabla f(x))\leq f(x)$.
-The chance is better the smaller $\alpha$ is.
-We see that if we iterate that process, we get a sequence $x_1,...,x_n,...$ such that $f(x_{i+1})\leq f(x_i)$ for every $i$.
-Since $m\leq f(x_i)\leq f(x_0)$ for every $i$, we get a bounded decreasing sequence of real number, which converges.
-
-Caveat: this ideal situation doesn't allways happen.
-
-This iterative process is {\em (Stocastic) gradient descent}.
-
-Suppose $f$ is a complsition of two functions, and that we can write $f(x,w)= g(h(x, w_1), w_2)$.
-Then we can write $\nabla_w f(x, w)=\nabla g(h(x, w_1), w_2)\cdot \nabla h(x, w_1)$. Each level
-of composition corresponds to a layer of neural network.  The chain rule transforms function composition
-into a product. If a network becomes deep, then the derivative of the loss function becomes a long
-product.  When gradients become small in norm (like they do when we approach a min), then the gradient
-becomes very close to $0$, and the update rule for gradient descent stops changing the weights.
-
-This is the vanishing gradient problem for very deep networks, and makes convergence slow.
-
-\subsection{Backward Propagation Through Time}
-How do we train recurrent neural networks?
-we can use back propagation, just like a regular neural network.
-Heuristically, we unroll the recurrent network ``infinitely'' many times, until it looks like an ordinary very deep neural network.
-In practice, since computers only have a finite amount of resources, we only unroll the network a large but finite number of times, and treat it like an ordinary neural network.
-
-\begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
-    \tikzstyle{arrow} = [thick,->,>=stealth]
-    \tikzstyle{startstop} = [rectangle, minimum width=1.5cm, minimum height=.6cm, text centered, draw=black, fill=black!20]
-    \node (input) [] at (0cm,1cm) {};
-    \node (output) [] at (0cm,-1cm) {};
-    \node (s-input-A) [] at (-2cm,0cm) {};
-    \node (s-output-A) [] at (10cm,0cm) {};
-    %\node (s-input-B) [] at (-2cm,-1cm) {};
-    %\node (s-output-B) [] at (10cm,-1cm) {};
-    %\node (s-input-C) [] at (-2cm,-2cm) {};
-    %\node (s-output-C) [] at (10cm,-2cm) {};
-    \path[] node[startstop] (A-0) at (0cm,0cm) {};
-    %\path[] node[startstop] (B-0) at (0cm,-1cm) {};
-    %\path[] node[startstop] (C-0) at (0cm,-2cm) {};
-    \draw [arrow] (input) -- (A);
-    \draw [arrow] (s-input-A) -- (A-0);
-    %\draw [arrow] (s-input-B) -- (B-0);
-    %\draw [arrow] (s-input-C) -- (C-0);
-    %\draw [arrow] (A) -- (B);
-    %\draw [arrow] (B) -- (C);
-    \draw [arrow] (A) -- (output);
-
-     \foreach \i/\j in {1/0,2/1,3/2,4/3}
-     {
-       \node (input-\i) [] at (2*\i cm,1cm) {};
-       \node (output-\i) [] at (2*\i cm,-1cm) {};
-       \path[] node[startstop] (A-\i) at (2*\i cm,0cm) {};
-    %   \path[] node[startstop] (B-\i) at (2*\i cm,-1cm) {};
-  %     \path[] node[startstop] (C-\i) at (2*\i cm,-2cm) {};
-  %     \draw [arrow] (A-\i) -- (B-\i);
-%       \draw [arrow] (B-\i) -- (C-\i);
-       \draw [arrow] (input-\i) -- (A-\i);
-       \draw [arrow] (A-\j) -- (A-\i);
-%       \draw [arrow] (B-\j) -- (B-\i);
-%       \draw [arrow] (C-\j) -- (C-\i);
-       \draw [arrow] (A-\i) -- (output-\i);
-     }
-     \draw [arrow] (A-4) -- (s-output-A);
-     %\draw [arrow] (B-4) -- (s-output-B);
-     %\draw [arrow] (C-4) -- (s-output-C);
-
-\end{tikzpicture}
-
-\section{Types of Recurrent Cells}
-
-There are three main types of recurrent cells which are used to build recurrent neural networks.
-Let us describe them, and give some of their basic properties, along with a na\"ive implementation in tensorflow.
-
-\subsection{Basic Recurrent Cell} We begin with the most basic of recurrent cell.
-Abstractly, a function $f:U\times S\to V\times S$ can be defined using two functions $u:\R^n\times \R^\ell\to \R^n$ and $v:\R^n\times\R^\ell \to \R^\ell$.
-We se ethe former function as providing the output, and the latter function as a state updating function.
-The most common definition for $u$ and $v$ for basic recurrent cells is as follows: $u(x, s) = f(A_ux+B_us)$, where $A$ nad $B$ are matrices, and $f$ is a non-linear function, and $v(x, s) = \tanh(A_vx+B_vs)$.
-Te parameters in this definition are $A_u$, $B_u$, $A_v$ and $B_v$.
-In tehsorflow, the code looks like:
-
-\begin{minted}{python}
-def basic_rnn_cell(input_tensor, state_tensor):
-  A_u = tf.Variable(shape=[N, L])
-  B_u = tf.Variable(shape=[L, L])
-  A_v = tf.Variable(shape=[N, L])
-  B_v = tf.Variable(shape=[L, L])
-  output_tensor = tf.relu(tf.matmul(input_tensor, A_u) +
+\defverbatim[colored]\lstI{
+\begin{lstlisting}[language=Python,basicstyle=\ttfamily,keywordstyle=\color{red}]
+def basic_rnn_cell(input_tensor, state_tensor, output_dimension):
+  input_dimension = input_tensor.get_shape()[1]
+  state_dimension = input_tensor.get_shape()[1]
+  A_u = tf.Variable(shape=[input_dimension, output_dimension])
+  B_u = tf.Variable(shape=[state_dimension, output_dimension])
+  A_v = tf.Variable(shape=[input_dimension, state_dimension])
+  B_v = tf.Variable(shape=[state_dimension, state_dimension])
+  output_tensor = tf.relu(tf.matmul(input_tensor, A_u) + \
                             tf.matmul(state_tensor, B_u))
@@
   return output_tensor, new_state_tensor
-\end{minted}
+\end{lstlisting}
+}
+\begin{frame}{The Basic RNN Cell (code)}
+\lstI
+\end{frame}
 
-\subsection{Long Short-term Memory Cell} There are a few problems with the architecture described above.
-First, it suffers greatly from the vanishing gradient problem, since there is no way to prevent gradients from becoming very small.
-Secondly, simple recurrent networks have a hard time remembering facts about the input sequence.
-To remedy this situation, long short-term memory cells were introduced.
-The overall structure of an LSTM cell is almost the same as the basic RNN cell.
-This time we take the previous output into account, which gives the structure as $f:\R^n\times \R^m\times \R^{\ell}\to \R^m\times \R^{\ell}$ which can be divided as two function $u:\R^{n}\times\R^m\times \R^{\ell}\to \R^m$ and $v:\R^n\times \R^m\times\R^{\ell}\to \R^{\ell}$.
-To define the functione $u$ and $v$, we use auxiliary functions $F$, $I$, and $O$ defined as follows
-\begin{eqnarray*}
-  F(x, y) &=& \sigma(A_Fx + B_Fy + b_F)\\
-  I(x, y) &=& \sigma(A_Ix + B_Iy + b_I)\\
-  O(x, y) &=& \sigma(A_Ox + B_Oy + b_O)\\
-  S(x, y) &=& \sigma(A_Ox + B_Oy + b_O)\\
-\end{eqnarray*}
-The state update is given by $v(x, y, s) = F(x, y)\circ s + I(x, y)\circ \sigma_v(A_vx + B_vy + b_v)$, where $\circ$ denotes pointwise multiplication of vectors, and finally, the output can be defined as $u(x, y, s)=O(x, y)\circ \sigma_u(v(x, y, s))$.
+\begin{frame}{The LSTM Cell}
+\end{frame}
 
-\begin{minted}{python}
+\defverbatim[colored]\lstI{
+\begin{lstlisting}[language=Python,basicstyle=\ttfamily,keywordstyle=\color{red}]
 def lstm_gate(input_tensor, previous_output, port_op):
@@
   return output, new_state
-\end{minted}
+\end{lstlisting}
+}
+\begin{frame}{The LSTM Cell (code)}
+\lstI
+\end{frame}
 
-%The function $u$ is defined by $u(x, o, s)=o()$
+\begin{frame}{The Basic GRU Cell}
+\end{frame}
 
-%: a function $f:\R^n\times \R^{\ell}\to \R^m\times \R^{\ell}$, which can be divided as two function $u:\R^{n}\times \R^{\ell}\to \R^m$ and $v:\R^n\times\R^{\ell}\to \R^{\ell_1}$.
-We can se the function $u$, like before, as providing the output.
-The function $v$ can be seen as the state update function.
-The function $u$ is defined by $u(x, s)=\sigma(A_U x+)$
-
-% above, but the state space is further divided as $\R^\ell=\R^{\ell_1}\times\R^{\ell_2}$, which gives the cell the structure of a function $f:\R^n\times \R^{\ell_1}\times\R^{\ell_2}\to \R^m\times \R^{\ell_1}\times\R^{\ell_2}$, which can be divided as three function $u:\R^{n}\times \R^{\ell_1}\times\R^{\ell_2}\to \R^m$,  $v:\R^n\times\R^{\ell_1}\times\R^{\ell_2} \to \R^{\ell_1}$ and $w:\R^n\times\R^{\ell_1}\times\R^{\ell_2} \to \R^{\ell_2}$.
-We can se the function $u$, like before, as providing the output.
-The function $v$ can be seen as the state update function, and the function $w$ controls a memory tape.
-
-\subsection{Gated Recurrent Unit Cell} A common variant of the long short term memory cell is the {\em gated recurrent unit cell}, more commonly known as GRU cells.
-The philosophy behind their design is similar to the long short term memory.
-Once again, each step of the computation takes into account a state vector and the output of the previous iteration.
-GRU's are functions  $f:\R^n\times \R^m\times \R^{\ell}\to \R^m\times \R^{\ell}$,  which can be divided as two function $u:\R^{n}\times\R^m\times \R^{\ell}\to \R^m$ and $v:\R^n\times \R^m\times\R^{\ell}\to \R^{\ell}$.
-To define the functione $u$ and $v$, we use auxiliary functions $U$ and  $R$ defined as follows
-\begin{eqnarray*}
-  U(x, y) &=& \sigma(A_Ux + B_Uy + b_U)\\
-  R(x, y) &=& \sigma(A_Rx + B_Ry + b_R)\\
-\end{eqnarray*}
-The state update is given by $v(x, y, s) = U(x, y)\circ s + (1-s)\circ \sigma_h(A_v x + B_v(R(x, y)\circ y) + b_v)$.
-
-\begin{minted}{python}
+\defverbatim[colored]\lstI{
+\begin{lstlisting}[language=Python,basicstyle=\ttfamily,keywordstyle=\color{red}]
 def gru_gate(input_tensor, previous_output, port_op):
@@
   return tf.mul(R, output) + tf.mul((1-R)O)
-\end{minted}
+\end{lstlisting}
+}
+\begin{frame}{The GRU Cell (code)}
+\lstI
+\end{frame}
 
-%$$
-%$$
-%
-%I(x, y)\circ \sigma_v(A_vx + B_vy + b_v)$, where $\circ$ denotes pointwise multiplication of vectors, and finally, the output can be defined as $u(x, y, s)=O(x, y)\circ \sigma_u(v(x, y, s))$.
+\begin{frame}{A many-to-one example}
+\end{frame}
 
-\section{Examples}
+\begin{frame}{A many-to-one example (model architecture)}
+  \begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
+      \tikzstyle{arrow} = [thick,->,>=stealth]
+       \tikzstyle{every pin edge}=[<-,shorten <=1pt]
+       \tikzstyle{neuron}=[circle,fill=black!50,minimum size=6pt,inner sep=0pt]
 
-\subsection{Many-to-One}
+      \tikzstyle{startstop} = [rectangle, minimum width=1.5cm, minimum height=.6cm, text centered, draw=black, fill=black!20]
+      \node (input) [] at (0cm,1cm) {};
+      \node (output) [] at (0cm,-2.25cm) {};
+      \node (s-input-A) [] at (-2cm,0cm) {};
+      \node (s-output-A) [] at (10cm,0cm) {};
+      \node (s-input-B) [] at (-2cm,-1cm) {};
+      \node (s-output-B) [] at (10cm,-1cm) {};
+      \node (s-input-C) [] at (-2cm,-2cm) {};
+      \node (s-output-C) [] at (10cm,-2cm) {};
+      \path[] node[startstop] (A-0) at (0cm,0cm) {};
+      \path[] node[startstop] (B-0) at (0cm,-1cm) {};
+      \path[] node[startstop] (C-0) at (0cm,-2cm) {};
+      \draw [arrow] (input) -- (A-0);
+      \draw [arrow] (s-input-A) -- (A-0);
+      \draw [arrow] (s-input-B) -- (B-0);
+      \draw [arrow] (s-input-C) -- (C-0);
+      \draw [arrow] (A-0) -- (B-0);
+      \draw [arrow] (B-0) -- (C-0);
+      %\draw [arrow] (C) -- (output);
 
-The Buzzometer sentiment analysis tool uses a bi-directional GRU model, in which the output a 4-layer
-bidirectional recurrent network is fed into a two-layer fully connected network which separates the input into
-four classes, corresponding to neutral, positive, negative and irrelevant messages. The only non-linearities
- in the network are inside the GRU cells.  Graphically, the network can be represented as:
+       \foreach \i/\j in {1/0,2/1,3/2,4/3}
+       {
+         \node (input-\i) [] at (2*\i cm,1cm) {};
+         \node (output-\i) [] at (2*\i cm,-2.25cm) {};
+         \path[] node[startstop] (A-\i) at (2*\i cm,0cm) {};
+         \path[] node[startstop] (B-\i) at (2*\i cm,-1cm) {};
+         \path[] node[startstop] (C-\i) at (2*\i cm,-2cm) {};
+         \draw [arrow] (A-\i) -- (B-\i);
+         \draw [arrow] (B-\i) -- (C-\i);
+         \draw [arrow] (input-\i) -- (A-\i);
+         \draw [arrow] (A-\j) -- (A-\i);
+         \draw [arrow] (B-\j) -- (B-\i);
+         \draw [arrow] (C-\j) -- (C-\i);
+         %5\draw [arrow] (C-\i) -- (output-\i);
+       }
+       \draw [arrow] (A-4) -- (s-output-A);
+       \draw [arrow] (B-4) -- (s-output-B);
+       \draw [arrow] (C-4) -- (s-output-C);
 
-%(GRAPHICAL REPRESENTATION OF NETWORK)
+       \foreach \name / \x in {1,...,15} {
+           \path[xshift=-1cm]
+               node[neuron] (I-\name) at (0.3*\x,-3.5cm) {};
+               \path (output) edge (I-\name);
+               }
+
+       \foreach \name / \x in {1,...,15} {
+           \path[xshift=4.5cm]
+               node[neuron] (J-\name) at (0.3*\x,-3.5cm) {};
+              \path (output-4) edge (J-\name); }
+
+       \foreach \name / \x in {1,...,15}{
+           \path[xshift=1.5cm]
+               node[neuron] (K-\name) at (0.3*\x,-4.5cm) {};
+          \path (I-\name) edge (K-\name);
+          \path (J-\name) edge (K-\name); }
+
+          \foreach \name / \x in {1,...,4}{
+              \path[xshift=3cm] node[neuron] (O-\name) at (0.3*\x,-5.25cm) {};
+              \foreach \j in {1,...,15}{
+             \path (K-\j) edge (O-\name);}
+             }
+  \end{tikzpicture}
+
+\end{frame}
+
+\defverbatim[colored]\lstI{
+\begin{lstlisting}[language=Python,basicstyle=\ttfamily,keywordstyle=\color{red}]
+SEQ_LENGTH = 256
+E_DIM = 128
+STATE_DIM = 512
+NUM_CLASSES = 4
+def inference():
+    model_input = tf.placeholder('uint8', shape=[None, SEQ_LENGTH])
+    _ = tf.one_hot(Globals.model_input, depth=E_DIM, axis=-1)
+    _ = tf.reshape(_, [-1, SEQ_LENGTH, E_DIM])
+    fw = multi_layer_rnn(N_LAYERS, STATE_DIM)
+    bw = multi_layer_rnn(N_LAYERS, STATE_DIM)
+    output, _ = tf.nn.bidirectional_dynamic_rnn(fw, bw, _, dtype=tf.float32)
+    fw_output = tf.reshape(output[0][:, -1:], [-1, STATE_DIM])
+    bw_output = tf.reshape(output[1][:, :1], [-1, STATE_DIM])
+    f = project(fw_output, E_DIM)
+    b = project(bw_output, E_DIM)
+    e = tf.add(f, b)
+    Globals.model_output = project(e, NUM_CLASSES)
+    Globals.prediction = tf.cast(tf.argmax(Globals.model_output, 1), tf.uint8)
+    return Globals.model_input, Globals.model_output
+\end{lstlisting}
+}
+
+\begin{frame}{A many-to-one example (code)}
+\lstI
+\end{frame}
+
+\begin{frame}{A many-to-many example}
+\end{frame}
+
+\begin{frame}{A many-to-many example (model architecture)}
+  \begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
+      \tikzstyle{arrow} = [thick,->,>=stealth]
+      \tikzstyle{startstop} = [rectangle, minimum width=1.5cm, minimum height=.6cm, text centered, draw=black, fill=black!20]
+      \tikzstyle{neuron}=[circle,fill=black!50,minimum size=6pt,inner sep=0pt]
+
+      \node (input) [] at (0cm,1cm) {};
+      \node (output) [] at (0cm,-3cm) {};
+      \node (s-input-A) [] at (-2cm,0cm) {};
+      \node (s-output-A) [] at (10cm,0cm) {};
+      \node (s-input-B) [] at (-2cm,-1cm) {};
+      \node (s-output-B) [] at (10cm,-1cm) {};
+      \node (s-input-C) [] at (-2cm,-2cm) {};
+      \node (s-output-C) [] at (10cm,-2cm) {};
+      \path[] node[startstop] (A-0) at (0cm,0cm) {};
+      \path[] node[startstop] (B-0) at (0cm,-1cm) {};
+      \path[] node[startstop] (C-0) at (0cm,-2cm) {};
+      \draw [arrow] (input) -- (A-0);
+      \draw [arrow] (s-input-A) -- (A-0);
+      \draw [arrow] (s-input-B) -- (B-0);
+      \draw [arrow] (s-input-C) -- (C-0);
+      \draw [arrow] (A-0) -- (B-0);
+      \draw [arrow] (B-0) -- (C-0);
+      \draw [arrow] (C-0) -- (output);
+
+       \foreach \i/\j in {1/0,2/1,3/2,4/3}
+       {
+         \node (input-\i) [] at (2*\i cm,1cm) {};
+         \node (output-\i) [] at (2*\i cm,-2.25cm) {};
+         \path[] node[startstop] (A-\i) at (2*\i cm,0cm) {};
+         \path[] node[startstop] (B-\i) at (2*\i cm,-1cm) {};
+         \path[] node[startstop] (C-\i) at (2*\i cm,-2cm) {};
+         \draw [arrow] (A-\i) -- (B-\i);
+         \draw [arrow] (B-\i) -- (C-\i);
+         \draw [arrow] (input-\i) -- (A-\i);
+         \draw [arrow] (A-\j) -- (A-\i);
+         \draw [arrow] (B-\j) -- (B-\i);
+         \draw [arrow] (C-\j) -- (C-\i);
+         \draw [arrow] (C-\i) -- (output-\i);
+       }
+       \draw [arrow] (A-4) -- (s-output-A);
+       \draw [arrow] (B-4) -- (s-output-B);
+       \draw [arrow] (C-4) -- (s-output-C);
+
+       \foreach \name / \x in {1,...,15}{
+           \path[xshift=0cm]
+               node[neuron] (K-\name) at (0.5*\x,-3.5cm) {};
+          \path (output-2) edge (K-\name);
+          %\path (J-\name) edge (K-\name);
+          }
+
+          \foreach \name / \x in {1,...,10}{
+              \path[xshift=1cm] node[neuron] (O-\name) at (0.6*\x,-5.75cm) {};
+              \foreach \j in {1,...,15}{
+             \path (K-\j) edge (O-\name);}
+             %\path (J-\name) edge (K-\name);
+             }
+  \end{tikzpicture}
+\end{frame}
+
+\defverbatim[colored]\lstI{
+\begin{lstlisting}[language=Python,basicstyle=\ttfamily,keywordstyle=\color{red}]
+SEQ_LENGTH = 256
+E_DIM = 128
+STATE_DIM = 512
+N_LAYERS = 3
+
+def inference():
+    model_input = tf.placeholder('uint8', shape=[None, SEQ_LENGTH])
+    _ = tf.one_hot(Globals.model_input, depth=E_DIM, axis=-1)
+    encode = multi_layer_rnn(N_LAYERS, STATE_DIM)
+    state_tuple = tuple(tf.unstack(Globals.initial_state, axis=0))
+    output, state = tf.nn.dynamic_rnn(encode, _,
+                                      dtype=tf.float32,
+                                      initial_state=state_tuple)
+    output = tf.reshape(output, [-1, STATE_DIM])
+    output = project(output, E_DIM)
+    out = tf.cast(tf.argmax(output, 1), tf.uint8)
+    out = tf.reshape(out, [-1, SEQ_LENGTH])
+    Globals.generated_sequence = out
+    Globals.generated_characters = tf.nn.softmax(output)
+    Globals.model_output = output
+    Globals.state = state
+\end{lstlisting}
+}
+
+\begin{frame}{A many-to-one example (code)}
+\lstI
+\end{frame}
+
+\defverbatim[colored]\lstI{
+\begin{lstlisting}[language=Python,basicstyle=\ttfamily,keywordstyle=\color{red}]
+def generate_text(length, session=None):
+    generated_text = ''
+    character = [[ord(' ')]]
+    istate = np.zeros([N_LAYERS, 1, STATE_DIM])
+    while len(generated_text) < length:
+        feed_dict = {Globals.model_input: character,
+                     Globals.initial_state: istate}
+        next_char, state = session.run([Globals.generated_characters,
+                                        Globals.state],
+                                       feed_dict=feed_dict)
+        next_char = np.asarray(next_char).astype('float64')
+        next_char = next_char / next_char.sum()
+        op = np.random.multinomial
+        next_char_id = op(1, next_char.squeeze(), 1).argmax()
+        next_char_id = next_char_id if chr(next_char_id) in \
+                          string.printable else ord(" ")
+        generated_text += chr(next_char_id)
+        character = [[next_char_id]]
+        istate = state
+    return generated_text
+\end{lstlisting}
+}
+
+\begin{frame}{A many-to-many example (text generating)}
+\lstI
+\end{frame}
+
+ \begin{frame}{Many-to-one-to-many example}
+
+\begin{itemize}
+\item  A recurrent neural network that can sort sequences
+\item Two parts: an encoder, and a decoder
+\item The encoder encodes sequences into fixed length vectors
+\item The decoder transforms this vector into a sorted list of numbers.
+\end{itemize}
+
+ \end{frame}
+
+ \begin{frame}{Many-to-one-to-many example}
+
    \begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
@@
        \path[] node[startstop] (C-0) at (0cm,-2cm) {};
-    \draw [arrow] (input) -- (A);
+       \draw [arrow] (input) -- (A-0);
        \draw [arrow] (s-input-A) -- (A-0);
@@
        \draw [arrow] (s-input-C) -- (C-0);
-    \draw [arrow] (A) -- (B);
-    \draw [arrow] (B) -- (C);
-    \draw [arrow] (C) -- (output);
+       \draw [arrow] (A-0) -- (B-0);
+       \draw [arrow] (B-0) -- (C-0);
+       \draw [arrow] (C-0) -- (output);
 
@@
    \end{tikzpicture}
+ \end{frame}
 
-\begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
-    \tikzstyle{every pin edge}=[<-,shorten <=1pt]
-    \tikzstyle{neuron}=[circle,fill=black!50,minimum size=10pt,inner sep=0pt]
-
-    % Draw the hidden layer nodes
-    \foreach \name / \x in {1,...,5}
-        \path[yshift=0.5cm]
-            node[neuron] (I-\name) at (\x,-0cm) {};
-
-    \foreach \name / \x in {1,...,5}
-        \path[yshift=0.5cm]
-            node[neuron] (H-\name) at (\x,-2cm) {};
-
-    \foreach \source in {1,...,5}
-        \foreach \dest in {1,...,5}
-            \path (I-\source) edge (H-\dest);
-
-    \foreach \name / \x in {1,...,5}
-        \path[yshift=0.5cm]
-            node[neuron] (K-\name) at (\x,-4 cm) {};
-
-    \foreach \source in {1,...,5}
-        \foreach \dest in {1,...,5}
-            \path (H-\source) edge (K-\dest);
-\end{tikzpicture}
-
-The base input ofor this network is a (unicode) character, which is one-hot encoded before being fed into the
- recurrent layers.  For training, the network is unrolled to a length of 256 characters,
- which is about twice the length  of an average message in our database, and all strings are padded or truncated
- to a length of 256 characters.
-
-\begin{minted}{python}
-  SEQ_LENGTH = 256
-  E_DIM = 128
-  STATE_DIM = 512
-  NUM_CLASSES = 4
-  def inference():
-      model_input = tf.placeholder('uint8', shape=[None, SEQ_LENGTH])
-      _ = tf.one_hot(Globals.model_input, depth=E_DIM, axis=-1)
-      _ = tf.reshape(_, [-1, SEQ_LENGTH, E_DIM])
-      fw = multi_layer_rnn(N_LAYERS, STATE_DIM)
-      bw = multi_layer_rnn(N_LAYERS, STATE_DIM)
-      output, _ = tf.nn.bidirectional_dynamic_rnn(fw, bw, _, dtype=tf.float32)
-      fw_output = tf.reshape(output[0][:, -1:], [-1, STATE_DIM])
-      bw_output = tf.reshape(output[1][:, :1], [-1, STATE_DIM])
-      f = project(fw_output, E_DIM)
-      b = project(bw_output, E_DIM)
-      e = tf.add(f, b)
-      Globals.model_output = project(e, NUM_CLASSES)
-      Globals.prediction = tf.cast(tf.argmax(Globals.model_output, 1), tf.uint8)
-      return Globals.model_input, Globals.model_output
-\end{minted}
-
-%first hour:
-
-%Loss: 1.331 Accuracy: 45.50%
-%0	Negative	Neutral	Positive	Irrelevant
-%Negative	291	113	50	127
-%Neutral	108	164	62	85
-%Positive	0	0	0	0
-%Irrelevant	0	0	0	0
-\begin{tabular}{lcccc}
-  {} & {\bf Negative} & {\bf Neutral} & {\bf Positive} & {\bf Irrelevant}\\
-  {\bf Negative}   & {291} & {113} & {50} & {127}\\
-  {\bf Neutral}    & {108} & {113} & {50} & {85}\\
-  {\bf Positive}   & {0} & {0} & {0} & {0}\\
-  {\bf Irrelevant} & {0} & {0} & {0} & {0}\\
-\end{tabular}
-
-%Loss: 0.933 Accuracy: 60.30%
-%0	Negative	Neutral	Positive	Irrelevant
-%Negative	292	52	36	55
-%Neutral	61	176	21	62
-%Positive	5	15	65	15
-%Irrelevant	33	28	14	70
-\begin{tabular}{lcccc}
-  {} & {\bf Negative} & {\bf Neutral} & {\bf Positive} & {\bf Irrelevant}\\
-  {\bf Negative}   & {292} & {52} & {36} & {55}\\
-  {\bf Neutral}    & {61} & {176} & {21} & {62}\\
-  {\bf Positive}   & {5} & {15} & {65} & {15}\\
-  {\bf Irrelevant} & {33} & {28} & {14} & {70}\\
-\end{tabular}
-
-% second hour
-% Loss: 0.729 Accuracy: 71.80%
-% 0	Negative	Neutral	Positive	Irrelevant
-% Negative	340	67	18	65
-% Neutral	22	182	18	36
-% Positive	10	7	91	10
-% Irrelevant	8	13	8	105
-% \begin{tabular}{lcccc}
-%   {} & {\bf Negative} & {\bf Neutral} & {\bf Positive} & {\bf Irrelevant}\\
-%   {\bf Negative}   & {} & {} & {} & {}\\
-%   {\bf Neutral}    & {} & {} & {} & {}\\
-%   {\bf Positive}   & {} & {} & {} & {}\\
-%   {\bf Irrelevant} & {} & {} & {} & {}\\
-% \end{tabular}
-
-%third hour
-%Loss: 0.611 Accuracy: 76.70%
-%0	Negative	Neutral	Positive	Irrelevant
-%Negative	319	38	11	43
-%Neutral	20	188	2	20
-%Positive	10	18	91	10
-%Irrelevant	27	27	7	169
-\begin{tabular}{lcccc}
-  {} & {\bf Negative} & {\bf Neutral} & {\bf Positive} & {\bf Irrelevant}\\
-  {\bf Negative}   & {319} & {38} & {11} & {43}\\
-  {\bf Neutral}    & {20} & {188} & {2} & {20}\\
-  {\bf Positive}   & {10} & {18} & {91} & {10}\\
-  {\bf Irrelevant} & {27} & {27} & {7} & {169}\\
-\end{tabular}
-
-% Loss: 0.468 Accuracy: 85.20%
-% 0	Negative	Neutral	Positive	Irrelevant
-% Negative	334	20	3	18
-% Neutral	14	242	0	9
-% Positive	20	10	125	17
-% Irrelevant	26	10	1	151
-% \begin{tabular}{lcccc}
-%   {} & {\bf Negative} & {\bf Neutral} & {\bf Positive} & {\bf Irrelevant}\\
-%   {\bf Negative}   & {} & {} & {} & {}\\
-%   {\bf Neutral}    & {} & {} & {} & {}\\
-%   {\bf Positive}   & {} & {} & {} & {}\\
-%   {\bf Irrelevant} & {} & {} & {} & {}\\
-% \end{tabular}
-%
-
-%0	Negative	Neutral	Positive	Irrelevant
-%Negative	334	4	4	9
-%Neutral	3	283	2	7
-%Positive	1	0	125	2
-%Irrelevant	1	0	4	221
-\begin{tabular}{lcccc}
-  {} & {\bf Negative} & {\bf Neutral} & {\bf Positive} & {\bf Irrelevant}\\
-  {\bf Negative}   & {334} & {4} & {4} & {9}\\
-  {\bf Neutral}    & {3} & {283} & {2} & {7}\\
-  {\bf Positive}   & {1} & {0} & {125} & {2}\\
-  {\bf Irrelevant} & {1} & {0} & {4} & {221}\\
-\end{tabular}
-
-% Loss: 0.530 Accuracy: 84.60%
-% 0	Negative	Neutral	Positive	Irrelevant
-% Negative	361	18	13	29
-% Neutral	22	246	10	18
-% Positive	6	6	95	6
-% Irrelevant	11	5	10	144
-% \begin{tabular}{lcccc}
-%   {} & {\bf Negative} & {\bf Neutral} & {\bf Positive} & {\bf Irrelevant}\\
-%   {\bf Negative}   & {} & {} & {} & {}\\
-%   {\bf Neutral}    & {} & {} & {} & {}\\
-%   {\bf Positive}   & {} & {} & {} & {}\\
-%   {\bf Irrelevant} & {} & {} & {} & {}\\
-% \end{tabular}
-
-%Loss: 0.275 Accuracy: 90.00%
-%0	Negative	Neutral	Positive	Irrelevant
-%Negative	321	16	3	20
-%Neutral	8	286	4	6
-%Positive	4	2	125	16
-%Irrelevant	11	9	1	168
-\begin{tabular}{lcccc}
-  {} & {\bf Negative} & {\bf Neutral} & {\bf Positive} & {\bf Irrelevant}\\
-  {\bf Negative}   & {321} & {16} & {3} & {20}\\
-  {\bf Neutral}    & {8} & {286} & {4} & {6}\\
-  {\bf Positive}   & {4} & {2} & {125} & {16}\\
-  {\bf Irrelevant} & {11} & {9} & {1} & {168}\\
-\end{tabular}
-
-\subsection{Many-to-many}
-
-\begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
-    \tikzstyle{arrow} = [thick,->,>=stealth]
-    \tikzstyle{startstop} = [rectangle, minimum width=1.5cm, minimum height=.6cm, text centered, draw=black, fill=black!20]
-    \node (input) [] at (0cm,1cm) {};
-    \node (output) [] at (0cm,-3cm) {};
-    \node (s-input-A) [] at (-2cm,0cm) {};
-    \node (s-output-A) [] at (10cm,0cm) {};
-    \node (s-input-B) [] at (-2cm,-1cm) {};
-    \node (s-output-B) [] at (10cm,-1cm) {};
-    \node (s-input-C) [] at (-2cm,-2cm) {};
-    \node (s-output-C) [] at (10cm,-2cm) {};
-    \path[] node[startstop] (A-0) at (0cm,0cm) {};
-    \path[] node[startstop] (B-0) at (0cm,-1cm) {};
-    \path[] node[startstop] (C-0) at (0cm,-2cm) {};
-    \draw [arrow] (input) -- (A);
-    \draw [arrow] (s-input-A) -- (A-0);
-    \draw [arrow] (s-input-B) -- (B-0);
-    \draw [arrow] (s-input-C) -- (C-0);
-    \draw [arrow] (A) -- (B);
-    \draw [arrow] (B) -- (C);
-    \draw [arrow] (C) -- (output);
-
-     \foreach \i/\j in {1/0,2/1,3/2,4/3}
-     {
-       \node (input-\i) [] at (2*\i cm,1cm) {};
-       \node (output-\i) [] at (2*\i cm,-3cm) {};
-       \path[] node[startstop] (A-\i) at (2*\i cm,0cm) {};
-       \path[] node[startstop] (B-\i) at (2*\i cm,-1cm) {};
-       \path[] node[startstop] (C-\i) at (2*\i cm,-2cm) {};
-       \draw [arrow] (A-\i) -- (B-\i);
-       \draw [arrow] (B-\i) -- (C-\i);
-       \draw [arrow] (input-\i) -- (A-\i);
-       \draw [arrow] (A-\j) -- (A-\i);
-       \draw [arrow] (B-\j) -- (B-\i);
-       \draw [arrow] (C-\j) -- (C-\i);
-       \draw [arrow] (C-\i) -- (output-\i);
-     }
-     \draw [arrow] (A-4) -- (s-output-A);
-     \draw [arrow] (B-4) -- (s-output-B);
-     \draw [arrow] (C-4) -- (s-output-C);
-
-\end{tikzpicture}
-
-\begin{minted}{python}
-  SEQ_LENGTH = 256
-  E_DIM = 128
-  STATE_DIM = 512
-  N_LAYERS = 3
-
-  def inference():
-      model_input = tf.placeholder('uint8', shape=[None, SEQ_LENGTH])
-      _ = tf.one_hot(Globals.model_input, depth=E_DIM, axis=-1)
-      encode = multi_layer_rnn(N_LAYERS, STATE_DIM)
-      state_tuple = tuple(tf.unstack(Globals.initial_state, axis=0))
-      output, state = tf.nn.dynamic_rnn(encode, _,
-                                        dtype=tf.float32,
-                                        initial_state=state_tuple)
-      output = tf.reshape(output, [-1, STATE_DIM])
-      output = project(output, E_DIM)
-      out = tf.cast(tf.argmax(output, 1), tf.uint8)
-      out = tf.reshape(out, [-1, SEQ_LENGTH])
-      Globals.generated_sequence = out
-      Globals.generated_characters = tf.nn.softmax(output)
-      Globals.model_output = output
-      Globals.state = state
-\end{minted}
-
-\begin{minted}{python}
-def generate_text(length, session=None):
-    generated_text = ''
-    character = [[ord(' ')]]
-    istate = np.zeros([N_LAYERS, 1, STATE_DIM])
-    while len(generated_text) < length:
-        feed_dict = {Globals.model_input: character,
-                     Globals.initial_state: istate}
-        next_char, state = session.run([Globals.generated_characters,
-                                        Globals.state],
-                                       feed_dict=feed_dict)
-        next_char = np.asarray(next_char).astype('float64')
-        next_char = next_char / next_char.sum()
-        op = np.random.multinomial
-        next_char_id = op(1, next_char.squeeze(), 1).argmax()
-        next_char_id = next_char_id if chr(next_char_id) in \
-                          string.printable else ord(" ")
-        generated_text += chr(next_char_id)
-        character = [[next_char_id]]
-        istate = state
-    return generated_text
-\end{minted}
-
-\subsection{Many-to-one-to-many}
-
-As an example of a recurrent neural network performing a task that ordinary neural network
-should not be able to perform, we present a sorting function which uses GRU cells. The architecture
-of the network is the reason we call it a many-to-one-to-many network.  One level takes a list
-of numbers, and produces a vector which we can see as representing all the numbers in the list,
-and a second layer uses this output vector as input, and produces the original list, sorted in
-increasing order. The training was done using sequences of natural numbers $n\in\{1,...,32\}$, and each
-of the training sequences of length $32$.  After about 6 days of training, the accuracy of the network
-stabilized at around 95\%, and it is worth noting that eventhough it took a long time to breach
-the 90\% accuracy mark, very early on in the training did the network output increasing sequences.
-
-\begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
-    \tikzstyle{arrow} = [thick,->,>=stealth]
-    \tikzstyle{startstop} = [rectangle, minimum width=1.5cm, minimum height=.6cm, text centered, draw=black, fill=black!20]
-    \node (input) [] at (0cm,1cm) {};
-    \node (output) [] at (0cm,-3cm) {};
-    \node (s-input-A) [] at (-2cm,0cm) {};
-    \node (s-output-A) [] at (10cm,0cm) {};
-    \node (s-input-B) [] at (-2cm,-1cm) {};
-    \node (s-output-B) [] at (10cm,-1cm) {};
-    \node (s-input-C) [] at (-2cm,-2cm) {};
-    \node (s-output-C) [] at (10cm,-2cm) {};
-    \path[] node[startstop] (A-0) at (0cm,0cm) {};
-    \path[] node[startstop] (B-0) at (0cm,-1cm) {};
-    \path[] node[startstop] (C-0) at (0cm,-2cm) {};
-    \draw [arrow] (input) -- (A);
-    \draw [arrow] (s-input-A) -- (A-0);
-    \draw [arrow] (s-input-B) -- (B-0);
-    \draw [arrow] (s-input-C) -- (C-0);
-    \draw [arrow] (A) -- (B);
-    \draw [arrow] (B) -- (C);
-    \draw [arrow] (C) -- (output);
-
-     \foreach \i/\j in {1/0,2/1,3/2,4/3}
-     {
-       \node (input-\i) [] at (2*\i cm,1cm) {};
-       \node (output-\i) [] at (2*\i cm,-3cm) {};
-       \path[] node[startstop] (A-\i) at (2*\i cm,0cm) {};
-       \path[] node[startstop] (B-\i) at (2*\i cm,-1cm) {};
-       \path[] node[startstop] (C-\i) at (2*\i cm,-2cm) {};
-       \draw [arrow] (A-\i) -- (B-\i);
-       \draw [arrow] (B-\i) -- (C-\i);
-       \draw [arrow] (input-\i) -- (A-\i);
-       \draw [arrow] (A-\j) -- (A-\i);
-       \draw [arrow] (B-\j) -- (B-\i);
-       \draw [arrow] (C-\j) -- (C-\i);
-       \draw [arrow] (C-\i) -- (output-\i);
-     }
-     \draw [arrow] (A-4) -- (s-output-A);
-     \draw [arrow] (B-4) -- (s-output-B);
-     \draw [arrow] (C-4) -- (s-output-C);
-
-\end{tikzpicture}
-
-\begin{minted}{python}
+ \defverbatim[colored]\lstI{
+ \begin{lstlisting}[language=Python,basicstyle=\ttfamily,keywordstyle=\color{red}]
    SEQ_LENGTH = 256
@@
    def inference():
-      model_input = tf.placeholder('uint8', shape=[None, SEQ_LENGTH])
-      _ = tf.one_hot(Globals.model_input, depth=E_DIM, axis=-1)
+       model_input = tf.placeholder('uint8',
+                                    shape=[None, SEQ_LENGTH])
+       _ = tf.one_hot(model_input, depth=E_DIM, axis=-1)
        _ = tf.reshape(_, [-1, SEQ_LENGTH, E_DIM])
        encode = multi_layer_rnn(N_LAYERS, STATE_DIM)
-      encoded_input, state = tf.nn.dynamic_rnn(encode,
-                                               _,
-                                               dtype=tf.float32)
+       OP = tf.nn.dynamic_rnn
+       encoded_input, state = OP(encode,_,dtype=tf.float32)
        Globals.encoder_output = state
@@
            Globals.decoder_input = _
-\end{minted}
+ \end{lstlisting}
+ }
+ \begin{frame}{Many-to-one-to-many example code}
+   \lstI
+ \end{frame}
 

### 2017-08-28 deeplearn/twitter_sentiment/models/text_generator/text_generator.py

@@ -201,6 +201,12 @@ def generate_text(length, session=None):
         istate = state
     return generated_text
 
+def initialize_uninitialized_variables(session=None):
+    global_variables = tf.global_variables()
+    is_initialized = session.run([tf.is_variable_initialized(x) for x in global_variables])
+    uninitialized_variables = [x for (x, v) in zip(global_variables, is_initialized) if not v]
+    if len(uninitialized_variables) > 0:
+        session.run(tf.variables_initializer(uninitialized_variables))
 
 if __name__ == '__main__':
     root = os.path.dirname(__file__)
@@ -215,12 +221,14 @@ if __name__ == '__main__':
     index = 0
     istate = np.zeros([Hyperparameters.n_layers, 32, Hyperparameters.hidden_states])
     with tf.Session() as session:
-        session.run(tf.global_variables_initializer())
+        #session.run(tf.global_variables_initializer())
         if not os.path.exists(chkpt):
             os.makedirs(chkpt)
+            session.run(tf.global_variables_initializer())
         else:
             s = tf.train.Saver()
-            s.save(session, os.path.join(chkpt, "t_gen_weights"))
+            s.restore(session, os.path.join(chkpt, "t_gen_weights"))
+            initialize_uninitialized_variables(session)
             print('Model restored')
 
         for batch in rnn_minibatch_sequencer(data, 32, Hyperparameters.sequence_length, nb_epochs=1000):
@@ -233,12 +241,12 @@ if __name__ == '__main__':
             print('x', end='', flush=True)
             istate = ostate
             #print(index, batch['epoch_number'])
-            if index % (batch['epoch_number']*150) == 0:
+            if index % 300 == 0:
                 file_name = "text_gen-{}".format(index)
                 file_name = os.path.join(save, file_name)
                 with open(file_name, 'w') as test_file:
                     test_file.write(generate_text(4000, session))
-            if index % 250 == 0:
-                s = tf.train.Saver()
-                s.save(session, os.path.join(chkpt, "t_gen_weights"))
+            #if index % 250 == 0:
+            #    s = tf.train.Saver()
+            #    s.save(session, os.path.join(chkpt, "t_gen_weights"))
             index += 1

### 2017-08-28 deeplearn/presentation/presentation_slides.tex
presentation update

@@
 
-\usepackage{minted}
+%\usepackage{minted}
 
@@
 
+\begin{frame}{A Bird's Eye View Of Classical Neural Networks}
+Let $f:\R^n\to \R^m$ be a function, and consider a class $\C$ of functions $g: \R^n\to \R^m$, which we assume to be parametrized by $\R^\ell$, in the sense that there is a function $F(\x,\w):\R^n\times \R^\ell\to\R^n$ such that $\C=\{\x\mapsto F(\x,\w):\w\in\R^n\}$.
+The generic problem of machine learning is to find a vector $\w\in\R^{\ell}$ such that $\ell(\w) = D(f(\x), F(\x, \w))$ is as small as possible, where $D$ represents a notion of distance between functions.
+ Given the definition of $D$, we see that $\ell(\x)\geq 0$, and therefore $\ell$ has a global minimum (which is not necessarily unique).
+In general, the global minimum of $\ell$ is not attained, in the sense that if $\epsilon=\min\{\ell(\w):\x\in\R^n\}$, then there does not necessarily exist a value $\w_0\in\R^\ell$ such that $\epsilon=\ell(\w_0)$.
+However, it can be approximated, that is to say there is a sequence $\w_i$ such that $\ell(\w_i)\to\epsilon$ as $i\to\infty$.
+\end{frame}
+
+\begin{frame}{A Bird's Eye View Of Classical Neural Networks}
+In tensorflow, the function we are trying to approximate is given in the form of a relation $f(\x)=\y$.
+We provide {\em placeholders} for the values of $\x$ and $\y$.
+The function $F(\x,\w)$ is defined by a computation graph.
+For example, here is a linear regression, in which we try to approximate a function $f:\R^{10}\to\R^{15}$ using a function of the form $F(\x, \W\b)=\W\x+\b$.
+Here $\W$ is an $15\times 10$ matrix, and $\b\in\R^{15}$
+\end{frame}
+
+\defverbatim[colored]\lstI{
+\begin{lstlisting}[language=Python,basicstyle=\ttfamily,keywordstyle=\color{red}]
+x = tf.placeholder(shape=[None, 10])
+y = tf.placeholder(shape=[None, 15])
+W = tf.Variable(shape=[None, 10, 15])
+b = tf.Variable(shape=[None, 15])
+F = tf.matmul(x, W) + b
+loss = tf.mean_square_error(y, F)
+\end{lstlisting}
+}
+\begin{frame}{A Bird's Eye View Of Classical Neural Networks}
+\lstI
+\end{frame}
+
+\begin{frame}{Sequences}
+Classical neural networks are great for classification problems involving either fixed (finite) sets $f:A\to B$, where $A$ represents the set to be classified, and $B$ is the set of labels, or functions $f:\R^n\to B$.
+An example of the latter is image classification, since an $n\times m$ image can be directly represented as a vector in $\R^{\ell mn}$.
+In many cases, however, the data to be classified is better represented as a (finite) sequence.
+For example, sentences are sequences of words, words are sequences of letters, movies are sequences of images, and sound files are sequences of amplitudes.
+It is possible to deal with sequences with ordinary neural networks, but we run into difficulties when we try to model dependency between different elements of a sequence.
+\end{frame}
+
+\begin{frame}{Sequences}
+Let $A$ be a finite set.
+A finite sequence of elements of $A$ is a tuple $(a_1,...,a_n)$, where $a_i\in A$ for every $i$.
+Here the number $n$ is allowed to change.
+The set of all finite sequences of elements of $A$ is denoted $A^*$.
+From the definition we just gave, we have $A^*=\bigcup_{n\geq 0} A^n$, which is a good enough definition for our purpose.
+If we are given two finite sets $A$ and $B$, there are many situations we can consider if we also have access to $A^*$ and $B^*$.
+\end{frame}
+
+\begin{frame}{Sequences}
+  \begin{enumerate}
+    \item A function $f:A\to B^*$, which given an element of $A$ outputs a sequence of elements of $B$.
+This is called a {\em one-to-many} function.
+We use this architecture in a sequence sorting neural network.
+The Tweet2Vec autoencoder also uses a one-to-many neural network.
+    \item A function $f:A^*\to B$, which is given a sequence of elments of $A$ and produces a single element of $B$.
+This is called a {\em many-to-one} function.
+A good example of sich a function is the Buzzometer sentiment analysis tool which assigns a single sentiment value to messages which can vary in length.
+    \item A function $f:A^*\to B^*$ which is given a seuqence of elements of $A$< and outputs a sequence of elements of $B$.
+This is called a {\em many-to-many} function.
+As an example of such a function we will show the implementation of a simple text generator.
+Translation, video captioning, part-of-speech tagging and speech recognition are all examples of many-to-many functions.
+  \end{enumerate}
+\end{frame}
+
+\begin{frame}{Functions on Sequences}
+We can of course define functions $A\to B^*$, $A^*\to B$ and $A^*\to B^*$ directly (they are just sets, after all), but it is more interesting and instructive to make use of the structure of $A^*$ and $B^*$ as sets of sequences of $A$ and $B$, and see how one can use a function (or class of functions) $A\to B$ to define a function $A^*\to B^*$.
+In particular, we will me making use of the iterative nature of the construction $A^*$ and $B^*$.
+Here is the simplest example: if $A=B$, and $f:A\to A$ is a function, then we can iterate $f$: $f^0(a)=a$ for every $a\in A$, and $f^{n+1}(a) = f(f^n(a))$.
+We terminate the iteration at the first value of $n$ for which $f^n(a)=f^{n-1}(a)$ (note that there is no guarantee that this will ever happen).
+This gives a function $f:A\to A^\omega$.
+\end{frame}
+
+\begin{frame}{Functions on Sequences}
+We describe a more general situation.
+Let $S$ be a (finite) set of {\em states} with a distinguished element $\perp\in S$, and consider a function $f:A\times S\to B\times S$.
+Let $a_1,...,a_n$ be a finite sequence of elements of $A$.
+We define two sequences $b_1,...,b_n$ and $s_1,...,s_n$ of elements of $B$ and $S$ respectively as follows.
+We first define $b_1, s_1 = f(a_1, \perp)$.
+Suppose that we have defined the elements $b_1,...,b_n$ and $s_1,...,s_n$.
+We define $b_{n+1}$ and $s_{n+1}$ via $b_{n+1}, s_{n+1} = f(a_{n+1}, s_n)$.
+We define $f^*(a_1...a_n)=b_1...b_n$, with $b_1,...,b_n$ defined as above.
+This is a function $f^*:A^*\to B^*$, and it has the property that the length of the output sequence is the same as that of the input sequence.
+An interesting special case of this is when $S=B$.
+In this case, we can have $f:A\times B\to B$.
+We can define $b_1 = f(a_1, \perp)$, and $b_{n+1}=f(a_{n+1}, b_n)$.
+This is in fact the main type of function we use in our examples later.
+Note that this is a state machine, close to a deterministic finite automaton, except that here the state space $S$ is allowed to be infinite.
+\end{frame}
+
+\begin{frame}{Functions on Sequences}
+We use the exact same method to define a many-to-one function.
+If $f:A\times S\to B\times S$ is a function, and  $a_1,...,a_n$, $b_1,...,b_n$ and $s_1,...,s_n$ are defined as above, then we define $f^*(a_1,...,a_n)=b_n$.
+In other words, we discard all elements of $B$ produced during the iteration, and keep only the last one.
+\end{frame}
+
+\begin{frame}{Functions on Sequences}
+Finally, we tackle the case of a one-to-many function.
+Again we let $f:A\times S\to B\times S$ be any function.
+We define sequences $b_1,...,b_i,...$ and $s_1,...,s_i,...$ as follows.
+$b_1,s_1=f(a,\perp)$, and for every $n$, $b_{n+1}, s_{n+1}=f(a, s_n)$.
+In other words, we keep iterating $f$, feeding $a$ as input at every step.
+An imoprtant variant of this is when $a$ itself has a distinguished element $\perp_A$, in which case, after feeding $a$ as an input to $f$ in the first iteration, we subsequently feed it $\perp_A$, and get $b_{n+1}, s_{n+1}=f(\perp_A, s_n)$.
+We stop the iteration whenever $s_n=\perp$ (or any other predetermined element of $S$).
+Note that this may never happen, i.e.
+the sequence $b_n$ may go on forever.
+In fact, there is no way to give a general pattern of definition for $f:A\times S\to B\times S$ which will guarantee that $f^*(a)$ is finite for every $a$.
+\end{frame}
+
+\begin{frame}{Recurrent Neural Networks}
+  In general, a recurrent neural network is just a presceiption for a class of functions of the form $f:\R^n\times \R^\ell\times\R^m\to \R^m\times \R^\ell$, where $\R^m$ is the space of parameters.
+The set $\{0,1\}^*$ of all finite sequences of $0$'s and $1$'s is countable, and $\R^\ell$, being a real vector space,
+  is uncountable.  Therefore, we can encode every element $w\in\{0,1\}^*$ as a vector in $S$.  Informally,
+  $\R^\ell$ is enough to encode any finite state space, and every possible value for the content of the tape of a
+  Turing machine. We get:
+
+  \begin{theorem}
+  Recurrent neural networks are Turing-complete.
+  \end{theorem}
+
+  This explains why recurrent neural networks seem to be able to produce results that other networks can't.
+Training a recurrent network is the same as producing a Turing machine.
+A common way to represent a recurrent node is to use a box, the inside of which represent the definition of $f$.
+The vertical arrows represent the input and onput of the node, and the horisontal arrows represent the input and output state of the node.
+\end{frame}
+
+\begin{frame}{Recurrent Neural Networks}
+  \begin{center}
+    \begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
+      \tikzstyle{arrow} = [thick,->,>=stealth]
+      \tikzstyle{startstop} = [rectangle, minimum width=2cm, minimum height=.75cm, text centered, draw=black, fill=black!20]
+      \node (input) [] at (0cm,1cm) {};
+      \node (output) [] at (0cm,-1cm) {};
+      \node (s-input) [] at (-2cm,0cm) {};
+      \node (s-output) [] at (2cm,0cm) {};
+      \path[] node[startstop] (N) at (0cm,0cm) {};
+      \draw [arrow] (input) -- (N);
+      \draw [arrow] (s-input) -- (N);
+      \draw [arrow] (N) -- (output);
+      \draw [arrow] (N) -- (s-output);
+  \end{tikzpicture}
+  \end{center}
+\end{frame}
+
+\begin{frame}{Recurrent Neural Networks}
+  ust like any regular neural network layer, recurrent nodes can be composed.
+The composition of two recurrent nodes is done by feeding the output of one node into the input of the other, and concatenating their state space.
+Formally, if $f:\R^n\times \R^\ell\times\to \R^m\times \R^\ell$ and $g:\R^n\times \R^\ell\times\to \R^m\times \R^\ell$ are two recurrent nodes, then we have the node $(f;g):\R^m\times\R^\ell\times\R^L\to\R^k\times\R^\ell\times\R^L$ defined by $(f;g)(\x, \s_1, \s_2) = (\y, \s'_1, \s'_2)$ where $f(\x, \s_1) = \x', \s'_1$ and $g(\x', \s_2) = \y, \s'_2$.
+Graphically, we can represent this by stacking the boxes representing $f$ and $g$ on top of one another:
+
+  \begin{center}
+  \begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
+      \tikzstyle{arrow} = [thick,->,>=stealth]
+      \tikzstyle{startstop} = [rectangle, minimum width=1.5cm, minimum height=.6cm, text centered, draw=black, fill=black!20]
+      \node (input) [] at (0cm,1cm) {};
+      \node (output) [] at (0cm,-3cm) {};
+      \node (s-input-A) [] at (-2cm,0cm) {};
+      \node (s-output-A) [] at (2cm,0cm) {};
+      \node (s-input-B) [] at (-2cm,-1cm) {};
+      \node (s-output-B) [] at (2cm,-1cm) {};
+      \node (s-input-C) [] at (-2cm,-2cm) {};
+      \node (s-output-C) [] at (2cm,-2cm) {};
+      \path[] node[startstop] (A) at (0cm,0cm) {};
+      \path[] node[startstop] (B) at (0cm,-1cm) {};
+      \path[] node[startstop] (C) at (0cm,-2cm) {};
+      \draw [arrow] (input) -- (A);
+      \draw [arrow] (s-input-A) -- (A);
+      \draw [arrow] (s-input-B) -- (B);
+      \draw [arrow] (s-input-C) -- (C);
+      \draw [arrow] (A) -- (s-output-A);
+      \draw [arrow] (B) -- (s-output-B);
+      \draw [arrow] (C) -- (s-output-C);
+      \draw [arrow] (A) -- (B);
+      \draw [arrow] (B) -- (C);
+      \draw [arrow] (C) -- (output);
+  \end{tikzpicture}
+  \end{center}
+\end{frame}
+
 \begin{frame}{Training RNN's}
+  How do we train recurrent neural networks?
+we can use back propagation, just like a regular neural network.
+Heuristically, we unroll the recurrent network ``infinitely'' many times, until it looks like an ordinary very deep neural network.
+In practice, since computers only have a finite amount of resources, we only unroll the network a large but finite number of times, and treat it like an ordinary neural network.
+
+  \begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
+      \tikzstyle{arrow} = [thick,->,>=stealth]
+      \tikzstyle{startstop} = [rectangle, minimum width=1.5cm, minimum height=.6cm, text centered, draw=black, fill=black!20]
+      \node (input) [] at (0cm,1cm) {};
+      \node (output) [] at (0cm,-1cm) {};
+      \node (s-input-A) [] at (-2cm,0cm) {};
+      \node (s-output-A) [] at (10cm,0cm) {};
+      %\node (s-input-B) [] at (-2cm,-1cm) {};
+      %\node (s-output-B) [] at (10cm,-1cm) {};
+      %\node (s-input-C) [] at (-2cm,-2cm) {};
+      %\node (s-output-C) [] at (10cm,-2cm) {};
+      \path[] node[startstop] (A-0) at (0cm,0cm) {};
+      %\path[] node[startstop] (B-0) at (0cm,-1cm) {};
+      %\path[] node[startstop] (C-0) at (0cm,-2cm) {};
+      \draw [arrow] (input) -- (A);
+      \draw [arrow] (s-input-A) -- (A-0);
+      %\draw [arrow] (s-input-B) -- (B-0);
+      %\draw [arrow] (s-input-C) -- (C-0);
+      %\draw [arrow] (A) -- (B);
+      %\draw [arrow] (B) -- (C);
+      \draw [arrow] (A) -- (output);
+
+       \foreach \i/\j in {1/0,2/1,3/2,4/3}
+       {
+         \node (input-\i) [] at (2*\i cm,1cm) {};
+         \node (output-\i) [] at (2*\i cm,-1cm) {};
+         \path[] node[startstop] (A-\i) at (2*\i cm,0cm) {};
+      %   \path[] node[startstop] (B-\i) at (2*\i cm,-1cm) {};
+    %     \path[] node[startstop] (C-\i) at (2*\i cm,-2cm) {};
+    %     \draw [arrow] (A-\i) -- (B-\i);
+  %       \draw [arrow] (B-\i) -- (C-\i);
+         \draw [arrow] (input-\i) -- (A-\i);
+         \draw [arrow] (A-\j) -- (A-\i);
+  %       \draw [arrow] (B-\j) -- (B-\i);
+  %       \draw [arrow] (C-\j) -- (C-\i);
+         \draw [arrow] (A-\i) -- (output-\i);
+       }
+       \draw [arrow] (A-4) -- (s-output-A);
+       %\draw [arrow] (B-4) -- (s-output-B);
+       %\draw [arrow] (C-4) -- (s-output-C);
+
+  \end{tikzpicture}
+
 \end{frame}
@@
 \begin{frame}{The Basic RNN Cell}
+  We begin with the most basic of recurrent cell.
+Abstractly, a function $f:U\times S\to V\times S$ can be defined using two functions $u:\R^n\times \R^\ell\to \R^n$ and $v:\R^n\times\R^\ell \to \R^\ell$.
+We se ethe former function as providing the output, and the latter function as a state updating function.
+The most common definition for $u$ and $v$ for basic recurrent cells is as follows: $u(x, s) = f(A_ux+B_us)$, where $A$ nad $B$ are matrices, and $f$ is a non-linear function, and $v(x, s) = \tanh(A_vx+B_vs)$.
+Te parameters in this definition are $A_u$, $B_u$, $A_v$ and $B_v$.
+In tehsorflow, the code looks like:
 \end{frame}
@@
 \begin{frame}{The LSTM Cell}
+  There are a few problems with the architecture described above.
+First, it suffers greatly from the vanishing gradient problem, since there is no way to prevent gradients from becoming very small.
+Secondly, simple recurrent networks have a hard time remembering facts about the input sequence.
+To remedy this situation, long short-term memory cells were introduced.
+The overall structure of an LSTM cell is almost the same as the basic RNN cell.
+This time we take the previous output into account, which gives the structure as $f:\R^n\times \R^m\times \R^{\ell}\to \R^m\times \R^{\ell}$ which can be divided as two function $u:\R^{n}\times\R^m\times \R^{\ell}\to \R^m$ and $v:\R^n\times \R^m\times\R^{\ell}\to \R^{\ell}$.
+To define the functione $u$ and $v$, we use auxiliary functions $F$, $I$, and $O$ defined as follows
+  \begin{eqnarray*}
+    F(x, y) &=& \sigma(A_Fx + B_Fy + b_F)\\
+    I(x, y) &=& \sigma(A_Ix + B_Iy + b_I)\\
+    O(x, y) &=& \sigma(A_Ox + B_Oy + b_O)\\
+    S(x, y) &=& \sigma(A_Ox + B_Oy + b_O)\\
+  \end{eqnarray*}
+  The state update is given by $v(x, y, s) = F(x, y)\circ s + I(x, y)\circ \sigma_v(A_vx + B_vy + b_v)$, where $\circ$ denotes pointwise multiplication of vectors, and finally, the output can be defined as $u(x, y, s)=O(x, y)\circ \sigma_u(v(x, y, s))$.
+
 \end{frame}
@@
 \begin{frame}{The Basic GRU Cell}
+  A common variant of the long short term memory cell is the {\em gated recurrent unit cell}, more commonly known as GRU cells.
+The philosophy behind their design is similar to the long short term memory.
+Once again, each step of the computation takes into account a state vector and the output of the previous iteration.
+GRU's are functions  $f:\R^n\times \R^m\times \R^{\ell}\to \R^m\times \R^{\ell}$,  which can be divided as two function $u:\R^{n}\times\R^m\times \R^{\ell}\to \R^m$ and $v:\R^n\times \R^m\times\R^{\ell}\to \R^{\ell}$.
+To define the functione $u$ and $v$, we use auxiliary functions $U$ and  $R$ defined as follows
+  \begin{eqnarray*}
+    U(x, y) &=& \sigma(A_Ux + B_Uy + b_U)\\
+    R(x, y) &=& \sigma(A_Rx + B_Ry + b_R)\\
+  \end{eqnarray*}
+  The state update is given by $v(x, y, s) = U(x, y)\circ s + (1-s)\circ \sigma_h(A_v x + B_v(R(x, y)\circ y) + b_v)$.
 \end{frame}
@@
 
+\begin{frame}{A many-to-one example (output)}
+
+\end{frame}
+
 \begin{frame}{A many-to-many example}
@@
 
-\begin{frame}{A many-to-one example (code)}
+\begin{frame}{A many-to-many example (code)}
 \lstI
+\end{frame}
+
+\begin{frame}{A many-to-many example (output)}
+
 \end{frame}
@@
 
+ \begin{frame}{A many-to-one-to-many example (output)}
+ \end{frame}
+
+ \begin{frame}{A many-to-one-to-many example (output)}
+ \end{frame}
+
+ \begin{frame}{A many-to-one-to-many example (output)}
+ \end{frame}
+
+ \begin{frame}{A many-to-one-to-many example (output)}
+ \end{frame}
+
+ \begin{frame}{A many-to-one-to-many example (output)}
+ \end{frame}
+
+ \begin{frame}{A many-to-one-to-many example (output)}
+ \end{frame}
+
 \end{document}

### 2017-08-30 deeplearn/presentation/presentation_slides.tex
presentation slides

--- 2017-08-28 deeplearn/presentation/presentation_slides.tex
@@
 \usepackage{booktabs}
-\usetikzlibrary{shapes.geometric, arrows}
+\usetikzlibrary{shapes.geometric,arrows}
 \usepackage{listings}
 
-\usepackage{minted}
+%\usepackage{minted}
 
@@
 
+\begin{frame}{A Bird's Eye View Of Classical Neural Networks}
+
+\begin{itemize}
+  \item A neural network is a fancy way to describe a class of functions $f:\R^n\to \R^m$ parametrized by $\R^\ell$.
+  \item The layers of a NN represent function composition.
+  \item If $f$ is a function $F(\x, \w):\R^n\times \R^\ell\to\R^n$ such that $\C=\{\x\mapsto F(\x, \w):\w\in\R^n\}$.
+  \item Each vector $\w\in\R^\ell$ is in some sense a code for an element of $\C$.
+  \item When we are given a function $f:\R^n\to \R^m$, we try to find which element of $\C$ best represents $f$.
+
+%  \item That is to say, find $\w$ such that
+\end{itemize}
+
+%Let $f:\R^n\to \R^m$ be a function, and consider a class $\C$ of functions $g: \R^n\to \R^m$, which we assume to be parametrized by $\R^\ell$, in the sense that there is a function $F(\x, \w):\R^n\times \R^\ell\to\R^n$ such that $\C=\{\x\mapsto F(\x, \w):\w\in\R^n\}$.
+The generic problem of machine learning is to find a vector $\w\in\R^{\ell}$ such that $\ell(\w) = D(f(\x), F(\x, \w))$ is as small as possible, where $D$ represents a notion of distance between functions.
+% Given the definition of $D$, we see that $\ell(\x)\geq 0$, and therefore $\ell$ has a global minimum (which is not necessarily unique).
+In general, the global minimum of $\ell$ is not attained, in the sense that if $\epsilon=\min\{\ell(\w):\x\in\R^n\}$, then there does not necessarily exist a value $\w_0\in\R^\ell$ such that $\epsilon=\ell(\w_0)$.
+However, it can be approximated, that is to say there is a sequence $\w_i$ such that $\ell(\w_i)\to\epsilon$ as $i\to\infty$.
+\end{frame}
+
+\begin{frame}{A Bird's Eye View Of Classical Neural Networks}
+  \begin{itemize}
+    \item In tensorflow, the function we are trying to approximate is given in the form of a relation $f(\x)=\y$.
+    \item We provide {\em placeholders} for the values of $\x$ and $\y$.
+    \item The function $F(\x, \w)$ is defined by a {\em computation graph}.
+    \item For example, here is a linear regression, in which we try to approximate a function $f:\R^{10}\to\R^{15}$ using a function of the form $F(\x, \W\b)=\W\x+\b$.
+Here $\W$ is an $15\times 10$ matrix, and $\b\in\R^{15}$
+  \end{itemize}
+%In tensorflow, the function we are trying to approximate is given in the form of a relation $f(\x)=\y$.
+We provide {\em placeholders} for the values of $\x$ and $\y$.
+The function $F(\x, \w)$ is defined by a computation graph.
+For example, here is a linear regression, in which we try to approximate a function $f:\R^{10}\to\R^{15}$ using a function of the form $F(\x, \W\b)=\W\x+\b$.
+Here $\W$ is an $15\times 10$ matrix, and $\b\in\R^{15}$
+\end{frame}
+
+\defverbatim[colored]\lstI{
+\begin{lstlisting}[language=Python, basicstyle=\ttfamily, keywordstyle=\color{red}]
+x = tf.placeholder(shape=[None, 10])
+y = tf.placeholder(shape=[None, 15])
+W = tf.Variable(shape=[10, 15])
+b = tf.Variable(shape=[1, 15])
+F = tf.matmul(x, W) + b
+D = tf.mean_square_error(y, F)
+\end{lstlisting}
+}
+\begin{frame}{A Bird's Eye View Of Classical Neural Networks}
+\lstI
+\begin{itemize}
+  \item In tensorflow (and also Theano), vectors are represented as rows instead of columns
+  \item The first element in the {\em shape} parameter is there for batching, and {\tt None} indicates that the dimension can vary.
+  \item Classical neural networks are great for classification problems involving either fixed (finite) sets $f:A\to B$, where $A$ represents the set to be classified, and $B$ is the set of labels, or functions $f:\R^n\to B$.
+  %\item An example of the latter is image classification, since an $n\times m$ image can be directly represented as a vector in $\R^{\ell mn}$.
+
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Sequences}
+  \begin{itemize}
+%\item Classical neural networks are great for classification problems involving either fixed (finite) sets $f:A\to B$, where $A$ represents the set to be classified, and $B$ is the set of labels, or functions $f:\R^n\to B$.
+%\item An example of the latter is image classification, since an $n\times m$ image can be directly represented as a vector in $\R^{\ell mn}$.
+\item In many cases, however, the data to be classified is better represented as a (finite) sequence.
+\item For example,
+  \begin{itemize}
+    \item sentences are sequences of words,
+    \item words are sequences of letters,
+    \item movies are sequences of images,
+    \item sound files are sequences of amplitudes.
+  \end{itemize}
+\item It is possible to deal with sequences with ordinary neural networks, but we run into difficulties when we try to model dependency between different elements of a sequence.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Sequences}
+  Let $A$ be a finite set.
+  \begin{itemize}
+\item A finite sequence of elements of $A$ is a tuple $(a_1, ..., a_n)$, where $a_i\in A$ for every $i$.  Here the number $n$ is allowed to change.
+\item There is a special sequence, the empty sequence, denoted by $\epsilon$.
+\item The set of all finite sequences of elements of $A$ is denoted $A^*$.
+%\item From the definition we just gave, we have $A^*=\bigcup_{n\geq 0} A^n$, which is a good enough definition for our purpose.
+%If we are given two finite sets $A$ and $B$, there are many situations we can consider if we also have access to $A^*$ and $B^*$.
+\end{itemize}
+There are several types of functions on sequences:
+\begin{enumerate}
+  \item A function $f:A\to B^*$, which given an element of $A$ outputs a sequence of elements of $B$.
+This is called a {\em one-to-many} function.
+%We use this architecture in a sequence sorting neural network.
+The Tweet2Vec autoencoder also uses a one-to-many neural network.
+  \item A function $f:A^*\to B$, which is given a sequence of elments of $A$ and produces a single element of $B$.
+This is called a {\em many-to-one} function.
+%A good example of sich a function is the Buzzometer sentiment analysis tool which assigns a single sentiment value to messages which can vary in length.
+  \item A function $f:A^*\to B^*$ which is given a seuqence of elements of $A$< and outputs a sequence of elements of $B$.
+This is called a {\em many-to-many} function.
+%As an example of such a function we will show the implementation of a simple text generator.
+Translation, video captioning, part-of-speech tagging and speech recognition are all examples of many-to-many functions.
+\end{enumerate}
+\end{frame}
+
+%\begin{frame}{Sequences}
+%  \begin{enumerate}
+%    \item A function $f:A\to B^*$, which given an element of $A$ outputs a sequence of elements of $B$.
+This is called a {\em one-to-many} function.
+We use this architecture in a sequence sorting neural network.
+The Tweet2Vec autoencoder also uses a one-to-many neural network.
+%    \item A function $f:A^*\to B$, which is given a sequence of elments of $A$ and produces a single element of $B$.
+This is called a {\em many-to-one} function.
+A good example of sich a function is the Buzzometer sentiment analysis tool which assigns a single sentiment value to messages which can vary in length.
+%    \item A function $f:A^*\to B^*$ which is given a seuqence of elements of $A$< and outputs a sequence of elements of $B$.
+This is called a {\em many-to-many} function.
+As an example of such a function we will show the implementation of a simple text generator.
+Translation, video captioning, part-of-speech tagging and speech recognition are all examples of many-to-many functions.
+%  \end{enumerate}
+%\end{frame}
+
+\begin{frame}{Functions on Sequences}
+
+\begin{itemize}
+\item We can of course define functions $A\to B^*$, $A^*\to B$ and $A^*\to B^*$ directly (they are just sets, after all),
+
+\item it is more interesting and instructive to make use of the structure of $A^*$ and $B^*$ as sets of sequences of $A$ and $B$, and see how one can use a function $A\to B$ to define a function $A^*\to B^*$.
+%\item In particular, we will me making use of the iterative nature of the construction $A^*$ and $B^*$.
+\item Here is the simplest example: if $A=B$, and $f:A\to A$ is a function, then we can iterate $f$: $f^0(a)=a$ for every $a\in A$, and $f^{n+1}(a) = f(f^n(a))$.
+We terminate the iteration at the first value of $n$ for which $f^n(a)=f^{n-1}(a)$ (note that there is no guarantee that this will ever happen).
+\item This gives a function $f^\omega:A\to A^\omega$.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Functions on Sequences}
+We describe a more general situation.
+
+\begin{itemize}
+\item  Let $S$ be a (finite) set of {\em states} with a distinguished element $\perp\in S$, and
+
+\item  consider a function $f:A\times S\to B\times S$.
+
+\item  Let $a_1, ..., a_n$ be a finite sequence of elements of $A$.
+
+\item  We define two sequences $b_1, ..., b_n$ and $s_1, ..., s_n$ of elements of $B$ and $S$ respectively as follows:
+\begin{enumerate}
+\item $b_1, s_1 = f(a_1, \perp)$.
+
+\item  %Suppose that we have defined the elements $b_1, ..., b_n$ and $s_1, ..., s_n$.
+ %We define $b_{n+1}$ and $s_{n+1}$ via
+ $b_{n+1}, s_{n+1} = f(a_{n+1}, s_n)$.
+ \item We define $f^*(a_1...a_n)=b_1...b_n$, with $b_1, ..., b_n$ defined as above.
+\end{enumerate}
+%\item  This is a function $f^*:A^*\to B^*$, and it has the property that the length of the output sequence is the same as that of the input sequence.
+
+ %An interesting special case of this is when $S=B$. In this case, we can have $f:A\times B\to B$. We can define $b_1 = f(a_1, \perp)$, and $b_{n+1}=f(a_{n+1}, b_n)$.
+%
+% This is in fact the main type of function we use in our examples later.
+
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Functions on Sequences}
+  An interesting special case of this is when $S=B$. In this case, we can have $f:A\times B\to B$. We can define
+  \begin{enumerate}
+    \item $b_1 = f(a_1, \perp)$,
+    \item $b_{n+1}=f(a_{n+1}, b_n)$.
+  \end{enumerate}
+
+  This is in fact the main type of function we use in our examples later. Here's a note:
+\begin{itemize}
+  \item Defining a function $f:A\to B\times C$ is exactly the same as defining two function, $f_1:A\to B$ and $f_2:A\to C$.
+
+  \item In the case of $f:A\times S\to B\times S$, we are defining two function $f_1:A\times S\to B$ and $f_2:A\times S\to S$.
+\end{itemize}
+  The former takes an input and a state to produce an outputs.  The latter, taking the same state and input, produces a new state.
+\end{frame}
+
+%\begin{frame}{Functions on Sequences}
+%Finally, we tackle the case of a one-to-many function.
+Again we let $f:A\times S\to B\times S$ be any function.
+We define sequences $b_1, ..., b_i, ...$ and $s_1, ..., s_i, ...$ as follows.
+$b_1, s_1=f(a, \perp)$, and for every $n$, $b_{n+1}, s_{n+1}=f(a, s_n)$.
+In other words, we keep iterating $f$, feeding $a$ as input at every step.
+An imoprtant variant of this is when $a$ itself has a distinguished element $\perp_A$, in which case, after feeding $a$ as an input to $f$ in the first iteration, we subsequently feed it $\perp_A$, and get $b_{n+1}, s_{n+1}=f(\perp_A, s_n)$.
+We stop the iteration whenever $s_n=\perp$ (or any other predetermined element of $S$).
+Note that this may never happen, i.e.
+the sequence $b_n$ may go on forever.
+In fact, there is no way to give a general pattern of definition for $f:A\times S\to B\times S$ which will guarantee that $f^*(a)$ is finite for every $a$.
+%\end{frame}
+
+\begin{frame}{Recurrent Neural Networks}
+  \begin{itemize}
+  \item Recurrent neural networks are just a prescription for a class of functions of the form $f:\R^n\times \R^\ell\times\R^m\to \R^\ell\times \R^m$, where $\R^m$ is the space of parameters.
+  \item The set $\{0, 1\}^*$ of all finite sequences of $0$'s and $1$'s is countable,
+  \item and $\R^\ell$, is a real vector space,
+  is uncountable.
+  \item Therefore, we can encode every element $w\in\{0, 1\}^*$ as a vector in $S$.  Informally,
+  $\R^\ell$ is enough to encode any finite state space, and every possible value for the content of the tape of a Turing machine. We get:
+  {\em Recurrent neural networks are Turing-complete.}
+  \item This explains why recurrent neural networks seem to be able to produce results that other networks can't.
+  \item Training a recurrent network is the same as producing a Turing machine.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Recurrent Neural Networks}
+   A common way to represent a recurrent node is to use a box, the inside of which represent the definition of $f$.
+The vertical arrows represent the input and onput of the node, and the horizontal arrows represent the input and output state of the node.
+  \begin{center}
+    \begin{tikzpicture}[shorten >=1pt, ->, draw=black!50, node distance=1cm]
+      \tikzstyle{arrow} = [thick, ->, >=stealth]
+      \tikzstyle{startstop} = [rectangle, minimum width=2cm, minimum height=.75cm, text centered, draw=black, fill=black!20]
+      \node (input) [] at (0cm, 1cm) {};
+      \node (output) [] at (0cm, -1cm) {};
+      \node (s-input) [] at (-2cm, 0cm) {};
+      \node (s-output) [] at (2cm, 0cm) {};
+      \path[] node[startstop] (N) at (0cm, 0cm) {};
+      \draw [arrow] (input) -- (N);
+      \draw [arrow] (s-input) -- (N);
+      \draw [arrow] (N) -- (output);
+      \draw [arrow] (N) -- (s-output);
+  \end{tikzpicture}
+  \end{center}
+\end{frame}
+
+\begin{frame}{Recurrent Neural Networks}
+  \begin{itemize}
+  \item Just like any regular neural network layer, recurrent nodes can be composed.
+  \item The composition of two recurrent nodes is done by feeding the output of one node into the input of the other, and concatenating their state space.
+  \item Formally, if $f:\R^n\times \R^\ell\times\R^m\to \R^k\times \R^\ell$ and $g:\R^k\times \R^\ell\times\R^m\to \R^q\times \R^\ell$ are two recurrent nodes, $$(f;g):\R^m\times\R^\ell\times\R^m\to\R^q\times\R^\ell\times\R^l$$ defined by $$(f;g)(\x, \s_1, \s_2) = (\y, \s'_1, \s'_2)$$ where $$f(\x, \s_1) = \x', \s'_1$$ and $$g(\x', \s_2) = \y, \s'_2$$.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Recurrent Neural Networks}
+  Graphically, we can represent this by stacking the boxes representing $f$ and $g$ on top of one another:
+
+  \begin{center}
+  \begin{tikzpicture}[shorten >=1pt, ->, draw=black!50, node distance=1cm]
+      \tikzstyle{arrow} = [thick, ->, >=stealth]
+      \tikzstyle{startstop} = [rectangle, minimum width=1.5cm, minimum height=.6cm, text centered, draw=black, fill=black!20]
+      \node (input) [] at (0cm, 1cm) {};
+      \node (output) [] at (0cm, -3cm) {};
+      \node (s-input-A) [] at (-2cm, 0cm) {};
+      \node (s-output-A) [] at (2cm, 0cm) {};
+      \node (s-input-B) [] at (-2cm, -1cm) {};
+      \node (s-output-B) [] at (2cm, -1cm) {};
+      \node (s-input-C) [] at (-2cm, -2cm) {};
+      \node (s-output-C) [] at (2cm, -2cm) {};
+      \path[] node[startstop] (A) at (0cm, 0cm) {};
+      \path[] node[startstop] (B) at (0cm, -1cm) {};
+      \path[] node[startstop] (C) at (0cm, -2cm) {};
+      \draw [arrow] (input) -- (A);
+      \draw [arrow] (s-input-A) -- (A);
+      \draw [arrow] (s-input-B) -- (B);
+      \draw [arrow] (s-input-C) -- (C);
+      \draw [arrow] (A) -- (s-output-A);
+      \draw [arrow] (B) -- (s-output-B);
+      \draw [arrow] (C) -- (s-output-C);
+      \draw [arrow] (A) -- (B);
+      \draw [arrow] (B) -- (C);
+      \draw [arrow] (C) -- (output);
+  \end{tikzpicture}
+  \end{center}
+\end{frame}
+
 \begin{frame}{Training RNN's}
+  How do we train recurrent neural networks?
+We use a variant of back propagation, just like a regular neural network.
+Heuristically, we unroll the recurrent network ``infinitely'' many times, until it looks like an ordinary very deep neural network.
+In practice, we only unroll the network a large but finite number of times, and treat it like an ordinary neural network.
+
+  \begin{tikzpicture}[shorten >=1pt, ->, draw=black!50, node distance=1cm]
+      \tikzstyle{arrow} = [thick, ->, >=stealth]
+      \tikzstyle{startstop} = [rectangle, minimum width=1.5cm, minimum height=.6cm, text centered, draw=black, fill=black!20]
+      \node (input) [] at (0cm, 1cm) {};
+      \node (output) [] at (0cm, -1cm) {};
+      \node (s-input-A) [] at (-2cm, 0cm) {};
+      \node (s-output-A) [] at (9.5cm, 0cm) {};
+      %\node (s-input-B) [] at (-2cm, -1cm) {};
+      %\node (s-output-B) [] at (10cm, -1cm) {};
+      %\node (s-input-C) [] at (-2cm, -2cm) {};
+      %\node (s-output-C) [] at (10cm, -2cm) {};
+      \path[] node[startstop] (A-0) at (0cm, 0cm) {};
+      %\path[] node[startstop] (B-0) at (0cm, -1cm) {};
+      %\path[] node[startstop] (C-0) at (0cm, -2cm) {};
+      \draw [arrow] (input) -- (A);
+      \draw [arrow] (s-input-A) -- (A-0);
+      %\draw [arrow] (s-input-B) -- (B-0);
+      %\draw [arrow] (s-input-C) -- (C-0);
+      %\draw [arrow] (A) -- (B);
+      %\draw [arrow] (B) -- (C);
+      \draw [arrow] (A) -- (output);
+
+       \foreach \i/\j in {1/0, 2/1, 3/2, 4/3}
+       {
+         \node (input-\i) [] at (2*\i cm, 1cm) {};
+         \node (output-\i) [] at (2*\i cm, -1cm) {};
+         \path[] node[startstop] (A-\i) at (2*\i cm, 0cm) {};
+      %   \path[] node[startstop] (B-\i) at (2*\i cm, -1cm) {};
+    %     \path[] node[startstop] (C-\i) at (2*\i cm, -2cm) {};
+    %     \draw [arrow] (A-\i) -- (B-\i);
+  %       \draw [arrow] (B-\i) -- (C-\i);
+         \draw [arrow] (input-\i) -- (A-\i);
+         \draw [arrow] (A-\j) -- (A-\i);
+  %       \draw [arrow] (B-\j) -- (B-\i);
+  %       \draw [arrow] (C-\j) -- (C-\i);
+         \draw [arrow] (A-\i) -- (output-\i);
+       }
+       \draw [arrow] (A-4) -- (s-output-A);
+       %\draw [arrow] (B-4) -- (s-output-B);
+       %\draw [arrow] (C-4) -- (s-output-C);
+
+  \end{tikzpicture}
+
 \end{frame}
@@
 \begin{frame}{The Basic RNN Cell}
+  We begin with the most basic of recurrent cell.
+  \begin{itemize}
+    \item Abstractly, a function $f:U\times S\to V\times S$ can be defined using two functions $u:\R^n\times \R^\ell\to \R^n$ and $v:\R^n\times\R^\ell \to \R^\ell$.
+    \item The former function as providing the output
+    \item The latter function as a state updating function.
+    \item The most common definition for $u$ and $v$ for basic recurrent cells is as follows: $u(x, s) = f(A_ux+B_us)$, and $v(x, s) = \tanh(A_vx+B_vs)$, where
+    \begin{itemize}
+      \item $A_u$, $A_v$, $B_u$ and $B_v$ are matrices,
+      \item $f$ is a non-linear function
+    \end{itemize}
+  \end{itemize}
 \end{frame}
@@
 \defverbatim[colored]\lstI{
-\begin{lstlisting}[language=Python,basicstyle=\ttfamily,keywordstyle=\color{red}]
-def basic_rnn_cell(input_tensor, state_tensor, output_dimension):
-  input_dimension = input_tensor.get_shape()[1]
-  state_dimension = input_tensor.get_shape()[1]
-  A_u = tf.Variable(shape=[input_dimension, output_dimension])
-  B_u = tf.Variable(shape=[state_dimension, output_dimension])
-  A_v = tf.Variable(shape=[input_dimension, state_dimension])
-  B_v = tf.Variable(shape=[state_dimension, state_dimension])
-  output_tensor = tf.relu(tf.matmul(input_tensor, A_u) + \
-                            tf.matmul(state_tensor, B_u))
-  new_state_tensor = tf.tanh(tf.matmul(input_tensor, A_v) + \
-                                  tf.matmul(state_tensor, B_v))
-  return output_tensor, new_state_tensor
+\begin{lstlisting}[language=Python, basicstyle=\ttfamily, keywordstyle=\color{red}]
+def basic_rnn_cell(i_tensor, s_tensor, o_dim):
+  i_dim = input_tensor.get_shape()[1]
+  s_dim = input_tensor.get_shape()[1]
+  A_u = tf.Variable(shape=[i_dim, o_dim])
+  B_u = tf.Variable(shape=[s_dim, o_dim])
+  A_v = tf.Variable(shape=[i_dim, s_dim])
+  B_v = tf.Variable(shape=[s_dim, s_dim])
+  o_tensor = tf.relu(tf.matmul(i_tensor, A_u) + \
+                     tf.matmul(s_tensor, B_u))
+  ns_tensor = tf.tanh(tf.matmul(i_tensor, A_v) + \
+                      tf.matmul(s_tensor, B_v))
+  return o_tensor, ns_tensor
 \end{lstlisting}
@@
 \begin{frame}{The Basic RNN Cell (code)}
+  In tensorflow, the code looks like:
+
 \lstI
@@
 
-\begin{frame}{The LSTM Cell}
+\begin{frame}{The Long-Short-Term-Memory Cell}
+
+\begin{itemize}
+  \item It is a cell that can remember elements of a sequence it has already seen
+  \item Its state has two parts: a state and a memory vector;
+  \item A special {\em forget gate} controls how much of the memory vector gets forgotten
+  \item An {\em input gate} controls how new information gets stored into the memory vector
+  \item The usual state update and output functions
+\end{itemize}
+\end{frame}
+
+\begin{frame}{The Long-Short-Term-Memory Cell}
+We define:
+  %There are a few problems with the architecture described above.
+First, it suffers greatly from the vanishing gradient problem, since there is no way to prevent gradients from becoming very small.
+Secondly, simple recurrent networks have a hard time remembering facts about the input sequence.
+To remedy this situation, long short-term memory cells were introduced.
+  %The overall structure of an LSTM cell is almost the same as the basic RNN cell: $f:\R^n\times \times \R^{\ell}\to \R^m\times \R^{\ell}$ which can be divided as two function
+  \begin{itemize}
+  \item $u:\R^{n}\times\R^m\times \R^{\ell}\to \R^m$ and
+  \item $v:\R^n\times \R^m\times\R^{\ell}\to \R^{\ell}$
+  \end{itemize}
+
+  using auxiliary functions $F$, $I$, $O$, and $S$ defined as follows
+  \begin{eqnarray*}
+    F(x, y) &=& \sigma(A_Fx + B_Fy + b_F)\\
+    I(x, y) &=& \sigma(A_Ix + B_Iy + b_I)\\
+    O(x, y) &=& \sigma(A_Ox + B_Oy + b_O)\\
+    S(x, y) &=& \tanh(A_Ox + B_Oy + b_O)\\
+  \end{eqnarray*}
+%\end{frame}
+%
+%
+%
+%\begin{frame}{The LSTM Cell}
+  The state update is given by $$v(x, y, s) = F(x, y)\circ s + I(x, y)\circ \sigma_v(A_vx + B_vy + b_v)$$ where $\circ$ denotes pointwise multiplication of vectors, and finally, the output can be defined as $$u(x, y, s)=O(x, y)\circ \sigma_u(v(x, y, s))$$\
+
 \end{frame}
@@
 \defverbatim[colored]\lstI{
-\begin{lstlisting}[language=Python,basicstyle=\ttfamily,keywordstyle=\color{red}]
-def lstm_gate(input_tensor, previous_output, port_op):
-  A = tf.Variable(shape=[N, L])
-  B = tf.Variable(shape=[L, L])
-  b = tf.Variable(shape=[L, L])
-  x = tf.matmul(input_tensor, A)+ tf.matmul(previous_output, B) + b
-  return post_op(x)
+\begin{lstlisting}[language=Python, basicstyle=\ttfamily, keywordstyle=\color{red}]
+def lstm_gate(input_tensor, previous_output, op):
+  _, N = input_tensor.get_shape()
+  _, output_dim = previous_output.get_shape()
+  A = tf.Variable(shape=[N, output_dim])
+  B = tf.Variable(shape=[output_dim, output_dim])
+  b = tf.Variable(shape=[1, output_dim])
+  x = tf.matmul(input_tensor, A) + \
+        tf.matmul(previous_output, B) + b
+  return op(x)
 
-def lstm_cell(input_tensor, output, state):
+def lstm_cell(input_tensor, output):
+  _, output_dim = output.get_shape()
   F = lstm_gate(input_tensor, output, tf.sigmoid)
@@
   O = lstm_gate(input_tensor, output, tf.sigmoid)
-  S = lstm_gate(input_tensorm output, tf.tanh)
+  S = lstm_gate(input_tensor, output, tf.tanh)
   new_state = tf.mul(output, F) + tf.mul(I, S)
@@
 \begin{frame}{The Basic GRU Cell}
+  %A common variant of the long short term memory cell is the {\em gated recurrent unit cell}, more commonly known as GRU cells.
+The philosophy behind their design is similar to the long short term memory.
+Once again, each step of the computation takes into account a state vector and the output of the previous iteration.
+  \begin{itemize}
+    \item GRU's are a simplification of the LSTM cell.
+    \item Compared to LSTM the GRU lacks an output gate.
+    \item Its performance is on par with LSTM cells in most applications.
+  \end{itemize}
+  %$f:\R^n\times \R^m\times \R^{\ell}\to \R^m\times \R^{\ell}$, which can be divided as two function
+  %\begin{itemize}
+  %  $u:\R^{n}\times\R^m\times \R^{\ell}\to \R^m$
+  %   $v:\R^n\times \R^m\times\R^{\ell}\to \R^{\ell}$
+  %   \end{itemize}
+  To define the functions $u$ and $v$, we use auxiliary functions $U$(pdate) and  $R$(eset) defined as follows
+  \begin{eqnarray*}
+    U(x, y) &=& \sigma(A_Ux + B_Uy + b_U)\\
+    R(x, y) &=& \sigma(A_Rx + B_Ry + b_R)\\
+  \end{eqnarray*}
+  The state update (and output) is given by $$v(x, y, s) = U(x, y)\circ s + (1-s)\circ \sigma_h(A_v x + B_v(R(x, y)\circ y) + b_v)$$
+  %The output is the same as the state.
 \end{frame}
@@
 \defverbatim[colored]\lstI{
-\begin{lstlisting}[language=Python,basicstyle=\ttfamily,keywordstyle=\color{red}]
+\begin{lstlisting}[language=Python, basicstyle=\ttfamily, keywordstyle=\color{red}]
 def gru_gate(input_tensor, previous_output, port_op):
-  A = tf.Variable(shape=[N, L])
-  B = tf.Variable(shape=[L, L])
-  b = tf.Variable(shape=[L, L])
-  x = tf.matmul(input_tensor, A)+ tf.matmul(previous_output, B) + b
+  _, N = input_tensor.get_shape()
+  _, output_dim = previous_output.get_shape()
+  A = tf.Variable(shape=[N, output_dim])
+  B = tf.Variable(shape=[output_dim, output_dim])
+  b = tf.Variable(shape=[output_dim, output_dim])
+  x = tf.matmul(input_tensor, A) + \
+        tf.matmul(previous_output, B) + b
   return post_op(x)
@@
   O = gru_gate(input, tf.mul(R, output))
-  return tf.mul(R, output) + tf.mul((1-R)O)
+  return [tf.mul(R, output) + tf.mul((1-R), O)]*2
 \end{lstlisting}
@@
 \begin{frame}{A many-to-one example}
+\begin{itemize}
+  \item As an example of a many-to-one RNN, consider the Buzzometer sentiment analysis tool.
+  \item On input of a sequence of {\bf characters}, we output one of 4 classes: negative, neutral, positive and irrelevant.
+  \item For training, the network was unrolled to 256 characters, which is twise the average length of a message in out database.
+  \item Longer messaages were truncated, and shorter messages were padded with $0$.
+\end{itemize}
+
+The architecture is:
+\begin{itemize}
+  \item A 3 layer {\em bi-directional RNN}
+  \item We keep the last output of the forward and the backward networks, and combine them linearly: $\w=W_1 \v_f + W_2\v_b$
+  \item The vector $\w$ is then projected to $\R^4$.
+\end{itemize}
 \end{frame}
@@
 \begin{frame}{A many-to-one example (model architecture)}
-  \begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
-      \tikzstyle{arrow} = [thick,->,>=stealth]
-       \tikzstyle{every pin edge}=[<-,shorten <=1pt]
-       \tikzstyle{neuron}=[circle,fill=black!50,minimum size=6pt,inner sep=0pt]
+  \begin{tikzpicture}[shorten >=1pt, ->, draw=black!50, node distance=1cm]
+      \tikzstyle{arrow} = [thick, ->, >=stealth]
+       \tikzstyle{every pin edge}=[<-, shorten <=1pt]
+       \tikzstyle{neuron}=[circle, fill=black!50, minimum size=6pt, inner sep=0pt]
 
       \tikzstyle{startstop} = [rectangle, minimum width=1.5cm, minimum height=.6cm, text centered, draw=black, fill=black!20]
-      \node (input) [] at (0cm,1cm) {};
-      \node (output) [] at (0cm,-2.25cm) {};
-      \node (s-input-A) [] at (-2cm,0cm) {};
-      \node (s-output-A) [] at (10cm,0cm) {};
-      \node (s-input-B) [] at (-2cm,-1cm) {};
-      \node (s-output-B) [] at (10cm,-1cm) {};
-      \node (s-input-C) [] at (-2cm,-2cm) {};
-      \node (s-output-C) [] at (10cm,-2cm) {};
-      \path[] node[startstop] (A-0) at (0cm,0cm) {};
-      \path[] node[startstop] (B-0) at (0cm,-1cm) {};
-      \path[] node[startstop] (C-0) at (0cm,-2cm) {};
+      \node (input) [] at (0cm, 1cm) {};
+      \node (output) [] at (0cm, -2.25cm) {};
+      \node (s-input-A) [] at (-2cm, 0cm) {};
+      \node (s-output-A) [] at (10cm, 0cm) {};
+      \node (s-input-B) [] at (-2cm, -1cm) {};
+      \node (s-output-B) [] at (10cm, -1cm) {};
+      \node (s-input-C) [] at (-2cm, -2cm) {};
+      \node (s-output-C) [] at (10cm, -2cm) {};
+      \path[] node[startstop] (A-0) at (0cm, 0cm) {};
+      \path[] node[startstop] (B-0) at (0cm, -1cm) {};
+      \path[] node[startstop] (C-0) at (0cm, -2cm) {};
       \draw [arrow] (input) -- (A-0);
@@
 
-       \foreach \i/\j in {1/0,2/1,3/2,4/3}
+       \foreach \i/\j in {1/0, 2/1, 3/2, 4/3}
        {
-         \node (input-\i) [] at (2*\i cm,1cm) {};
-         \node (output-\i) [] at (2*\i cm,-2.25cm) {};
-         \path[] node[startstop] (A-\i) at (2*\i cm,0cm) {};
-         \path[] node[startstop] (B-\i) at (2*\i cm,-1cm) {};
-         \path[] node[startstop] (C-\i) at (2*\i cm,-2cm) {};
+         \node (input-\i) [] at (2*\i cm, 1cm) {};
+         \node (output-\i) [] at (2*\i cm, -2.25cm) {};
+         \path[] node[startstop] (A-\i) at (2*\i cm, 0cm) {};
+         \path[] node[startstop] (B-\i) at (2*\i cm, -1cm) {};
+         \path[] node[startstop] (C-\i) at (2*\i cm, -2cm) {};
          \draw [arrow] (A-\i) -- (B-\i);
@@
 
-       \foreach \name / \x in {1,...,15} {
+       \foreach \name / \x in {1, ..., 15} {
            \path[xshift=-1cm]
-               node[neuron] (I-\name) at (0.3*\x,-3.5cm) {};
+               node[neuron] (I-\name) at (0.3*\x, -3.5cm) {};
                \path (output) edge (I-\name);
@@
 
-       \foreach \name / \x in {1,...,15} {
+       \foreach \name / \x in {1, ..., 15} {
            \path[xshift=4.5cm]
-               node[neuron] (J-\name) at (0.3*\x,-3.5cm) {};
+               node[neuron] (J-\name) at (0.3*\x, -3.5cm) {};
               \path (output-4) edge (J-\name); }
 
-       \foreach \name / \x in {1,...,15}{
+       \foreach \name / \x in {1, ..., 15}{
            \path[xshift=1.5cm]
-               node[neuron] (K-\name) at (0.3*\x,-4.5cm) {};
+               node[neuron] (K-\name) at (0.3*\x, -4.5cm) {};
           \path (I-\name) edge (K-\name);
@@
 
-          \foreach \name / \x in {1,...,4}{
-              \path[xshift=3cm] node[neuron] (O-\name) at (0.3*\x,-5.25cm) {};
-              \foreach \j in {1,...,15}{
+          \foreach \name / \x in {1, ..., 4}{
+              \path[xshift=3cm] node[neuron] (O-\name) at (0.3*\x, -5.25cm) {};
+              \foreach \j in {1, ..., 15}{
              \path (K-\j) edge (O-\name);}
@@
 \defverbatim[colored]\lstI{
-\begin{lstlisting}[language=Python,basicstyle=\ttfamily,keywordstyle=\color{red}]
-SEQ_LENGTH = 256
-E_DIM = 128
-STATE_DIM = 512
-NUM_CLASSES = 4
-def inference():
-    model_input = tf.placeholder('uint8', shape=[None, SEQ_LENGTH])
-    _ = tf.one_hot(Globals.model_input, depth=E_DIM, axis=-1)
+\begin{lstlisting}[language=Python, basicstyle=\ttfamily, keywordstyle=\color{red}]
+model_input = tf.placeholder(shape=[SEQ_LENGTH])
+_ = tf.one_hot(model_input, depth=E_DIM, axis=-1)
 _ = tf.reshape(_, [-1, SEQ_LENGTH, E_DIM])
@@
 bw = multi_layer_rnn(N_LAYERS, STATE_DIM)
-    output, _ = tf.nn.bidirectional_dynamic_rnn(fw, bw, _, dtype=tf.float32)
-    fw_output = tf.reshape(output[0][:, -1:], [-1, STATE_DIM])
-    bw_output = tf.reshape(output[1][:, :1], [-1, STATE_DIM])
+OP = tf.nn.bidirectional_dynamic_rnn
+output, _ = OP(fw, bw, _, dtype=tf.float32)
+fw_output = tf.reshape(output[0][:, -1:],
+                            [-1, STATE_DIM])
+bw_output = tf.reshape(output[1][:, :1],
+                            [-1, STATE_DIM])
 f = project(fw_output, E_DIM)
@@
 e = tf.add(f, b)
-    Globals.model_output = project(e, NUM_CLASSES)
-    Globals.prediction = tf.cast(tf.argmax(Globals.model_output, 1), tf.uint8)
-    return Globals.model_input, Globals.model_output
+model_output = project(e, NUM_CLASSES)
+prediction = tf.argmax(model_output, 1)
 \end{lstlisting}
 }
-
 \begin{frame}{A many-to-one example (code)}
@@
 
+\begin{frame}{A many-to-one example (output)}
+\begin{center}
+  \begin{tabular}{lcccc}
+    {} & {\bf Negative} & {\bf Neutral} & {\bf Positive} & {\bf Irrelevant}\\
+    {\bf Negative}   & {291} & {113} & {50} & {127}\\
+    {\bf Neutral}    & {108} & {113} & {50} & {85}\\
+    {\bf Positive}   & {0} & {0} & {0} & {0}\\
+    {\bf Irrelevant} & {0} & {0} & {0} & {0}\\
+  \end{tabular}
+\end{center}
+%\end{frame}
+%
+%\begin{frame}{A many-to-one example (output)}
+  \begin{center}
+  \begin{tabular}{lcccc}
+    {} & {\bf Negative} & {\bf Neutral} & {\bf Positive} & {\bf Irrelevant}\\
+    {\bf Negative}   & {292} & {52} & {36} & {55}\\
+    {\bf Neutral}    & {61} & {176} & {21} & {62}\\
+    {\bf Positive}   & {5} & {15} & {65} & {15}\\
+    {\bf Irrelevant} & {33} & {28} & {14} & {70}\\
+  \end{tabular}
+\end{center}
+
+%\end{frame}
+%
+%\begin{frame}{A many-to-one example (output)}
+  \begin{center}
+  \begin{tabular}{lcccc}
+    {} & {\bf Negative} & {\bf Neutral} & {\bf Positive} & {\bf Irrelevant}\\
+    {\bf Negative}   & {319} & {38} & {11} & {43}\\
+    {\bf Neutral}    & {20} & {188} & {2} & {20}\\
+    {\bf Positive}   & {10} & {18} & {91} & {10}\\
+    {\bf Irrelevant} & {27} & {27} & {7} & {169}\\
+  \end{tabular}
+\end{center}
+\end{frame}
+
+\begin{frame}{A many-to-one example (output)}
+  \begin{center}
+  \begin{tabular}{lcccc}
+    {} & {\bf Negative} & {\bf Neutral} & {\bf Positive} & {\bf Irrelevant}\\
+    {\bf Negative}   & {334} & {4} & {4} & {9}\\
+    {\bf Neutral}    & {3} & {283} & {2} & {7}\\
+    {\bf Positive}   & {1} & {0} & {125} & {2}\\
+    {\bf Irrelevant} & {1} & {0} & {4} & {221}\\
+  \end{tabular}
+\end{center}
+%
+%\end{frame}
+%
+%\begin{frame}{A many-to-one example (output)}
+  \begin{center}
+  \begin{tabular}{lcccc}
+    {} & {\bf Negative} & {\bf Neutral} & {\bf Positive} & {\bf Irrelevant}\\
+    {\bf Negative}   & {321} & {16} & {3} & {20}\\
+    {\bf Neutral}    & {8} & {286} & {4} & {6}\\
+    {\bf Positive}   & {4} & {2} & {125} & {16}\\
+    {\bf Irrelevant} & {11} & {9} & {1} & {168}\\
+  \end{tabular}
+\end{center}
+\end{frame}
+
 \begin{frame}{A many-to-many example}
+  \begin{itemize}
+    \item we define an architecture that generates text in the style of a particular author, or body of text.
+    \item The architecture is very simple, and consists of three stacked GRU cells.
+    \item This model will also expose some of the challenges in training recurrent networks.
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}{A many-to-many example (model architecture)}
-  \begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
-      \tikzstyle{arrow} = [thick,->,>=stealth]
+  \begin{tikzpicture}[shorten >=1pt, ->, draw=black!50, node distance=1cm]
+      \tikzstyle{arrow} = [thick, ->, >=stealth]
       \tikzstyle{startstop} = [rectangle, minimum width=1.5cm, minimum height=.6cm, text centered, draw=black, fill=black!20]
-      \tikzstyle{neuron}=[circle,fill=black!50,minimum size=6pt,inner sep=0pt]
+      \tikzstyle{neuron}=[circle, fill=black!50, minimum size=6pt, inner sep=0pt]
 
-      \node (input) [] at (0cm,1cm) {};
-      \node (output) [] at (0cm,-3cm) {};
-      \node (s-input-A) [] at (-2cm,0cm) {};
-      \node (s-output-A) [] at (10cm,0cm) {};
-      \node (s-input-B) [] at (-2cm,-1cm) {};
-      \node (s-output-B) [] at (10cm,-1cm) {};
-      \node (s-input-C) [] at (-2cm,-2cm) {};
-      \node (s-output-C) [] at (10cm,-2cm) {};
-      \path[] node[startstop] (A-0) at (0cm,0cm) {};
-      \path[] node[startstop] (B-0) at (0cm,-1cm) {};
-      \path[] node[startstop] (C-0) at (0cm,-2cm) {};
+      \node (input) [] at (0cm, 1cm) {};
+      \node (output) [] at (0cm, -3cm) {};
+      \node (s-input-A) [] at (-2cm, 0cm) {};
+      \node (s-output-A) [] at (10cm, 0cm) {};
+      \node (s-input-B) [] at (-2cm, -1cm) {};
+      \node (s-output-B) [] at (10cm, -1cm) {};
+      \node (s-input-C) [] at (-2cm, -2cm) {};
+      \node (s-output-C) [] at (10cm, -2cm) {};
+      \path[] node[startstop] (A-0) at (0cm, 0cm) {};
+      \path[] node[startstop] (B-0) at (0cm, -1cm) {};
+      \path[] node[startstop] (C-0) at (0cm, -2cm) {};
       \draw [arrow] (input) -- (A-0);
@@
 
-       \foreach \name / \x in {1,...,15}{
+       \foreach \name / \x in {1, ..., 15}{
            \path[xshift=0cm]
-               node[neuron] (K-\name) at (0.5*\x,-3.5cm) {};
+               node[neuron] (K-\name) at (0.5*\x, -3.5cm) {};
           \path (output-2) edge (K-\name);
@@
 
-          \foreach \name / \x in {1,...,10}{
-              \path[xshift=1cm] node[neuron] (O-\name) at (0.6*\x,-5.75cm) {};
-              \foreach \j in {1,...,15}{
+          \foreach \name / \x in {1, ..., 10}{
+              \path[xshift=1cm] node[neuron] (O-\name) at (0.6*\x, -5.75cm) {};
+              \foreach \j in {1, ..., 15}{
              \path (K-\j) edge (O-\name);}
@@
 
+\begin{frame}{A many-to-one example (output)}
+\begin{itemize}
+  \item For training we network was unrolled to 30 characters, and the training was done in batches of 32 strings.
+  \item The main challenge for training is that the strings should continue from batch to batch.
+\item For example, consider training a similar network unrolled to 2 characters with a batch size of 3.
+  \item Consider the sentence {\tt The quick brown fox jumps over the lazy dog}.
+  \item The first step is to cut the string into substrings of length 2:
+  {\tt |Th|e\textvisiblespace |qu|ic|k\textvisiblespace |br|ow|n\textvisiblespace |fo|x\textvisiblespace |ju|mp|s\textvisiblespace |ov|er| \textvisiblespace t|he|\textvisiblespace l|az|y\textvisiblespace |do|g.|}.
+  \item In a normal batching situation, we would then cut this liet of strings into chunks of length 3, like so:
+{\tt |(Th|e\textvisiblespace |qu)|(ic|k\textvisiblespace |br)|(ow|n\textvisiblespace |fo)|(x\textvisiblespace |ju|mp)| (s\textvisiblespace |ov|er)|(\textvisiblespace t|he|\textvisiblespace l)|(az|y\textvisiblespace |do)|} but this causes a problem.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{A many-to-many example (output)}
+
+    {\tt |Th|e\textvisiblespace |qu|ic|k\textvisiblespace |br|ow|n\textvisiblespace |fo|x\textvisiblespace |ju|mp|s\textvisiblespace |ov|er|\textvisiblespace t|he|\textvisiblespace l|az|y |do|g.|}.
+
+    {\tt |(Th|e\textvisiblespace |qu)|(ic|k\textvisiblespace |br)|(ow|n\textvisiblespace |fo)|(x |ju|mp)|(s\textvisiblespace |ov|er)|(\textvisiblespace t|he|\textvisiblespace l)|(az|y\textvisiblespace |do)|}
+
+  \begin{itemize}
+\item The first sequence of the first batch is {\tt Th}, which will leave the network in a certain state $s$.
+\item For this state to be updated properly, the first element of the second batch should be {\tt |e\textvisiblespace |}, and {\em not} {\tt |ic|}.
+\item We must therefore use a different batching strategy.
+The string will give us 7 batches in total, so we number the subsequences with a batch index from 1 to 7 in order, starting over at 1 when we run out of indices.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{A many-to-one example (output)}
+  \begin{center}
+\begin{tabular}{ccccccc}
+  {1}&{2}&{3}&{4}&{5}&{6}&{7}\\\hline
+  Th & e\textvisiblespace &  qu & ic & k\textvisiblespace &  br & ow \\
+  n\textvisiblespace & fo & x\textvisiblespace &  ju & mp & s\textvisiblespace &  ov \\
+   er & \textvisiblespace t & he & \textvisiblespace l & az & y\textvisiblespace &  do \\
+  g. &  {}&{}&{}&{}&{}&{}
+\end{tabular}
+\end{center}
+\end{frame}
+
 \defverbatim[colored]\lstI{
-\begin{lstlisting}[language=Python,basicstyle=\ttfamily,keywordstyle=\color{red}]
-SEQ_LENGTH = 256
-E_DIM = 128
-STATE_DIM = 512
-N_LAYERS = 3
-
-def inference():
-    model_input = tf.placeholder('uint8', shape=[None, SEQ_LENGTH])
-    _ = tf.one_hot(Globals.model_input, depth=E_DIM, axis=-1)
+\begin{lstlisting}[language=Python, basicstyle=\ttfamily, keywordstyle=\color{red}]
+model_input = tf.placeholder(shape=[None, SEQ_LENGTH])
+initial_state = tf.placeholder(shape=[N_LAYERS, S_DIM])
+_ = tf.one_hot(model_input, depth=E_DIM, axis=-1)
 encode = multi_layer_rnn(N_LAYERS, STATE_DIM)
-    state_tuple = tuple(tf.unstack(Globals.initial_state, axis=0))
-    output, state = tf.nn.dynamic_rnn(encode, _,
+state_tuple = tuple(tf.unstack(initial_state, axis=0))
+OP = tf.nn.dynamic_rnn
+output, state = OP(encode, _,
                    dtype=tf.float32,
@@
 output = project(output, E_DIM)
-    out = tf.cast(tf.argmax(output, 1), tf.uint8)
 out = tf.reshape(out, [-1, SEQ_LENGTH])
-    Globals.generated_sequence = out
-    Globals.generated_characters = tf.nn.softmax(output)
-    Globals.model_output = output
-    Globals.state = state
+model_output = tf.nn.softmax(output)
+output = tf.argmax(output, 1)
 \end{lstlisting}
@@
 
-\begin{frame}{A many-to-one example (code)}
+\begin{frame}{A many-to-many example (code)}
 \lstI
@@
 \defverbatim[colored]\lstI{
-\begin{lstlisting}[language=Python,basicstyle=\ttfamily,keywordstyle=\color{red}]
+\begin{lstlisting}[language=Python, basicstyle=\ttfamily, keywordstyle=\color{red}]
 def generate_text(length, session=None):
@@
     while len(generated_text) < length:
-        feed_dict = {Globals.model_input: character,
-                     Globals.initial_state: istate}
-        next_char, state = session.run([Globals.generated_characters,
-                                        Globals.state],
-                                       feed_dict=feed_dict)
-        next_char = np.asarray(next_char).astype('float64')
-        next_char = next_char / next_char.sum()
+        feed_dict = {model_input: character,
+                     initial_state: istate}
+        next_char, state = session.run(
+          [out, state], feed_dict=feed_dict)
         op = np.random.multinomial
-        next_char_id = op(1, next_char.squeeze(), 1).argmax()
-        next_char_id = next_char_id if chr(next_char_id) in \
+        next_char_id = op(1, next_char.squeeze(), 1)
+        next_char_id = next_char_id.argmax()
+        next_char_id = next_char_id \
+                if chr(next_char_id) in \
                     string.printable else ord(" ")
@@
 
+\begin{frame}{A many-to-many example (output)}
+
+\end{frame}
+
+\begin{frame}{A many-to-many example (output)}
+
+\end{frame}
+
+\begin{frame}{A many-to-many example (output)}
+
+\end{frame}
+
+\begin{frame}{A many-to-many example (output)}
+
+\end{frame}
+
+\begin{frame}{A many-to-many example (output)}
+
+\end{frame}
+
+\begin{frame}{A many-to-many example (output)}
+
+\end{frame}
+
+\begin{frame}{A many-to-many example (output)}
+
+\end{frame}
+
+\begin{frame}{A many-to-many example (output)}
+
+\end{frame}
+
+\begin{frame}{A many-to-many example (output)}
+
+\end{frame}
+
+\begin{frame}{A many-to-many example (output)}
+
+\end{frame}
+
+\begin{frame}{A many-to-many example (output)}
+
+\end{frame}
+
+\begin{frame}{A many-to-many example (output)}
+
+\end{frame}
+
 \begin{frame}{Many-to-one-to-many example}
@@
 \item The decoder transforms this vector into a sorted list of numbers.
+\item For simplicity, the model was restricted to sequences of length 32
+\item the elements of the sequence were all betweeen 1 and 128
 \end{itemize}
@@
 
-   \begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=1cm]
-       \tikzstyle{arrow} = [thick,->,>=stealth]
+   \begin{tikzpicture}[shorten >=1pt, ->, draw=black!50, node distance=1cm]
+       \tikzstyle{arrow} = [thick, ->, >=stealth]
        \tikzstyle{startstop} = [rectangle, minimum width=1.5cm, minimum height=.6cm, text centered, draw=black, fill=black!20]
-       \node (input) [] at (0cm,1cm) {};
-       \node (output) [] at (0cm,-3cm) {};
-       \node (s-input-A) [] at (-2cm,0cm) {};
-       \node (s-output-A) [] at (10cm,0cm) {};
-       \node (s-input-B) [] at (-2cm,-1cm) {};
-       \node (s-output-B) [] at (10cm,-1cm) {};
-       \node (s-input-C) [] at (-2cm,-2cm) {};
-       \node (s-output-C) [] at (10cm,-2cm) {};
-       \path[] node[startstop] (A-0) at (0cm,0cm) {};
-       \path[] node[startstop] (B-0) at (0cm,-1cm) {};
-       \path[] node[startstop] (C-0) at (0cm,-2cm) {};
-       \draw [arrow] (input) -- (A-0);
+       \node (input) [] at (0cm, 1cm) {};
+       \node (output) [] at (0cm, -1cm) {};
+       \node (s-input-A) [] at (-2cm, 0cm) {};
+       \node (s-output-A) [] at (7.5cm, 0cm) {};
+       \path[] node[startstop] (A-0) at (0cm, 0cm) {};
+       \draw [arrow] (input) -- (A);
        \draw [arrow] (s-input-A) -- (A-0);
-       \draw [arrow] (s-input-B) -- (B-0);
-       \draw [arrow] (s-input-C) -- (C-0);
-       \draw [arrow] (A-0) -- (B-0);
-       \draw [arrow] (B-0) -- (C-0);
-       \draw [arrow] (C-0) -- (output);
+       \draw [arrow] (A) -- (output);
 
-        \foreach \i/\j in {1/0,2/1,3/2,4/3}
+        \foreach \i/\j in {1/0, 2/1, 3/2}
         {
-          \node (input-\i) [] at (2*\i cm,1cm) {};
-          \node (output-\i) [] at (2*\i cm,-3cm) {};
-          \path[] node[startstop] (A-\i) at (2*\i cm,0cm) {};
-          \path[] node[startstop] (B-\i) at (2*\i cm,-1cm) {};
-          \path[] node[startstop] (C-\i) at (2*\i cm,-2cm) {};
-          \draw [arrow] (A-\i) -- (B-\i);
-          \draw [arrow] (B-\i) -- (C-\i);
+          \node (input-\i) [] at (2*\i cm, 1cm) {};
+          \node (output-\i) [] at (2*\i cm, -1cm) {};
+          \path[] node[startstop] (A-\i) at (2*\i cm, 0cm) {};
           \draw [arrow] (input-\i) -- (A-\i);
           \draw [arrow] (A-\j) -- (A-\i);
-          \draw [arrow] (B-\j) -- (B-\i);
-          \draw [arrow] (C-\j) -- (C-\i);
-          \draw [arrow] (C-\i) -- (output-\i);
+          \draw [arrow] (A-\i) -- (output-\i);
         }
-        \draw [arrow] (A-4) -- (s-output-A);
-        \draw [arrow] (B-4) -- (s-output-B);
-        \draw [arrow] (C-4) -- (s-output-C);
+
+        \draw [arrow] (A-3) -- (s-output-A);
+
+        \node (d-input) [] at (0cm, -2cm) {};
+        \node (d-output) [] at (0cm, -4cm) {};
+        \node (d-s-input-A) [] at (-2cm, -3cm) {};
+        \node (d-s-output-A) [] at (7.5cm, -3cm) {};
+        \path[] node[startstop] (d-A-0) at (0cm, -3cm) {};
+        \draw [arrow] (d-input) -- (d-A-0);
+        \draw [arrow] (d-s-input-A) -- (d-A-0);
+        \draw [arrow] (d-A-0) -- (d-output);
+
+         \foreach \i/\j in {1/0, 2/1, 3/2}
+         {
+           \node (d-input-\i) [] at (2*\i cm, -2cm) {};
+           \node (d-output-\i) [] at (2*\i cm, -4cm) {};
+           \path[] node[startstop] (d-A-\i) at (2*\i cm, -3cm) {};
+           \draw [arrow] (d-input-\i) -- (d-A-\i);
+           \draw [arrow] (d-A-\j) -- (d-A-\i);
+           \draw [arrow] (d-A-\i) -- (d-output-\i);
+         }
+         \draw [arrow] (d-A-3) -- (d-s-output-A);
+
+         \draw [arrow] (output-3) -- (d-input);
 
@@
  \defverbatim[colored]\lstI{
- \begin{lstlisting}[language=Python,basicstyle=\ttfamily,keywordstyle=\color{red}]
-   SEQ_LENGTH = 256
-   E_DIM = 128
-   STATE_DIM = 512
-   N_LAYERS = 4
-
-   def inference():
+ \begin{lstlisting}[language=Python, basicstyle=\ttfamily, keywordstyle=\color{red}]
 model_input = tf.placeholder('uint8',
@@
 OP = tf.nn.dynamic_rnn
-       encoded_input, state = OP(encode,_,dtype=tf.float32)
-       Globals.encoder_output = state
-       with tf.variable_scope('decoder'):
-           training_decoder_input = tf.zeros_like(Globals.model_input)
-           _ = tf.one_hot(training_decoder_input, depth=E_DIM, axis=-1)
-           _ = tf.reshape(_, [-1, SEQ_LENGTH, E_DIM])
-           decode = multi_layer_rnn(N_LAYERS, STATE_DIM)
-           decoded_output, state = tf.nn.dynamic_rnn(decode, _,
-                                                     dtype=tf.float32,
-                                                     initial_state=state)
-           decoded_output = tf.reshape(decoded_output, [-1, STATE_DIM])
-           output = project(decoded_output, E_DIM)
-           out = tf.cast(tf.argmax(output, 1), tf.uint8)
-           out = tf.reshape(out, [-1, SEQ_LENGTH])
-           Globals.training_decoder_input = training_decoder_input
-           Globals.model_output = output
-           Globals.prediction = out
-           Globals.decoder = decode
-           Globals.decoder_input = _
+encoded_input, state = OP(encode, _, dtype=tf.float32)
+encoder_output = state
  \end{lstlisting}
  }
- \begin{frame}{Many-to-one-to-many example code}
+ \begin{frame}{Many-to-one-to-many example (code for encoder)}
    \lstI
@@
 
+ \defverbatim[colored]\lstI{
+ \begin{lstlisting}[language=Python, basicstyle=\ttfamily, keywordstyle=\color{red}]
+with tf.variable_scope('decoder'):
+   training_decoder_input = \
+          tf.zeros_like(Globals.model_input)
+   _ = tf.one_hot(training_decoder_input,
+                  depth=E_DIM, axis=-1)
+   _ = tf.reshape(_, [-1, SEQ_LENGTH, E_DIM])
+   decode = multi_layer_rnn(N_LAYERS, STATE_DIM)
+   OP = tf.nn.dynamic_rnn
+   decoded_output, state = OP(decode, _,
+                              dtype=tf.float32,
+                              initial_state=state)
+   decoded_output = tf.reshape(decoded_output,
+                               [-1, STATE_DIM])
+   output = project(decoded_output, E_DIM)
+   out = tf.argmax(output, 1)
+ \end{lstlisting}
+ }
+ \begin{frame}{Many-to-one-to-many example (code for decoder)}
+   \lstI
+ \end{frame}
+
+ \begin{frame}{A many-to-one-to-many example (output)}
+   \begin{itemize}
+   \item 39, 9, 113, 57, 39, 124, 125, 86, 65, 30, 39, 124, 98, 65, 27, 62, 79, 102, 96, 45, 96, 42, 54, 92, 33, 61, 14, 106, 89, 4, 61, 95;
+   \item 57, 6, 6, 6, 6, 6, 96, 96, 96, 96, 96, 96, 96, 96, 96, 96, 96, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118
+  \end{itemize}
+  \begin{itemize}
+   \item 22, 31, 58, 43, 96, 83, 95, 65, 17, 48, 65, 122, 94, 109, 113, 49, 74, 61, 39, 127, 3, 14, 107, 91, 55, 2, 108, 7, 119, 60, 32, 21;
+   \item 14, 14, 14, 17, 17, 6, 6, 6, 96, 96, 96, 96, 96, 96, 96, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118
+ \end{itemize}
+ \begin{itemize}
+   \item 117, 15, 77, 4, 57, 121, 108, 76, 72, 119, 41, 4, 94, 44, 103, 124, 120, 20, 79, 121, 68, 12, 31, 97, 93, 19, 45, 27, 119, 92, 119, 92;
+   \item 65, 65, 95, 95, 95, 96, 96, 96, 96, 96, 96, 96, 96, 96, 96, 96, 96, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118
+ \end{itemize}
+ \end{frame}
+
+ \begin{frame}{A many-to-one-to-many example (output)}
+ \begin{itemize}
+   \item 75, 6, 57, 57, 108, 34, 78, 71, 112, 115, 108, 48, 67, 1, 14, 9, 14, 115, 83, 62, 86, 91, 61, 40, 105, 92, 86, 84, 30, 84, 19, 107;
+   \item 1, 6, 9, 14, 14, 19, 0, 34, 40, 48, 57, 57, 61, 62, 67, 71, 75, 78, 83, 84, 84, 85, 86, 91, 92, 105, 107, 107, 109, 111, 114, 115
+ \end{itemize}
+ \begin{itemize}
+   \item 59, 108, 56, 66, 76, 38, 100, 61, 47, 79, 102, 15, 24, 92, 8, 26, 126, 43, 5, 90, 41, 2, 60, 85, 2, 104, 86, 40, 35, 47, 61, 91;
+   \item 2, 2, 5, 8, 15, 24, 26, 35, 38, 40, 41, 42, 47, 47, 56, 59, 60, 61, 61, 66, 76, 79, 84, 86, 89, 91, 92, 100, 102, 104, 108, 126
+ \end{itemize}
+ \begin{itemize}
+   \item 25, 50, 64, 4, 40, 47, 6, 14, 97, 32, 87, 103, 44, 25, 84, 40, 95, 13, 113, 66, 38, 79, 106, 40, 26, 16, 74, 50, 119, 32, 80, 16;
+   \item 4, 6, 13, 14, 16, 16, 25, 25, 26, 32, 32, 38, 39, 40, 40, 44, 47, 50, 50, 64, 66, 74, 79, 80, 84, 86, 95, 97, 103, 106, 113, 119
+ \end{itemize}
+ \end{frame}
+
+ \begin{frame}{A many-to-one-to-many example (output)}
+ \begin{itemize}
+   \item 96, 59, 77, 29, 39, 112, 23, 79, 60, 110, 97, 69, 107, 13, 96, 124, 2, 12, 55, 16, 106, 110, 30, 118, 119, 52, 22, 37, 113, 93, 73, 58;
+   \item 2, 12, 13, 16, 22, 23, 29, 0, 37, 39, 52, 55, 58, 59, 60, 69, 73, 77, 79, 93, 96, 96, 97, 106, 107, 109, 111, 111, 113, 118, 119, 124
+ \end{itemize}
+ \begin{itemize}
+   \item 44, 117, 104, 116, 111, 115, 58, 79, 8, 9, 1, 39, 112, 43, 64, 31, 126, 12, 36, 10, 93, 87, 40, 5, 108, 92, 11, 75, 113, 104, 64, 109;
+   \item 1, 5, 8, 9, 10, 11, 12, 0, 36, 39, 40, 0, 44, 58, 64, 64, 75, 79, 0, 92, 93, 104, 104, 108, 109, 111, 113, 113, 115, 116, 117, 126
+ \end{itemize}
+ \begin{itemize}
+   \item 58, 19, 10, 52, 22, 61, 67, 96, 3, 7, 116, 96, 54, 24, 19, 45, 127, 124, 11, 114, 53, 75, 126, 84, 122, 41, 75, 1, 119, 18, 92, 51;
+   \item 1, 3, 7, 10, 11, 18, 19, 19, 22, 24, 41, 45, 51, 52, 53, 54, 58, 61, 67, 75, 75, 84, 92, 96, 96, 114, 116, 119, 122, 124, 126, 127
+ \end{itemize}
+ \end{frame}
+
+ \begin{frame}{A many-to-one-to-many example (output)}
+ \begin{itemize}
+   \item 34, 73, 95, 91, 93, 108, 43, 75, 38, 70, 66, 40, 108, 127, 25, 94, 34, 26, 89, 23, 95, 43, 2, 54, 11, 19, 105, 52, 108, 77, 93, 86;
+   \item 2, 11, 19, 23, 25, 26, 34, 34, 38, 40, 0, 0, 52, 54, 66, 70, 73, 75, 77, 86, 89, 91, 93, 93, 94, 95, 95, 105, 107, 108, 108, 127
+ \end{itemize}
+ \begin{itemize}
+   \item 95, 106, 119, 69, 40, 41, 28, 114, 12, 1, 106, 87, 117, 78, 54, 37, 110, 24, 9, 114, 107, 87, 33, 76, 5, 90, 29, 14, 96, 109, 1, 3;
+   \item 1, 1, 3, 5, 9, 12, 14, 24, 28, 29, 33, 37, 40, 41, 54, 69, 76, 78, 86, 0, 90, 95, 96, 106, 106, 107, 109, 110, 114, 114, 117, 119
+ \end{itemize}
+ \begin{itemize}
+   \item 122, 97, 90, 61, 72, 66, 60, 60, 25, 125, 84, 73, 114, 46, 112, 76, 110, 62, 58, 34, 126, 124, 102, 35, 11, 100, 47, 113, 85, 64, 22, 89;
+   \item 11, 22, 25, 34, 35, 46, 47, 58, 60, 60, 61, 62, 64, 66, 72, 73, 76, 84, 85, 89, 90, 97, 100, 102, 110, 113, 113, 114, 122, 124, 125, 126
+ \end{itemize}
+ \end{frame}
+
+ \begin{frame}{A many-to-one-to-many example (output)}
+ \begin{itemize}
+   \item 69, 41, 40, 61, 9, 120, 37, 85, 103, 47, 81, 77, 126, 72, 86, 72, 6, 72, 92, 7, 22, 100, 127, 19, 89, 72, 62, 6, 64, 26, 45, 15;
+   \item 6, 6, 7, 9, 15, 19, 22, 26, 37, 40, 41, 45, 47, 61, 62, 64, 69, 72, 72, 72, 72, 77, 81, 85, 86, 89, 92, 100, 103, 120, 126, 127
+ \end{itemize}
+ \begin{itemize}
+   \item 95, 40, 116, 120, 68, 75, 24, 114, 39, 81, 52, 37, 76, 10, 15, 105, 79, 15, 44, 89, 115, 5, 26, 11, 85, 53, 85, 114, 39, 29, 34, 16;
+   \item 5, 10, 11, 15, 15, 16, 24, 26, 29, 34, 37, 39, 39, 40, 44, 52, 53, 68, 75, 76, 79, 81, 85, 85, 89, 95, 105, 114, 114, 115, 116, 120
+ \end{itemize}
+ \begin{itemize}
+   \item 20, 61, 78, 15, 61, 8, 108, 36, 85, 96, 40, 80, 106, 24, 66, 82, 96, 43, 126, 33, 80, 116, 35, 86, 98, 81, 76, 89, 6, 103, 117, 16;
+   \item 6, 8, 15, 16, 20, 24, 33, 35, 36, 40, 0, 61, 61, 66, 76, 78, 80, 80, 81, 82, 85, 86, 89, 96, 96, 98, 103, 106, 108, 116, 117, 126
+ \end{itemize}
+ \end{frame}
 \end{document}

### 2017-08-31 deeplearn/presentation/presentation_slides.tex
Presentation slides

@@
 \begin{itemize}
-  \item A neural network is a fancy way to describe a class of functions $f:\R^n\to \R^m$ parametrized by $\R^\ell$.
-  \item The layers of a NN represent function composition.
-  \item If $f$ is a function $F(\x, \w):\R^n\times \R^\ell\to\R^n$ such that $\C=\{\x\mapsto F(\x, \w):\w\in\R^n\}$.
+  \item The task is to model functions $f:A\to B$, where $A$ and $B$ are sets, using functions $\R^n\to\R^m$.
+  \item A neural network is a fancy way to describe a class of functions $f:\R^n\to \R^m$ parametrized by $\R^\ell$, in the sense that there is a function $F(\x, \w):\R^n\times \R^\ell\to\R^n$ such that $\C=\{\x\mapsto F(\x, \w):\w\in\R^n\}$.
   \item Each vector $\w\in\R^\ell$ is in some sense a code for an element of $\C$.
   \item When we are given a function $f:\R^n\to \R^m$, we try to find which element of $\C$ best represents $f$.
-
 %  \item That is to say, find $\w$ such that
@@
 %\item In particular, we will me making use of the iterative nature of the construction $A^*$ and $B^*$.
-\item Here is the simplest example: if $A=B$, and $f:A\to A$ is a function, then we can iterate $f$: $f^0(a)=a$ for every $a\in A$, and $f^{n+1}(a) = f(f^n(a))$.
+\item Here is the simplest example: if $A=B$, and $f:A\to A$ is a function, then we can iterate $f$:
+\begin{itemize}
+  \item $f^0(a)=a$ for every $a\in A$,
+  \item $f^{n+1}(a) = f(f^n(a))$.
+\end{itemize}
 We terminate the iteration at the first value of $n$ for which $f^n(a)=f^{n-1}(a)$ (note that there is no guarantee that this will ever happen).
@@
       \node (s-input-A) [] at (-2cm, 0cm) {};
-      \node (s-output-A) [] at (10cm, 0cm) {};
+      \node (s-output-A) [] at (9.5cm, 0cm) {};
       \node (s-input-B) [] at (-2cm, -1cm) {};
-      \node (s-output-B) [] at (10cm, -1cm) {};
+      \node (s-output-B) [] at (9.5cm, -1cm) {};
       \node (s-input-C) [] at (-2cm, -2cm) {};
-      \node (s-output-C) [] at (10cm, -2cm) {};
+      \node (s-output-C) [] at (9.5cm, -2cm) {};
       \path[] node[startstop] (A-0) at (0cm, 0cm) {};
@@
 \begin{frame}{A many-to-many example (output)}
-
+%\begin{verbatim}
+%  W =YL nuf rs t Cgi3uaa nemk+ymeK  tWb elhr inseae iu  g (.
+  Faewn tron ooaes  rl n   u i o .or3itkdieesatint  bt u ahHhkoenn
+  etra+uHtiebnndrlyht sato  ioom y  n m,ee.soon ol e0  ,y a hnreliy?e
+  wree nf tlVthyi  oei oloynea .h.g8wtnsekhw r. ow AleiasfrhoeiryeswB
+  .ogor H ael,gol m Rno sn aa  galrm er o de3surr asadeatso  detn %
+  aml hrwedop orl aoirltsrggos,a  rnaurlewyYe". eld.k ujdt<ad losd\^roe
+  msnr.tou ods rr L-hieeee eagug oo5si y 6setS  osg e .d f dmd e p\^mraohe.
+  iaroh.let  etshdsseniolthg emeyndt neotg ue h iy hunthdg !me  oeA oer
+  te sl h t mnmdi t da3nfnspL>lndrr yhlehee mX  olaPfgdoye,l thot m.gitlc
+  eegn gseo(gn> Btehnnuasn eiffbggi ud  r nfet.m ri!iokuCeou  ?j
+  uraiAuuhwoslb|k  iho aeat- k y   . m neBnh.p ulhosah . vusLesagshio e o
+  alyeea rrae   sams   y phsu as  iaey ka ert   raetl1s  o m hweirLz/a
+  mldordrlehdphtr !eal  aawaoi,slrmdhooieit  rdath dhe r uohtn  giu hl  ,eo
 \end{frame}
@@
 \begin{frame}{A many-to-many example (output)}
-
+  st best neeoa, bet hublen  whin nevents, Yebear towhrath lit Goed P orrssss H
+  int. Revell, but hav ed asak MudgSe Poo
+  Marry hooling olkint of the torferyfwerdnole opotger as  Snd dot fufped a rivenad with a hedd Swering ald end the fose l had tra of oss tobe.
+Lhoush dame, haik  Dumbledore, bhents wayens is couldnbacking the tavirg ou at fre with if comelt intiVcrweroughtsit whed tagrtadh teer EontY Baid joom han sanledoo to falfin.
+Mrd hid you steedhesed that asd and Goomemis,.
+If save ti t  yeak, and ulststod, pay Chouehhed.
+I sswvect mech.
+makd thene whost well, shead Duwpledoalmingverya thin, will.houco itthatuardryd tle freled to prope ly bndchy bgrid gowntiing gronsidocw ng thes uthim mame f atd reeed the  wrye,gou deyer fonnt ere aarn whatsebyihe haugh.
+gedling his ames.diain.
+Bunnea oleyk.
+Herrioned
+  AIn, Prldy.fors.ead Monfurfry.ars.. Lur Mames. The le arnay
 \end{frame}
@@
 \begin{frame}{A many-to-many example (output)}
+  Yes, said Phich.
+I seem d tou.e,in t it.
+...
+Harry could not understracked on the sirlow quite.
+YoteirSh?
+Masi4  easley was tweir baaring a swarm onss o oy shem.
+Ho your veice!
+sr7 Mrs  WOr!!eIll saact ye as i tomgo enhuge f.
+.
+.
+is to have teen his siorys faked thobonsto tarelys pld the board to have seennttered th seen.
+Harry had gnven the dignaes oonce, tf ce was turning, mattering, irs dmorning your way at t tell only time  nankl?uldur ofntabally all about it, after a luot eo  inclneriitcr ed.
+How other ever tep ga se, ao  But   It Harlid, said The Soivers.
+.
+with aimioh ey iould had for them.to!
+Yus, iaid TBattern he mired  ever sackbd fandinred clldye  copfles, te aidenly told as averying to be seet but she was curling ilougain.t a  the ske was png khen because im inteirsion.
+O.
+love hes lef ...
+Harry said acgedally, his newdecime closs and hidding as he spoke, you right to use of if of cheverdiescand mot,h, haypened tuffleng her itions andly at horrgalm.
 
@@
 \begin{frame}{A many-to-many example (output)}
-
+  A fartoblind, shouing in coptless, rutching lis glasses alive with a tightly rhrugting.
+Fartyeys in the tibrage serched on green like iktin the wandow, crossed with hoggant.ba warrowes rkling with dhheul  and pon sever broatures, which was carved again frum the age of the glasm too.
+
+  There wase risonab look at thi splattday, eaen on the blurd aobodyheasal hed back ao a marmor tfualive with slacklines  o
+  Before that  ac will, sery mered, Fasy so  ln ihe staff table, and then  Yidmione sxploded, aan uring thi  no. But Dhe , whele y
 \end{frame}
@@
 \begin{frame}{A many-to-many example (output)}
-
+Filch ead eere with all tis hneethighb hair, seemed as hi had musmereded faxh apport ae Kruacher had the Slytherin house of tets before.yesterday, Rind how to see, great, and taure he llfore, that weth nded stay in the rormer Slugee Yoar?
+No  unknown, nlytherin, Hf ter go nowhing that , bixing   .ouve jHished to losk at himest, Hermione, George.
+Neledaayest Narcissa,in built.
+on theimsddle of my while.
+Albus, he  ...
+liveng and safeer ers oime fiustrw ..
+good fabily., or came f...
+Your famher roneared at he feared her wa rinht dowr behind the  dark and every step, their terrified sealing and
 \end{frame}
@@
 \begin{frame}{A many-to-many example (output)}
+  Theres no doubt to kateh  Harry, dont bet er toke toe naght he reckons the wayds can surprise Inean halp that icdobd side mn  those ehit apd Aunt Petenia Allum  jusped up Doth tis wands.
+Tre not anabl a Distinctly  I detds ed,miggry of knoph tifough ahfunl still boy nhat whin he pas comfrowling ay murder than usual.
+In siarel mistress in Doagon Illey ...
+thet wy subrotrs,had been giing and marious.teat sange  whet  aust betliout, said Ron nuietly, but he did not lant to have to fulded as he picked back.
 
@@
 \begin{frame}{A many-to-many example (output)}
-
+  Oh are stay-hamfered myny memold? Professor M Great Holl avice is shwpshelves  said Lupin, and shr knew out. H
+  g soudents faster: Treaicized aunitius!
+Hasmione!lives to she common got the Indie with myself?
+Who is a wording, Malfoy.
+as t-ttinhtto tell him that the stubents wll with Snape anybody oIll have ior Siriul the caad tnyohere, Ill belceve egomr  Inter Bllba rust Copy,of Vus Charms and yhan-ed Ocdonagss harrible  onceusl.
+Prowessor?
+Yaid a creaty ahirl and thr blonked ond powerfulsy.
+There was an adgiliai and the back of the two hide, then rinilumblid, and up the srmcktfo cans  flel down to the kead.
 \end{frame}
@@
 \begin{frame}{A many-to-many example (output)}
-
+  Traco Malfoy, well, and you add the Min, they go  Cho back indo Lur strits, gripping gaolies around its, aleared lood.
+They presaed no prefects for that, sho hoot ...
+he knepsda lhe tinistry least until I think,i You oont trink hesdid.
+.
+Looking kiidly with ut multipe serore where wrountsyou  hhat came to itvola yours l We can finds, anl right?.Horccux as furious by awrid brooer, fin you  Aaid Hateilita, lof him hut what he thinks it, you wesder any time theyd have lesven to do.
+Nice Hnd,lasten to me.
+And yeure going fon thelferhenks, have, said Mu.
+Weasley bliracciusty, as it had became very reliev, bugay n  a powntiof giren sp echo dpeliher.tents.
+He was listening sor.
+It was Dembledore desire.
+He had missed lu .
+He turned his foer never held bytween the sige, the huftled expluseli needr that followed.
 \end{frame}
@@
 \begin{frame}{A many-to-many example (output)}
-
+  Oh, he said gruntingly. N
+  Teel me somethn your oire, made it ttained from Hogwadts  ae doea, you saemed to be dropping Uis at Harry  hed Head .
+.
+.
+that he das !
+but Dobby is has faet, readuti, sireaking after a mans cauttleng  theyd hald to vosmentagainst tige of a Quifditch  alarming!y.
+Toeyre doing an earthy yaurs  And by name ac alr, you know .t Harrys lying thin s  hom mummerook!
+Maybe you wirldnt chugh eventmones later, Po two, did you?
+said Harry with his eyeb with flulf,rsnnd-solnger,lunk away at Sim.
 \end{frame}
@@
 \begin{frame}{A many-to-many example (output)}
-
+  arry could not look quickly at his hustened,pnd stile barsting with his hwig apen?
+Thatswat hing Umbridge the rnx saggins if you taen startanders, thay po it, sor knod.that  I think you  in iould be l bat an ahothcle at the weardnc term,at find Rim   Blat on prypticess not,to take a viry chocolate, take uhes roomed bine on thes The.
+Harry had left the stairs toomention, but Harry interreesed the sort of reaih was it,
 \end{frame}
@@
 \begin{frame}{A many-to-many example (output)}
-
-\end{frame}
-
-\begin{frame}{A many-to-many example (output)}
-
+  Varry shurted oot over to find Fudges oyes with hir pool eaery direction.
+You know where hes worth dirtction yeh,wuth mere moubtful upon the days.
+Iow may e ...
+but not hintless.
+...
+Shes strong,y uven losdlr, Professor, I hve look bank  gut it will remember wha fine- steeds to listen .
+.
+.
+blew an!and alr Mystlound vanished about Malfoy g indisaiing time.
+Im sorry th want to hear the Ministry,ofispers.to aonfll eavisdormitory, itswoll even if you  wanteto have home.
+A  peyed  very surpas off   Si  struxed th a haov and mede iim hn arinaelery distracled  Itmdonld numbil centaminate the katwien he himself.
+Harry, Ron, and Hermione, the okly oeniis lizarding moving ps their study that hhownd aany sotus of powdereathe faont paggent so tHa his speak aentttoon the higlstryc  igh a brocm, and none of them  ould seem sick lant temrh Itnever kmen in curse  fact, wher it somt of d whons gettling through their airection ttab with a thue issuses.
+I never enew you went every me and finalive rnd!
+He wasnt sngnified, said Harry.
 \end{frame}

### 2017-09-05 deeplearn/presentation/presentation_slides.tex
moved presentation

@@
   \begin{itemize}
-  \item Recurrent neural networks are just a prescription for a class of functions of the form $f:\R^n\times \R^\ell\times\R^m\to \R^\ell\times \R^m$, where $\R^m$ is the space of parameters.
+  \item Recurrent neural networks are just a prescription for a class of functions of the form $$f:\R^n\times \R^\ell\times\R^m\to \R^\ell\times \R^m$$ where $\R^m$ is the space of parameters.\pause
   \item The set $\{0, 1\}^*$ of all finite sequences of $0$'s and $1$'s is countable,
-  \item and $\R^\ell$, is a real vector space,
-  is uncountable.
-  \item Therefore, we can encode every element $w\in\{0, 1\}^*$ as a vector in $S$.  Informally,
+  \item and $\R^\ell$, is uncountable.
+  \item We can encode every element $w\in\{0, 1\}^*$ as a vector in $S$.  Informally,
   $\R^\ell$ is enough to encode any finite state space, and every possible value for the content of the tape of a Turing machine. We get:

### 2017-09-05 deeplearn/twitter_sentiment/scripts/count_chars.py
Database character counting

@@ -1,0 +1,31 @@
+import os
+import sys
+import pymysql
+import numpy as np
+
+ignore_bugs = '--ignore-bugs' in sys.argv
+
+# connection to cms database
+cms_connection = pymysql.connect(host='localhost', user='root', password='root',
+                                 db='sentiment_analysis_data', charset='utf8mb4',
+                                 cursorclass=pymysql.cursors.DictCursor)
+
+
+
+sql = """SELECT sanitized_text FROM neuronet_training_dataset"""
+
+chrs = {}
+with cms_connection.cursor() as cursor:
+    cursor.execute(sql)
+    data = cursor.fetchall()
+    for t in data:
+        text = t['sanitized_text']
+        for ch in text:
+            if ch not in chrs:
+                chrs[ch] = 0
+            chrs[ch] += 1
+    print(len(chrs))
+    print (chrs)
+    chs = sorted(chrs.keys(), key=lambda x:chrs[x])
+    for i, x in enumerate(chs):
+        print(i, x, '   ;', x.encode('utf8'), ord(x), chrs[x])

### 2017-09-05 deeplearn/twitter_sentiment/scripts/update_sentiment_training_table.py

@@ -111,9 +111,9 @@ def populate_table(table_name, source_query, source_connection, target_connectio
             target_connection.commit()
 
 if ignore_bugs:
-    training_table = 'neuronet_training_dataset'
-else:
     training_table = 'neuronet_training_dataset_no_bug'
+else:
+    training_table = 'neuronet_training_dataset'
 
 def fetch_cms_data(table_name):
     populate_table(table_name=table_name,
@@ -157,12 +157,12 @@ def shuffle_table_rows(table_name, connection):
         connection.commit()
 
 
-def truncate_table(table_name):
+def truncate_table(table_name, connection):
     with connection.cursor() as cursor:
         sql = """TRUNCATE {table_name}""".format(table_name=table_name)
         cursor.execute(sql)
 
-truncate_table(table_name)
+truncate_table(training_table, buzzometer_dev_connection)
 fetch_cms_data(training_table)
 fetch_buzzometer_data(training_table)
 shuffle_table_rows(training_table, buzzometer_dev_connection)
